---
title: "AI and the Labour Market in Croatian Media"
subtitle: "Media Framing Analysis (2021--2024)"
author: "Luka Sikic"
date: today
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    code-fold: true
    code-tools: true
    code-summary: "Show code"
    df-print: paged
    fig-width: 10
    fig-height: 6
    fig-dpi: 300
    embed-resources: true
execute:
  warning: false
  message: false
  echo: true
---

```{r}
#| label: setup
#| include: false

# ==============================================================================
# SETUP — load helpers, config, packages
# ==============================================================================

source("R/00_helpers.R")

load_packages(c(
  "dplyr", "tidyr", "stringr", "stringi", "lubridate", "forcats", "tibble",
  "ggplot2", "scales", "patchwork", "ggrepel",
  "knitr", "kableExtra",
  "quanteda", "quanteda.textstats",
  "tidytext"
))

options(dplyr.summarise.inform = FALSE, scipen = 999)

# Custom theme
theme_report <- theme_minimal(base_size = 12) +
  theme(
    plot.title    = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(color = "gray40", size = 11),
    legend.position  = "bottom",
    panel.grid.minor = element_blank(),
    strip.text       = element_text(face = "bold")
  )
theme_set(theme_report)

# Color palettes
frame_colors <- c(
  "JOB_LOSS"        = "#e41a1c",
  "JOB_CREATION"    = "#4daf4a",
  "TRANSFORMATION"  = "#ff7f00",
  "SKILLS"          = "#377eb8",
  "REGULATION"      = "#984ea3",
  "PRODUCTIVITY"    = "#f781bf",
  "INEQUALITY"      = "#a65628",
  "FEAR_RESISTANCE" = "#999999",
  "NONE"            = "gray80"
)
```

# Abstract {.unnumbered}

This document analyses how Croatian digital media frame the intersection of artificial intelligence and the labour market. The corpus is drawn from the Determ media-monitoring platform (2021--2024) and includes only articles containing both AI-related and labour-market-related keywords (intersection logic). Eight interpretive frames, six actor categories, and automated sentiment are examined across time and media outlet types.

::: {.callout-note}
## Data preparation
The corpus was extracted by `R/01_extract_corpus.R` and enriched by `R/02_add_diagnostics.R`. Run those scripts first if the data files are missing.
:::

# Introduction

## Background

Technological change has historically reshaped labour demand in ways that were difficult to predict at the time. Artificial intelligence poses a distinctive challenge because it affects not only routine manual tasks but also non-routine cognitive work such as programming, translation, data analysis, and creative writing.

Media do not merely relay facts; they select, omit, and emphasise, thereby actively shaping public perception. The way media frame the impact of AI on work can influence individual behaviour, corporate strategy, and public policy.

## Research questions

**RQ1 (Volume and dynamics)** How much media coverage does the AI--labour nexus receive, and how has coverage evolved over time?

**RQ2 (Frames)** Which interpretive frames dominate, and how has their prevalence shifted?

**RQ3 (Actors)** Whose perspectives appear most frequently?

**RQ4 (Sources)** Do different media outlet types frame the topic differently?

# Data

## Loading the corpus

```{r}
#| label: load-corpus

if (!file.exists(path_raw_corpus)) {
  stop("Corpus not found at: ", path_raw_corpus,
       "\n\nRun R/01_extract_corpus.R first.")
}

corpus_data <- readRDS(path_raw_corpus)

cat("Corpus loaded successfully\n")
cat("Total articles:", format(nrow(corpus_data), big.mark = ","), "\n")
cat("Date range:", as.character(min(corpus_data$DATE)), "to",
    as.character(max(corpus_data$DATE)), "\n")
cat("Columns:", ncol(corpus_data), "\n")
```

```{r}
#| label: corpus-overview

corpus_data$.text_lower <- stri_trans_tolower(
  paste(coalesce(corpus_data$TITLE, ""),
        coalesce(corpus_data$FULL_TEXT, ""),
        sep = " ")
)

if (!"year" %in% names(corpus_data)) {
  corpus_data$year <- year(corpus_data$DATE)
}
if (!"year_month" %in% names(corpus_data)) {
  corpus_data$year_month <- floor_date(corpus_data$DATE, "month")
}
if (!"quarter" %in% names(corpus_data)) {
  corpus_data$quarter <- quarter(corpus_data$DATE)
  corpus_data$year_quarter <- paste0(corpus_data$year, " Q", corpus_data$quarter)
}

corpus_data$word_count <- stri_count_regex(corpus_data$FULL_TEXT, "\\S+")

summary_stats <- tibble(
  Metric = c("Total articles", "Unique sources", "Date range",
             "Mean words/article", "Median words/article"),
  Value = c(
    format(nrow(corpus_data), big.mark = ","),
    format(n_distinct(corpus_data$FROM), big.mark = ","),
    paste(min(corpus_data$DATE), "to", max(corpus_data$DATE)),
    round(mean(corpus_data$word_count, na.rm = TRUE)),
    round(median(corpus_data$word_count, na.rm = TRUE))
  )
)

kable(summary_stats, col.names = c("Metric", "Value")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

## Source distribution

```{r}
#| label: source-distribution

if ("SOURCE_TYPE" %in% names(corpus_data)) {
  source_dist <- corpus_data |>
    count(SOURCE_TYPE, sort = TRUE) |>
    mutate(pct = round(n / sum(n) * 100, 1))

  kable(source_dist, col.names = c("Source type", "N", "%")) |>
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
}
```

```{r}
#| label: fig-top-sources
#| fig-cap: "Top 25 sources by article count"
#| fig-height: 7

top_sources <- corpus_data |>
  count(FROM, sort = TRUE) |>
  head(25)

ggplot(top_sources, aes(x = reorder(FROM, n), y = n)) +
  geom_col(fill = "#2c7bb6", alpha = 0.8) +
  coord_flip() +
  labs(title = "Top 25 sources", x = NULL, y = "Articles")
```

# Temporal dynamics

## Monthly volume

```{r}
#| label: fig-monthly-volume
#| fig-cap: "Monthly article volume with key events"
#| fig-height: 5

monthly_volume <- corpus_data |>
  count(year_month) |>
  filter(!is.na(year_month))

events_cfg <- CONFIG$events
events <- tibble(
  date  = as.Date(sapply(events_cfg, `[[`, "date")),
  label = sapply(events_cfg, `[[`, "label"),
  y_pos = max(monthly_volume$n, na.rm = TRUE) * seq(0.95, by = -0.10,
                                                     length.out = length(events_cfg))
)

ggplot(monthly_volume, aes(x = year_month, y = n)) +
  geom_col(fill = "#2c7bb6", alpha = 0.7) +
  geom_smooth(method = "loess", se = TRUE, color = "#d7191c", linewidth = 1) +
  geom_vline(data = events, aes(xintercept = date),
             linetype = "dashed", color = "gray40") +
  geom_label(data = events, aes(x = date, y = y_pos, label = label),
             size = 3, fill = "white", alpha = 0.9) +
  scale_x_date(date_breaks = "3 months", date_labels = "%b\n%Y") +
  labs(title = "AI and labour market coverage",
       subtitle = "Monthly volume with LOESS trend and key events",
       x = NULL, y = "Articles")
```

## Yearly volume

```{r}
#| label: fig-yearly-volume
#| fig-cap: "Annual article volume"

yearly_volume <- corpus_data |>
  count(year) |>
  filter(!is.na(year))

ggplot(yearly_volume, aes(x = factor(year), y = n)) +
  geom_col(fill = "#2c7bb6", alpha = 0.8) +
  geom_text(aes(label = format(n, big.mark = ",")), vjust = -0.5) +
  labs(title = "Annual volume", x = "Year", y = "Articles") +
  expand_limits(y = max(yearly_volume$n) * 1.1)
```

# Frame analysis

## Frame dictionaries

```{r}
#| label: frame-dictionaries

# Build dictionaries from config
frame_dictionaries <- lapply(CONFIG$frames, `[[`, "keywords")

frame_summary <- tibble(
  Frame       = names(frame_dictionaries),
  Description = sapply(CONFIG$frames, `[[`, "description"),
  Keywords    = sapply(frame_dictionaries, length)
)

kable(frame_summary, col.names = c("Frame", "Description", "N keywords")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

## Frame detection

```{r}
#| label: detect-frames

for (frame_name in names(frame_dictionaries)) {
  pattern <- paste(frame_dictionaries[[frame_name]], collapse = "|")
  corpus_data[[paste0("frame_", frame_name)]] <- stri_detect_regex(
    corpus_data$.text_lower, pattern
  )
}

frame_cols <- paste0("frame_", names(frame_dictionaries))
frame_counts <- corpus_data |>
  summarise(across(all_of(frame_cols), sum, na.rm = TRUE)) |>
  pivot_longer(everything(), names_to = "frame", values_to = "count") |>
  mutate(
    frame = str_remove(frame, "frame_"),
    pct   = round(count / nrow(corpus_data) * 100, 1)
  ) |>
  arrange(desc(count))

kable(frame_counts, col.names = c("Frame", "Articles", "% corpus")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

```{r}
#| label: fig-frame-distribution
#| fig-cap: "Frame prevalence"
#| fig-height: 5

ggplot(frame_counts, aes(x = reorder(frame, count), y = count, fill = frame)) +
  geom_col(alpha = 0.8) +
  geom_text(aes(label = paste0(pct, "%")), hjust = -0.1, size = 3.5) +
  coord_flip() +
  scale_fill_manual(values = frame_colors, guide = "none") +
  labs(title = "Interpretive frame prevalence",
       subtitle = "Percentage of articles containing frame keywords",
       x = NULL, y = "Articles") +
  expand_limits(x = max(frame_counts$count) * 1.15)
```

## Frame evolution

```{r}
#| label: fig-frame-evolution
#| fig-cap: "Frame evolution over time"
#| fig-height: 7

frame_monthly <- corpus_data |>
  group_by(year_month) |>
  summarise(
    n_total = n(),
    across(all_of(frame_cols), sum, na.rm = TRUE)
  ) |>
  pivot_longer(cols = all_of(frame_cols),
               names_to = "frame", values_to = "count") |>
  mutate(
    frame = str_remove(frame, "frame_"),
    pct   = count / n_total * 100
  ) |>
  filter(!is.na(year_month))

ggplot(frame_monthly, aes(x = year_month, y = pct, color = frame)) +
  geom_line(linewidth = 0.8, alpha = 0.8) +
  geom_smooth(method = "loess", se = FALSE, linewidth = 1.2, linetype = "dashed") +
  facet_wrap(~frame, ncol = 2, scales = "free_y") +
  scale_color_manual(values = frame_colors, guide = "none") +
  scale_x_date(date_breaks = "6 months", date_labels = "%b\n%Y") +
  labs(title = "Frame evolution",
       subtitle = "Monthly share of articles per frame with LOESS trend",
       x = NULL, y = "% articles") +
  theme(axis.text.x = element_text(size = 8))
```

# Actor analysis

```{r}
#| label: actor-detection

actor_dictionaries <- CONFIG$actors

for (actor_name in names(actor_dictionaries)) {
  pattern <- paste(actor_dictionaries[[actor_name]], collapse = "|")
  corpus_data[[paste0("actor_", actor_name)]] <- stri_detect_regex(
    corpus_data$.text_lower, pattern
  )
}

actor_cols <- paste0("actor_", names(actor_dictionaries))
actor_counts <- corpus_data |>
  summarise(across(all_of(actor_cols), sum, na.rm = TRUE)) |>
  pivot_longer(everything(), names_to = "actor", values_to = "count") |>
  mutate(
    actor = str_remove(actor, "actor_"),
    pct   = round(count / nrow(corpus_data) * 100, 1)
  ) |>
  arrange(desc(count))

kable(actor_counts, col.names = c("Actor", "Articles", "% corpus")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

```{r}
#| label: fig-actor-distribution
#| fig-cap: "Actor prevalence"

ggplot(actor_counts, aes(x = reorder(actor, count), y = count)) +
  geom_col(fill = "#377eb8", alpha = 0.8) +
  geom_text(aes(label = paste0(pct, "%")), hjust = -0.1, size = 3.5) +
  coord_flip() +
  labs(title = "Actor prevalence", x = NULL, y = "Articles") +
  expand_limits(x = max(actor_counts$count) * 1.15)
```

# Outlet classification

```{r}
#| label: outlet-classification

outlet_cfg <- CONFIG$outlet_types

corpus_data$outlet_type <- "Other"
for (type_name in names(outlet_cfg)) {
  for (pat in outlet_cfg[[type_name]]) {
    matches <- stri_detect_regex(stri_trans_tolower(corpus_data$FROM), pat)
    corpus_data$outlet_type[matches] <- type_name
  }
}

outlet_dist <- corpus_data |>
  count(outlet_type, sort = TRUE) |>
  mutate(pct = round(n / sum(n) * 100, 1))

kable(outlet_dist, col.names = c("Outlet type", "N", "%")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

```{r}
#| label: fig-frames-by-outlet
#| fig-cap: "Frame prevalence by outlet type"
#| fig-height: 6

min_articles <- CONFIG$analysis$min_articles_per_outlet

frames_by_outlet <- corpus_data |>
  group_by(outlet_type) |>
  summarise(
    n = n(),
    across(all_of(frame_cols), ~sum(.x, na.rm = TRUE) / n() * 100)
  ) |>
  filter(n >= min_articles) |>
  pivot_longer(cols = all_of(frame_cols),
               names_to = "frame", values_to = "pct") |>
  mutate(frame = str_remove(frame, "frame_"))

ggplot(frames_by_outlet, aes(x = frame, y = pct, fill = outlet_type)) +
  geom_col(position = "dodge", alpha = 0.8) +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Frame prevalence by outlet type",
       x = NULL, y = "% articles", fill = "Outlet type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Sentiment analysis

```{r}
#| label: sentiment-check

if ("AUTO_SENTIMENT" %in% names(corpus_data)) {
  sentiment_dist <- corpus_data |>
    count(AUTO_SENTIMENT, sort = TRUE) |>
    mutate(pct = round(n / sum(n) * 100, 1))

  kable(sentiment_dist, col.names = c("Sentiment", "N", "%")) |>
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
} else {
  cat("AUTO_SENTIMENT column not available in corpus.\n")
}
```

```{r}
#| label: fig-sentiment-time
#| fig-cap: "Sentiment distribution over time"
#| fig-height: 5

if ("AUTO_SENTIMENT" %in% names(corpus_data)) {
  sentiment_monthly <- corpus_data |>
    filter(!is.na(AUTO_SENTIMENT) & !is.na(year_month)) |>
    count(year_month, AUTO_SENTIMENT) |>
    group_by(year_month) |>
    mutate(pct = n / sum(n)) |>
    ungroup()

  ggplot(sentiment_monthly, aes(x = year_month, y = pct, fill = AUTO_SENTIMENT)) +
    geom_area(alpha = 0.7) +
    scale_fill_manual(values = c(
      "positive" = "#4daf4a", "Positive" = "#4daf4a",
      "neutral"  = "gray60",  "Neutral"  = "gray60",
      "negative" = "#e41a1c", "Negative" = "#e41a1c"
    )) +
    scale_y_continuous(labels = scales::percent) +
    scale_x_date(date_breaks = "3 months", date_labels = "%b\n%Y") +
    labs(title = "Sentiment over time",
         x = NULL, y = "Share", fill = "Sentiment")
}
```

# Sample articles

```{r}
#| label: sample-articles

set.seed(CONFIG$analysis$seed)
sample_articles <- corpus_data |>
  slice_sample(n = min(30, nrow(corpus_data))) |>
  select(DATE, TITLE, FROM) |>
  arrange(DATE)

kable(sample_articles, col.names = c("Date", "Title", "Source")) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE, font_size = 11) |>
  scroll_box(height = "400px")
```

# Summary

```{r}
#| label: summary-stats

n_articles    <- nrow(corpus_data)
date_range    <- paste(min(corpus_data$DATE), "to", max(corpus_data$DATE))
n_sources     <- n_distinct(corpus_data$FROM)
top_frame     <- frame_counts$frame[1]
top_frame_pct <- frame_counts$pct[1]

cat("ANALYSIS SUMMARY\n")
cat("================\n\n")
cat("Total articles:", format(n_articles, big.mark = ","), "\n")
cat("Period:", date_range, "\n")
cat("Sources:", format(n_sources, big.mark = ","), "\n")
cat("Dominant frame:", top_frame, "(", top_frame_pct, "% of articles)\n")
```

# Data export

```{r}
#| label: export
#| eval: false

export_data <- corpus_data |> select(-`.text_lower`)
saveRDS(export_data, path_analysed_corpus)
```

# Session info

```{r}
#| label: session-info

sessionInfo()
```
