---
title: "ChatGPT as an Exogenous Shock"
subtitle: "Media Coverage of AI's Labour-Market Implications in Croatia, 2021–2026"
author:
  - name: Petra Palić
    affiliations:
      - name: Croatian Catholic University
        department: University Department of Sociology
  - name: Luka Šikić
    affiliations:
      - name: Croatian Catholic University
        department: University Department of Communication Studies
  - name: Antea Barišić
    affiliations:
      - name: University of Zagreb
        department: Faculty of Economics and Business
date: today
bibliography: ../references.bib
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    code-fold: true
    code-tools: true
    code-summary: "Show code"
    df-print: paged
    fig-width: 10
    fig-height: 6
    fig-dpi: 300
    embed-resources: true
    include-in-header:
      text: |
        <style>.figure-note { font-size: 0.85em; color: #444; margin-top: -0.5em; margin-bottom: 1.5em; }</style>
execute:
  warning: false
  message: false
  echo: false
---

```{r}
#| label: setup
#| include: false

source("00_helpers.R")

required_packages <- c(
  "dplyr", "tidyr", "stringr", "stringi", "lubridate", "forcats", "tibble",
  "ggplot2", "scales", "patchwork", "ggrepel",
  "knitr", "kableExtra",
  "fixest",        # fixed-effects estimation
  "sandwich",      # HAC standard errors
  "lmtest",        # coefficient tests
  "strucchange",   # structural break detection
  "MASS",          # negative binomial
  "broom",         # tidy model output
  "zoo"            # rolling means
)
load_packages(required_packages)
select <- dplyr::select

options(dplyr.summarise.inform = FALSE, scipen = 999)
invisible(Sys.setlocale("LC_TIME", "C"))   # English month labels

theme_pub <- theme_classic(base_size = 11) +
  theme(
    plot.title       = element_blank(),
    plot.subtitle    = element_blank(),
    legend.position  = "bottom",
    legend.title     = element_blank(),
    panel.grid.major = element_line(color = "gray90", linewidth = 0.3),
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    strip.text       = element_text(face = "bold", size = 10),
    axis.title       = element_text(size = 11),
    axis.text        = element_text(color = "black")
  )
theme_set(theme_pub)

# Table with a short caption (set in the chunk) and the notes below it
kbl_note <- function(df, note, col.names = NA, digits = getOption("digits"), ...) {
  kable(df, col.names = col.names, digits = digits, ...) |>
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) |>
    footnote(general = note, general_title = "Note:", footnote_as_chunk = TRUE,
             threeparttable = TRUE)
}
fmt1 <- function(x) formatC(x, format = "f", digits = 1)
fmt2 <- function(x) formatC(x, format = "f", digits = 2)
fmt3 <- function(x) formatC(x, format = "f", digits = 3)
fmt4 <- function(x) formatC(x, format = "f", digits = 4)
fmtN <- function(x) format(round(x), big.mark = ",")
fmtP <- function(p) ifelse(p < 0.001, "< 0.001", formatC(p, format = "f", digits = 3))
```

```{r}
#| label: analysis
#| include: false

# ==========================================================================
# DATA — extended corpus from the merged Determ database (01_extract_corpus.R)
# All estimates are computed here once; later chunks only display them, and
# every number quoted in the text comes from the `kn` list at the end.
# ==========================================================================

if (!file.exists(path_raw_corpus_extended)) {
  stop("Corpus not found at ", path_raw_corpus_extended, ". Run R/01_extract_corpus.R first.")
}

CHATGPT_DATE <- as.Date(CONFIG$paper1_events$treatment)
PLACEBO_DATE <- as.Date(CONFIG$paper1_events$placebo)
events_df <- bind_rows(
  tibble(date = CHATGPT_DATE, label = "ChatGPT"),
  tibble(date  = as.Date(sapply(CONFIG$paper1_events$secondary, `[[`, "date")),
         label = sapply(CONFIG$paper1_events$secondary, `[[`, "label"))
)

corpus_data <- readRDS(path_raw_corpus_extended) |>
  mutate(
    DATE       = as.Date(DATE),
    year       = year(DATE),
    year_month = floor_date(DATE, "month"),
    word_count = stri_count_regex(FULL_TEXT, "\\S+")
  ) |>
  arrange(DATE)

frame_dictionaries <- lapply(CONFIG$frames, `[[`, "keywords")
all_frames <- names(frame_dictionaries)
frame_cols <- paste0("frame_", all_frames)
for (f in all_frames) {
  corpus_data[[paste0("frame_", f)]] <- stri_detect_regex(
    corpus_data$.text_lower, paste(frame_dictionaries[[f]], collapse = "|"))
}

corpus_data <- bind_cols(corpus_data, classify_outlets(corpus_data$FROM, corpus_data$SOURCE_TYPE)) |>
  mutate(
    outlet_type  = case_when(!is.na(outlet_type) ~ outlet_type,
                             SOURCE_TYPE == "web" ~ "Other web",
                             TRUE ~ "Non-web"),
    outlet_type  = factor(outlet_type, levels = c("Tabloid", "Quality", "Regional", "Public",
                                                  "Tech", "Business", "Other web", "Non-web")),
    threat       = frame_JOB_LOSS | frame_FEAR_RESISTANCE | frame_INEQUALITY,
    opportunity  = frame_JOB_CREATION | frame_PRODUCTIVITY | frame_TRANSFORMATION,
    post_chatgpt = as.integer(DATE >= CHATGPT_DATE),
    month_idx    = (year(year_month) - year(min(year_month))) * 12 +
                   month(year_month) - month(min(year_month)),
    rel_month    = interval(CHATGPT_DATE, year_month) %/% months(1)
  )
T_STAR <- unique(corpus_data$month_idx[corpus_data$year_month == CHATGPT_DATE])
hac <- function(m) coeftest(m, vcov = vcovHAC(m))

# --- Monthly volume ----------------------------------------------------------
monthly <- corpus_data |>
  count(year_month, month_idx) |>
  mutate(post = as.integer(year_month >= CHATGPT_DATE), t = month_idx,
         moy = factor(month(year_month)))

its_model  <- lm(n ~ t + post + I((t - T_STAR) * post), data = monthly)
its_robust <- hac(its_model)
monthly$fitted <- predict(its_model)

vol_m1 <- glm(n ~ post,           data = monthly, family = quasipoisson())
vol_m2 <- glm(n ~ post + t,       data = monthly, family = quasipoisson())
vol_m3 <- glm(n ~ post + t + moy, data = monthly, family = quasipoisson())
outlet_month_panel <- corpus_data |>
  count(FROM, year_month) |>
  mutate(post = as.integer(year_month >= CHATGPT_DATE), moy = factor(month(year_month)))
vol_m4 <- fixest::feglm(n ~ post | FROM + moy, data = outlet_month_panel,
                        family = poisson, cluster = "FROM")
vol_nb <- MASS::glm.nb(n ~ post + t, data = monthly)

# --- Structural breaks -------------------------------------------------------
fs_test   <- Fstats(n ~ t, data = monthly, from = 0.15)
bp_all    <- breakpoints(n ~ t, data = monthly, h = 0.15)
bp_ci     <- confint(bp_all)$confint
fs_log    <- Fstats(log(n) ~ t, data = monthly, from = 0.15)
cusum_test <- efp(n ~ t, data = monthly, type = "OLS-CUSUM")
ym_lab <- function(i) format(monthly$year_month[i], "%b %Y")

# --- Frame panel ---------------------------------------------------------------
frame_monthly <- corpus_data |>
  group_by(year_month, month_idx, rel_month) |>
  summarise(n_total = n(), across(all_of(frame_cols), ~ sum(.x)), .groups = "drop") |>
  pivot_longer(all_of(frame_cols), names_to = "frame", values_to = "count") |>
  mutate(frame = str_remove(frame, "frame_"), pct = 100 * count / n_total,
         post = as.integer(year_month >= CHATGPT_DATE), t = month_idx)

# Event study: three-month event-time bins, reference = Sep-Nov 2022
es_bin <- function(k) floor(k / 3)
es_results <- bind_rows(lapply(all_frames, function(f) {
  d  <- filter(frame_monthly, frame == f) |>
    mutate(q = relevel(factor(es_bin(rel_month)), ref = "-1"))
  ct <- hac(lm(pct ~ q, data = d))
  tibble(frame = f, term = rownames(ct), estimate = ct[, 1], se = ct[, 2], p = ct[, 4]) |>
    filter(str_detect(term, "^q")) |>
    mutate(bin = as.integer(str_remove(term, "q")),
           bin_start = CHATGPT_DATE %m+% months(3 * bin),
           conf.low = estimate - 1.96 * se, conf.high = estimate + 1.96 * se)
}))

# Pre-trend tests on the pre-launch months
pre_bins <- function(k) cut(k, breaks = c(-24, seq(-20, -2, by = 3)), right = TRUE)
pretrend_tbl <- bind_rows(lapply(all_frames, function(f) {
  d <- filter(frame_monthly, frame == f, rel_month <= -2)
  d$bin <- relevel(factor(pre_bins(d$rel_month)), ref = "(-5,-2]")
  a <- anova(lm(pct ~ t, d), lm(pct ~ bin + t, d))
  s <- summary(lm(pct ~ t, d))$coefficients
  tibble(Frame = f, F_stat = a$F[2], F_p = a$`Pr(>F)`[2], slope = s["t", 1], slope_p = s["t", 4])
}))

# Average post-launch effects (post + linear trend, HAC)
ate_tbl <- bind_rows(lapply(all_frames, function(f) {
  d <- filter(frame_monthly, frame == f)
  r <- hac(lm(pct ~ post + t, data = d))
  tibble(Frame = f, pre = mean(d$pct[d$post == 0]), post_mean = mean(d$pct[d$post == 1]),
         ate = r["post", 1], se = r["post", 2], p = r["post", 4])
})) |> arrange(desc(abs(ate)))

# Composite indices
composite_monthly <- corpus_data |>
  group_by(year_month, month_idx) |>
  summarise(n = n(), threat = 100 * mean(threat), opportunity = 100 * mean(opportunity),
            .groups = "drop") |>
  mutate(gap = threat - opportunity, post = as.integer(year_month >= CHATGPT_DATE),
         t = month_idx)
comp_fit <- lapply(c(threat = "threat", opportunity = "opportunity", gap = "gap"),
                   function(y) hac(lm(reformulate(c("post", "t"), y), data = composite_monthly)))

# Persistence: post-launch slopes
decay_tbl <- bind_rows(lapply(c(all_frames, "threat", "opportunity", "gap"), function(f) {
  d <- if (f %in% all_frames) filter(frame_monthly, frame == f, post == 1) |> mutate(y = pct)
       else filter(composite_monthly, post == 1) |> mutate(y = .data[[f]])
  d$ms <- d$t - T_STAR
  r <- hac(lm(y ~ ms, data = d))
  tibble(Series = f, level0 = r[1, 1], slope = r[2, 1], se = r[2, 2], p = r[2, 4])
}))

# --- Platforms --------------------------------------------------------------------
all_months <- distinct(monthly, year_month, month_idx)
platform_its <- bind_rows(lapply(sort(unique(corpus_data$SOURCE_TYPE)), function(p) {
  d <- corpus_data |> filter(SOURCE_TYPE == p) |> count(year_month, month_idx) |>
    right_join(all_months, by = c("year_month", "month_idx")) |>
    mutate(n = coalesce(n, 0L), post = as.integer(year_month >= CHATGPT_DATE), t = month_idx)
  pre_items <- sum(d$n[d$post == 0])
  out <- tibble(Platform = p, total = sum(d$n), pre_items = pre_items,
                pre_mo = mean(d$n[d$post == 0]), post_mo = mean(d$n[d$post == 1]),
                level = NA_real_, level_p = NA_real_, slope = NA_real_, slope_p = NA_real_)
  if (pre_items >= 30) {
    r <- hac(lm(n ~ t + post + I((t - T_STAR) * post), data = d))
    out$level <- r["post", 1]; out$level_p <- r["post", 4]
    out$slope <- r[4, 1];      out$slope_p <- r[4, 4]
  }
  out
})) |> arrange(desc(total))

# --- Tabloid vs quality DiD (web items, named outlets) -------------------------------
did_data <- corpus_data |>
  filter(outlet_type %in% c("Tabloid", "Quality")) |>
  mutate(threat_i = as.integer(threat), tabloid = as.integer(outlet_type == "Tabloid"),
         post_tab = post_chatgpt * tabloid)
did_hetero <- feols(threat_i ~ post_tab + tabloid | year_month, did_data, vcov = "hetero")
did_2way   <- feols(threat_i ~ post_tab + tabloid | year_month, did_data,
                    cluster = c("outlet", "year_month"))
did_dm     <- demean(cbind(y = did_data$threat_i, post_tab = did_data$post_tab,
                           tabloid = did_data$tabloid), did_data$year_month)
did_wild   <- wild_cluster_p(did_dm[, "y"], did_dm[, c("post_tab", "tabloid")],
                             did_data$outlet, "post_tab", B = 9999, weights = "webb")
did_n1 <- feols(threat_i ~ post_tab + tabloid | year_month,
                did_data |> mutate(tabloid = as.integer(tabloid == 1 | outlet == "N1 Hrvatska"),
                                   post_tab = post_chatgpt * tabloid),
                cluster = c("outlet", "year_month"))
did_means <- did_data |> group_by(outlet_type, post_chatgpt) |>
  summarise(threat = 100 * mean(threat_i), n = n(), .groups = "drop")
dm_ <- function(ty, p) did_means$threat[did_means$outlet_type == ty & did_means$post_chatgpt == p]

# Threat and opportunity by outlet type
type_tbl <- bind_rows(lapply(levels(corpus_data$outlet_type), function(ty) {
  d <- corpus_data |> filter(outlet_type == ty) |>
    group_by(year_month, month_idx) |>
    summarise(n = n(), threat = 100 * mean(threat), opp = 100 * mean(opportunity), .groups = "drop") |>
    mutate(post = as.integer(year_month >= CHATGPT_DATE), t = month_idx)
  x_ <- corpus_data |> filter(outlet_type == ty)
  rt <- hac(lm(threat ~ post + t, d)); ro <- hac(lm(opp ~ post + t, d))
  tibble(Type = ty, items = nrow(x_),
         thr_pre = 100 * mean(x_$threat[x_$post_chatgpt == 0]),
         thr_post = 100 * mean(x_$threat[x_$post_chatgpt == 1]),
         thr_ate = rt["post", 1], thr_p = rt["post", 4],
         opp_pre = 100 * mean(x_$opportunity[x_$post_chatgpt == 0]),
         opp_post = 100 * mean(x_$opportunity[x_$post_chatgpt == 1]),
         opp_ate = ro["post", 1], opp_p = ro["post", 4])
}))

# --- Mechanisms ------------------------------------------------------------------------
# (a) Near-duplicates across outlets: title-token Jaccard >= 0.7 within +-3 days.
#     Candidate pairs are blocked on each title's two rarest tokens.
.tok <- corpus_data |>
  mutate(id = row_number()) |>
  filter(!is.na(TITLE)) |>
  transmute(id, DATE, FROM, tok = stri_extract_all_regex(stri_trans_tolower(TITLE), "\\p{L}{4,}")) |>
  tidyr::unnest(tok) |>
  distinct(id, tok, .keep_all = TRUE)
.ntok <- count(.tok, id, name = "n_tok") |> filter(n_tok >= 3)
.tok  <- semi_join(.tok, .ntok, by = "id")
.keys <- .tok |> add_count(tok, name = "df") |> arrange(id, df, tok) |>
  group_by(id) |> slice_head(n = 2) |> ungroup() |>
  mutate(lo = DATE - 3, hi = DATE + 3)
.cand <- inner_join(select(.keys, id, tok, FROM, lo, hi),
                    select(.keys, id2 = id, tok, FROM2 = FROM, DATE2 = DATE),
                    by = join_by(tok, between(y$DATE2, x$lo, x$hi))) |>
  filter(id < id2, FROM != FROM2) |>
  distinct(id, id2)
.inter <- .cand |>
  inner_join(select(.tok, id, tok), by = "id", relationship = "many-to-many") |>
  semi_join(select(.tok, id2 = id, tok), by = c("id2", "tok")) |>
  count(id, id2, name = "inter") |>
  left_join(.ntok, by = "id") |>
  left_join(rename(.ntok, id2 = id, n_tok2 = n_tok), by = "id2") |>
  mutate(jac = inter / (n_tok + n_tok2 - inter)) |>
  filter(jac >= 0.7)
.dup_ids <- unique(c(.inter$id, .inter$id2))
wire_tbl <- corpus_data |>
  mutate(id = row_number(), near_dup = id %in% .dup_ids) |>
  filter(id %in% .ntok$id) |>
  mutate(Period = ifelse(post_chatgpt == 1, "Post-launch", "Pre-launch")) |>
  group_by(Period) |>
  summarise(items = n(), dups = sum(near_dup), share = 100 * mean(near_dup),
            threat_dup = 100 * mean(threat[near_dup]), threat_nodup = 100 * mean(threat[!near_dup]),
            .groups = "drop") |>
  arrange(desc(Period))

# (b) Reader engagement: interactions on web items
eng_data <- corpus_data |> filter(SOURCE_TYPE == "web", !is.na(INTERACTIONS))
eng_all  <- feols(log1p(INTERACTIONS) ~ threat + opportunity | FROM + year_month,
                  eng_data, cluster = "FROM")
eng_pos  <- feols(log(INTERACTIONS) ~ threat + opportunity | FROM + year_month,
                  filter(eng_data, INTERACTIONS > 0), cluster = "FROM")

# --- Robustness: threat composite, post + trend, HAC ------------------------------------
threat_ate <- function(df, var = "threat") {
  m <- df |> group_by(year_month, month_idx) |>
    summarise(y = 100 * mean(.data[[var]]), .groups = "drop") |>
    mutate(post = as.integer(year_month >= CHATGPT_DATE), t = month_idx)
  r <- hac(lm(y ~ post + t, data = m))
  c(est = r["post", 1], se = r["post", 2], p = r["post", 4])
}
windows <- c(6, 9, 12, 18, 24, 36)
window_tbl <- bind_rows(lapply(windows, function(w) {
  r <- threat_ate(filter(corpus_data, DATE >= CHATGPT_DATE %m-% months(w),
                         DATE <  CHATGPT_DATE %m+% months(w)))
  tibble(Window = paste0("±", w, " months"), est = r[1], se = r[2], p = r[3])
}))
placebo_data  <- monthly |> filter(year_month < CHATGPT_DATE) |>
  mutate(placebo_post = as.integer(year_month >= PLACEBO_DATE))
placebo_fit   <- hac(lm(n ~ placebo_post + t, data = placebo_data))
tabloid_names <- CONFIG$paper1_outlets$outlets |> Filter(f = function(o) o$type == "Tabloid") |>
  sapply(`[[`, "name")
alt_fear <- corpus_data |> mutate(threat_nofear = frame_JOB_LOSS | frame_INEQUALITY)

rb <- list(
  c("Full sample (reference)", threat_ate(corpus_data)),
  c("Web items only",          threat_ate(filter(corpus_data, SOURCE_TYPE == "web"))),
  c("Excluding January 2023",  threat_ate(filter(corpus_data, year_month != as.Date("2023-01-01")))),
  c("Threat without fear/resistance", threat_ate(alt_fear, "threat_nofear"))
)
for (o in tabloid_names)
  rb[[length(rb) + 1]] <- c(paste("Excluding", o), threat_ate(filter(corpus_data, is.na(outlet) | outlet != o)))
robust_tbl <- bind_rows(lapply(rb, function(r) tibble(Specification = r[1], est = as.numeric(r[2]),
                                                      se = as.numeric(r[3]), p = as.numeric(r[4]))))

# LPM with outlet FE, trend and month-of-year FE (year-month FE would absorb `post`)
lpm_fit <- feols(as.integer(threat) ~ post_chatgpt + month_idx | FROM + moy,
                 corpus_data |> mutate(moy = month(year_month)), cluster = "FROM")

# --- Anticipation -------------------------------------------------------------------------
.pre <- corpus_data$DATE < as.Date("2022-11-30")
pre_chatgpt <- sum(.pre & stri_detect_regex(corpus_data$.text_lower, "chat.?gpt"))
pre_gpt     <- sum(.pre & stri_detect_regex(corpus_data$.text_lower, "openai|open ai|\\bgpt\\b|gpt.?[34]"))

# --- Cross-platform duplicates -----------------------------------------------------------
.norm <- function(x) stri_trim(stri_replace_all_regex(stri_trans_tolower(x), "[^\\p{L}\\p{N} ]+", " "))
.fbw <- inner_join(
  corpus_data |> filter(SOURCE_TYPE == "facebook") |> transmute(k = .norm(TITLE), DF = DATE, id_f = row_number()),
  corpus_data |> filter(SOURCE_TYPE == "web") |> transmute(k = .norm(TITLE), DW = DATE),
  by = "k", relationship = "many-to-many") |>
  filter(abs(as.numeric(DF - DW)) <= 2, nchar(k) > 25)

# --- Key numbers for the text ---------------------------------------------------------------
es_ <- function(f, bins) mean(es_results$estimate[es_results$frame == f & es_results$bin %in% bins])
late_bins  <- 8:max(es_results$bin)          # December 2024 onward
early_bins <- 0:1                            # first six months
cy <- composite_monthly |> mutate(y = year(year_month)) |> group_by(y) |>
  summarise(threat = mean(threat), opportunity = mean(opportunity), .groups = "drop")
w24 <- window_tbl[window_tbl$Window == "±24 months", ]
w18 <- window_tbl[window_tbl$Window == "±18 months", ]
ate_ <- function(f, col) ate_tbl[[col]][ate_tbl$Frame == f]
forum_tab <- table(corpus_data$FROM[corpus_data$SOURCE_TYPE == "forum"])
bp_dates  <- ym_lab(bp_all$breakpoints)

kn <- list(
  N              = fmtN(nrow(corpus_data)),
  start          = format(min(corpus_data$DATE), "%B %Y"),
  end            = format(max(corpus_data$DATE), "%B %Y"),
  n_pre_months   = sum(monthly$post == 0),
  n_post_months  = sum(monthly$post == 1),
  pre_mean       = fmtN(mean(monthly$n[monthly$post == 0])),
  post_mean      = fmtN(mean(monthly$n[monthly$post == 1])),
  ratio          = fmt1(mean(monthly$n[monthly$post == 1]) / mean(monthly$n[monthly$post == 0])),
  last_year      = max(year(monthly$year_month)),
  last_year_mean = fmtN(mean(monthly$n[year(monthly$year_month) == max(year(monthly$year_month))])),
  first_year_mean = fmtN(mean(monthly$n[year(monthly$year_month) == 2021])),
  its_level      = fmt1(its_robust["post", 1]),
  its_level_p    = fmt2(its_robust["post", 4]),
  its_slope      = fmt1(its_robust[4, 1]),
  its_slope_se   = fmt1(its_robust[4, 2]),
  its_pretrend   = fmt1(its_robust["t", 1]),
  irr_raw        = fmt2(exp(coef(vol_m1)["post"])),
  irr_trend      = fmt2(exp(coef(vol_m2)["post"])),
  irr_trend_p    = fmt2(summary(vol_m2)$coefficients["post", 4]),
  irr_outlet     = fmt2(exp(coef(vol_m4)["post"])),
  supf           = fmt1(max(fs_test$Fstats)),
  supf_date      = ym_lab(fs_test$breakpoint),
  supf_log_date  = ym_lab(fs_log$breakpoint),
  bp1            = bp_dates[1],
  bp1_lo         = ym_lab(bp_ci[1, 1]), bp1_hi = ym_lab(bp_ci[1, 3]),
  bp2            = if (length(bp_dates) > 1) bp_dates[2] else NA,
  bp2_lo         = if (nrow(bp_ci) > 1) ym_lab(bp_ci[2, 1]) else NA,
  bp2_hi         = if (nrow(bp_ci) > 1) ym_lab(bp_ci[2, 3]) else NA,
  placebo        = fmt1(placebo_fit["placebo_post", 1]),
  placebo_p      = fmt2(placebo_fit["placebo_post", 4]),
  threat_pre     = fmt1(mean(composite_monthly$threat[composite_monthly$post == 0])),
  threat_post    = fmt1(mean(composite_monthly$threat[composite_monthly$post == 1])),
  opp_pre        = fmt1(mean(composite_monthly$opportunity[composite_monthly$post == 0])),
  opp_post       = fmt1(mean(composite_monthly$opportunity[composite_monthly$post == 1])),
  gap_pre        = fmt1(mean(composite_monthly$gap[composite_monthly$post == 0])),
  gap_post       = fmt1(mean(composite_monthly$gap[composite_monthly$post == 1])),
  threat_last    = fmt1(cy$threat[cy$y == max(cy$y)]),
  threat_ate     = fmt1(comp_fit$threat["post", 1]),  threat_ate_p = fmt2(comp_fit$threat["post", 4]),
  opp_ate        = fmt1(comp_fit$opportunity["post", 1]), opp_ate_p = fmt2(comp_fit$opportunity["post", 4]),
  gap_ate        = fmt1(comp_fit$gap["post", 1]),     gap_ate_p = fmt2(comp_fit$gap["post", 4]),
  w24            = fmt1(w24$est), w24_p = fmt2(w24$p),
  w18            = fmt1(w18$est), w18_p = fmt2(w18$p),
  fear_early     = fmt1(es_("FEAR_RESISTANCE", early_bins)),
  fear_late      = fmt1(es_("FEAR_RESISTANCE", late_bins)),
  skills_ate     = fmt1(ate_("SKILLS", "ate")), skills_p = fmtP(ate_("SKILLS", "p")),
  skills_late    = fmt1(es_("SKILLS", late_bins)),
  reg_peak       = fmt1(max(es_results$estimate[es_results$frame == "REGULATION"])),
  reg_peak_date  = format(es_results$bin_start[es_results$frame == "REGULATION"][
                     which.max(es_results$estimate[es_results$frame == "REGULATION"])], "%B %Y"),
  prod_early     = fmt1(es_("PRODUCTIVITY", early_bins)),
  prod_late      = fmt1(es_("PRODUCTIVITY", late_bins)),
  did_coef       = fmt1(100 * coef(did_2way)["post_tab"]),
  did_p          = fmt3(pvalue(did_2way)["post_tab"]),
  did_p_wild     = fmt3(did_wild$p),
  did_G          = did_wild$G,
  did_n1         = fmt1(100 * coef(did_n1)["post_tab"]),
  tab_pre = fmt1(dm_("Tabloid", 0)), tab_post = fmt1(dm_("Tabloid", 1)),
  qual_pre = fmt1(dm_("Quality", 0)), qual_post = fmt1(dm_("Quality", 1)),
  eng_threat     = fmt1(100 * (exp(coef(eng_all)["threatTRUE"]) - 1)),
  eng_opp        = fmt1(100 * (exp(coef(eng_all)["opportunityTRUE"]) - 1)),
  eng_threat_p   = fmtP(pvalue(eng_all)["threatTRUE"]),
  eng_opp_p      = fmtP(pvalue(eng_all)["opportunityTRUE"]),
  wire_pre       = fmt1(wire_tbl$share[wire_tbl$Period == "Pre-launch"]),
  wire_post      = fmt1(wire_tbl$share[wire_tbl$Period == "Post-launch"]),
  web_share      = fmt1(100 * mean(corpus_data$SOURCE_TYPE == "web")),
  named_share    = fmt1(100 * mean(!is.na(corpus_data$outlet))),
  forumhr_share  = fmt1(100 * forum_tab["forum.hr"] / sum(forum_tab)),
  fb_n           = fmtN(sum(corpus_data$SOURCE_TYPE == "facebook")),
  fb_dup         = fmtN(n_distinct(.fbw$id_f)),
  fb_dup_share   = fmt1(100 * n_distinct(.fbw$id_f) / sum(corpus_data$SOURCE_TYPE == "facebook")),
  pre_chatgpt_mentions = fmtN(pre_chatgpt),
  pre_gpt_mentions     = fmtN(pre_gpt),
  pre_gpt_share        = fmt2(100 * pre_gpt / sum(.pre)),
  nofear         = fmt1(robust_tbl$est[robust_tbl$Specification == "Threat without fear/resistance"]),
  nofear_p       = fmt2(robust_tbl$p[robust_tbl$Specification == "Threat without fear/resistance"])
)

# --- Guards: the prose below states these directions; stop if the data disagree ----------
stopifnot(
  its_robust[4, 1] > 0, its_robust[4, 4] < 0.05,         # post-launch slope increase
  its_robust["post", 4] > 0.10,                          # no significant level jump
  coef(did_2way)["post_tab"] < 0,                        # quality > tabloid response
  es_("FEAR_RESISTANCE", late_bins) > es_("FEAR_RESISTANCE", early_bins),
  ate_("SKILLS", "ate") < 0, ate_("SKILLS", "p") < 0.05,
  coef(eng_all)["threatTRUE"] > 0, coef(eng_all)["opportunityTRUE"] < 0,
  placebo_fit["placebo_post", 4] > 0.10,
  mean(composite_monthly$gap[composite_monthly$post == 1]) >
    mean(composite_monthly$gap[composite_monthly$post == 0])
)
```

# Abstract

ChatGPT's November 30, 2022 release made generative artificial intelligence
visible to a mass audience overnight. We treat the launch as an exogenous
shock and trace its effect on Croatian digital media coverage of AI and the
labour market, using `r kn$N` Croatian-language items published between
`r kn$start` and `r kn$end` and drawn from the Determ media-monitoring
database. Coverage did not jump at the launch; it accelerated. Monthly volume
shows no significant level shift in December 2022 but a sustained increase in
its trend (`r kn$its_slope` additional items per month, each month), rising
from `r kn$pre_mean` items a month before the launch to `r kn$post_mean`
after it, and a December 2021 placebo is null. The share of items carrying a
threat frame (job loss, fear, inequality) rose from `r kn$threat_pre`% to
`r kn$threat_post`%, while the opportunity share (job creation,
productivity, transformation) stayed flat; fear-and-resistance framing
continued to grow for more than three years, and skills and reskilling
framing fell and did not recover. Contrary to our prediction, quality
outlets increased threat framing more than tabloids. Threat-framed articles
attract more reader interactions than other articles, consistent with a
demand-side explanation of the shift.

# Introduction

Croatian-language digital coverage of AI and work grew from about
`r kn$pre_mean` items a month before ChatGPT's release to `r kn$post_mean`
a month afterwards, and by `r kn$last_year` it averaged `r kn$last_year_mean`
items a month. Its tone changed with its volume: the share of items
framing AI as a threat to jobs, as a source of fear, or as a driver of
inequality rose from `r kn$threat_pre`% before the launch to
`r kn$threat_post`% after it. Whether generative AI ultimately displaces,
complements, or transforms human work remains an open empirical question
[@acemoglu2018race; @autor2015why; @bessen2020automation]. Workers, firms,
and policymakers must nonetheless act under uncertainty, and the public
information environment that shapes those actions has moved measurably.

Task-level evidence suggests that large language models are exposed to a
wide and disproportionately white-collar share of occupations
[@eloundou2023gpts; @felten2021ai], and field experiments find that
generative AI raises productivity for some workers [@brynjolfsson2023generative].
In a small economy such as Croatia, where firm-level data on AI adoption are
sparse, the main channel through which workers and policymakers learn about
these capabilities is mediated coverage. Which frames that channel
transmits, how quickly, and how persistently is therefore an input to
belief formation and labour-market adjustment.

This paper estimates how the November 30, 2022 ChatGPT release changed
Croatian digital media coverage of AI and the labour market. We treat the
release as an exogenous shock to the salience of generative-AI capabilities
and follow coverage for `r kn$n_pre_months` months before and
`r kn$n_post_months` months after it. The corpus of `r kn$N` items is drawn
from the Determ media-monitoring database through a two-stage filter that
requires both an AI term and a labour-market term. We combine an
interrupted time series for volume, structural-break tests that locate
breaks without imposing the launch date, frame-level event studies, and a
tabloid–quality difference-in-differences.

Four findings stand out. First, the launch changed the *trajectory* of
coverage rather than its level. Monthly volume shows no significant jump in
December 2022 (`r kn$its_level` items, p = `r kn$its_level_p`) but its trend
increases by `r kn$its_slope` items per month, each month, after the launch.
The Bai–Perron procedure dates the first break to `r kn$bp1` (95% interval
`r kn$bp1_lo`–`r kn$bp1_hi`), within three months of the launch, and a
second to `r kn$bp2`; a placebo at December 2021 is null.

Second, the tone of coverage shifted toward threat. The threat share rose
from `r kn$threat_pre`% to `r kn$threat_post`% while the opportunity share
barely moved (`r kn$opp_pre`% to `r kn$opp_post`%). The shift was gradual
rather than abrupt: within 24 months of the launch the threat share is
`r kn$w24` percentage points higher than its trend would predict
(p = `r kn$w24_p`), but over the full sample a linear trend absorbs most of
the change. Fear-and-resistance framing drives the shift, rising by
`r kn$fear_early` points in the first six months and by `r kn$fear_late`
points from December 2024 onward, while skills and reskilling framing fell
by `r fmt1(abs(ate_("SKILLS", "ate")))` points and stayed down.

Third, the prediction that tabloids would amplify threat framing most is
rejected. Quality outlets, which started from a lower threat share,
increased it by more than tabloids did; the difference-in-differences
estimate is `r kn$did_coef` percentage points for tabloids relative to
quality outlets (p = `r kn$did_p` with two-way clustering;
`r kn$did_p_wild` with a wild-cluster bootstrap).

Fourth, readers reward threat framing. Among web articles from the same
outlet in the same month, threat-framed articles receive
`r kn$eng_threat`% more interactions and opportunity-framed articles
`r abs(as.numeric(kn$eng_opp))`% fewer, which is consistent with editors
responding to demand.

The paper contributes to three literatures. For work on automation
expectations [@acemoglu2018race; @autor2022work], it documents the supply
of public information that shapes those expectations in a small European
economy where direct measurement of AI adoption is not available. For work
on media framing of AI [@fasthorvitz2017; @chuan2019framing;
@nguyenhekman2024], most of it descriptive and on English-language media, it
provides event-study evidence around a dated shock, a post-period of almost
four years, and direct evidence on outlet type and reader demand. For work
on the labour-market consequences of generative AI [@eloundou2023gpts;
@felten2021ai; @brynjolfsson2023generative], it supplies a measurable input,
the tone of coverage, that can be linked to vacancy, employment, and
reskilling outcomes.

Section 3 describes the Croatian media market and the conceptual framework.
Section 4 describes the corpus, the frame dictionaries, and the outlet
classification. Section 5 sets out the empirical strategy. Section 6
reports the results, Section 7 the robustness checks, and Sections 8 and 9
discuss and conclude.

# Institutional and Conceptual Background

## The Croatian Media Landscape

Croatia's online news market is served by a small number of national
portals, most of which belong to a handful of domestic and foreign owners.
Styria Media Group, an Austrian company, owns the tabloid *24sata*, which
@perusko2024croatia describes as the best-selling newspaper and one that
attracts large numbers of readers online, as well as *Večernji list* and
*Poslovni dnevnik*. The domestically owned Hanza Media, the print market
leader, publishes *Jutarnji list* and *Slobodna Dalmacija*. United Group
operates Nova TV, whose news portal is *Dnevnik.hr*, and the cable news
channel N1, while the Czech-based CME has owned RTL Hrvatska since 2022
[@perusko2024croatia]. RTL Hrvatska took over the portal *Net.hr* from
Telegram Media Grupa in 2021 [@nethr2021rtl], which continues to publish
*Telegram*. *Index.hr* is independently owned, and *Tportal* is published by
Hrvatski Telekom.

The outlets we analyse by name are listed, with their owners and their
number of items in our corpus, in @tbl-outlets (Section 4.3). They are the
national portals of these groups plus the public broadcaster, two regional
dailies, and specialist technology and business titles. Items from all
other sources enter the aggregate analysis but not the outlet-level
comparisons.

The outlet typology used throughout the paper (Tabloid, Quality, Regional,
Public, Tech, Business) describes editorial model and audience rather than
ownership: an owner group can publish both a tabloid and a quality title, as
Styria does with *24sata* and *Večernji list*. Tabloid outlets (24sata,
Index.hr, Net.hr) prioritise high-volume, attention-driven coverage; quality
outlets (Jutarnji list, Večernji list, Dnevnik.hr, N1, Telegram, Tportal)
run longer analytical pieces. The Tabloid–Quality contrast is the basis for
the difference-in-differences in Section 6.8.

## ChatGPT as a Salience Shock

We treat ChatGPT's November 30, 2022 public release as an unanticipated
shock to the salience of generative-AI capabilities for Croatian editors and
readers. The release was announced and shipped on the same day. Section 5.3
reports the in-corpus evidence: only `r kn$pre_chatgpt_mentions` items in
our corpus mention ChatGPT before the release, and the pre-launch items that
mention OpenAI or GPT concern earlier models.

The treatment date is December 1, 2022, the first month in which every
day is post-release in the monthly aggregation our design uses.

## Conceptual Framework

A short verbal model of editorial choice under attention scarcity yields
three predictions. Editors fill a limited number of article slots to
maximise expected readership, and readers respond more strongly to a vivid
downside than to a measured upside [@soroka2015negativity;
@robertson2023negativity]. A salience shock expands the menu of story angles
and raises the expected return to AI-and-work coverage relative to other
topics competing for the same slots.

Three predictions follow. First, coverage volume rises after the shock,
because the expected return to the topic now exceeds the opportunity cost
of a slot (tested in Sections 6.1–6.2). Second, threat-leaning angles
(displacement, fear, inequality) gain more than opportunity-leaning angles
(job creation, productivity, transformation), because reader negativity
bias makes threat angles more profitable per slot (Sections 6.3–6.6).
Third, outlets whose business model depends most on attention-driven
revenue, tabloids, amplify the threat tilt more than quality outlets that
depend more on subscriptions and reputation [@esser1999tabloidization]
(Section 6.8). The demand-side mechanism behind the second prediction also
implies that threat-framed articles attract more reader engagement, which
we test in Section 6.9.

### Literature review (draft outline for the co-authors)

`[DRAFT OUTLINE — to be written up by the co-authors. All references below
are verified against CrossRef and are in references.bib; summaries of each
paper's findings must be checked against the paper before the prose is
written.]`

1. **Framing as selection and salience.** @entman1993framing defines
   framing as selecting some aspects of a perceived reality and making them
   more salient; @scheufele1999framing separates media frames (what we
   measure) from audience frames (what we do not);
   @devreese2005framing distinguishes generic from issue-specific frames.
   *Where we fit:* our eight frames are issue-specific frames for AI and
   work, measured as media frames only.
2. **Generic news frames.** @semetkovalkenburg2000 identify conflict,
   human-interest, economic-consequences, responsibility and morality
   frames. *Where we fit:* map our frames onto these (Section 4.2).
3. **Media framing of AI.** Long-run trends in US press coverage
   [@fasthorvitz2017]; benefit and risk framing in US newspapers
   [@chuan2019framing]; ethical issues in AI coverage
   [@ouchchy2020headlines]; how news discourse makes sense of automation
   [@nguyenhekman2024]; ChatGPT headlines in UK media [@roeperkins2023].
   *Where we fit:* this work is mostly descriptive and on English-language
   media; we give event-study evidence around a dated shock for a small
   non-English market, with almost four years of post-shock data.
4. **Negativity and the demand for threat framing.**
   @soroka2015negativity, @soroka2019crossnational and
   @robertson2023negativity on negativity bias in news attention and
   consumption; @mullainathanshleifer2005 and @gentzkowshapiro2010 on
   demand-driven media content. *Where we fit:* motivates prediction 2 and
   the engagement test in Section 6.9.
5. **Tabloidisation and outlet type.** @esser1999tabloidization.
   *Where we fit:* motivates prediction 3, which our data reject; the
   write-up should discuss why quality outlets responded more (Section 8).
6. **Labour-market exposure to generative AI.** @eloundou2023gpts,
   @felten2021ai, @brynjolfsson2023generative: the benchmark against which
   the tone of coverage can be judged.

# Data and Measurement

## Corpus Construction

The corpus draws on the Determ media-monitoring database, which holds
about 42 million records of Croatian-language digital content published
between January 2021 and September 2026. Each record is one published item:
a news-portal article, a Facebook post, a tweet, a YouTube video
description, a Reddit post, a forum post, or an item from another monitored
channel. We use every month from January 2021 to August 2026, the last
complete month, and exclude reader comments on news sites, which are
audience reactions rather than media content.

We construct the analytical corpus through a two-stage intersection filter.
An item enters the corpus only if its title or full text contains at least
one term from an AI-and-technology vocabulary *and* at least one term from a
labour-market vocabulary. The first vocabulary covers terms such as
*umjetna inteligencija* (artificial intelligence), *strojno učenje* (machine
learning), ChatGPT, GPT, OpenAI, *automatizacija*, *robotizacija*,
*algoritam*, *neuronska mreža*, *chatbot*, and related expressions. The
second covers labour-market terms including *posao* (job), *zaposlen*
(employed), *nezaposlenost* (unemployment), *radno mjesto* (workplace),
*tržište rada* (labour market), *vještine* (skills), *karijera* (career),
*produktivnost* (productivity), *otpuštanje* (lay-offs), *prekvalifikacija*
(reskilling), and *plaća* (wage). A fast SQL pre-filter selects candidates in
the database; stem-based regular expressions, which accommodate Croatian
inflection (the pattern `umjetn.*inteligencij` matches all case forms of
*umjetna inteligencija*), then confirm both conditions. After removing
duplicates on title, date, and source, the corpus contains `r kn$N` items.

The corpus records media supply rather than media consumption: we observe
what was published, not whether readers saw it or were influenced by it. We
treat it as a measure of the public information environment on the supply
side throughout.

**Platforms.** @tbl-platforms shows the corpus by platform. Web portals
supply `r kn$web_share`% of all items, so the aggregate results are, in
substance, results about news portals. Social-media items come from outlet
and public pages, channels and accounts. Forum items come mainly from the
Forum.hr discussion board (`r kn$forumhr_share`% of forum items), followed
by the Bug.hr forum and regional boards such as Osijek031. Some platforms enter the database only after the launch; they appear
in the descriptive tables but cannot contribute to a pre/post comparison.

```{r}
#| label: tbl-platforms
#| tbl-cap: "Corpus items by platform"

platform_tbl <- corpus_data |>
  count(SOURCE_TYPE, period = ifelse(post_chatgpt == 1, "post", "pre")) |>
  pivot_wider(names_from = period, values_from = n, values_fill = 0) |>
  mutate(total = pre + post,
         share = 100 * total / sum(total),
         first = sapply(SOURCE_TYPE, function(p)
           format(min(corpus_data$DATE[corpus_data$SOURCE_TYPE == p]), "%b %Y"))) |>
  arrange(desc(total)) |>
  transmute(Platform = SOURCE_TYPE, `Pre-launch` = fmtN(pre),
            `Post-launch` = fmtN(post), Total = fmtN(total),
            `Share (%)` = fmt1(share), `First item` = first)

kbl_note(platform_tbl, paste0(
  "Pre-launch: January 2021 to November 2022; post-launch: December 2022 to ",
  format(max(corpus_data$DATE), "%B %Y"), ". Reader comments are excluded. ",
  "First item is the month of the platform's first item in the corpus."))
```

**Duplicates across platforms.** An outlet can publish a story on its portal
and share it on its Facebook page, so the same story can enter the corpus
twice. We measure this directly: `r kn$fb_dup` of `r kn$fb_n` Facebook posts
(`r kn$fb_dup_share`%) repeat the headline of a web item published within two
days. Duplication is therefore negligible, and all outlet-level analyses use
web items only.

## Frame Detection

We operationalise eight interpretive frames, each a distinct way in which
the media discuss the relationship between AI and work. Curated Croatian
keyword dictionaries detect each frame; an item carries a frame if any of
its dictionary terms appears in the title or text. The threat-leaning frames
are *job loss* (*gubitak posla*, *ukidanje radnih mjesta*, *zamjena
radnika*, *otpuštanje*, *tehnološka nezaposlenost*), *fear and resistance*
(*strah od AI*, *prijetnja*, *opasnost*, *neizvjesnost*, *panika*), and
*inequality* (*nejednakost*, *digitalni jaz*, *polarizacija*, *nestanak
srednje klase*). The opportunity-leaning frames are *job creation* (*nova
radna mjesta*, *potražnja za stručnjacima*, *nova zanimanja*),
*productivity* (*učinkovitost*, *ušteda*, *gospodarski rast*,
*optimizacija*), and *task transformation* (*transformacija rada*,
*prilagodba*, *suradnja čovjeka i*, *evolucija poslova*). The remaining two
are *skills* (*prekvalifikacija*, *cjeloživotno učenje*, *digitalne
vještine*, *reskilling*) and *regulation* (*regulacija AI*, *AI Act*,
*zaštita radnika*, *prava radnika*, *sindikat*). Appendix A lists the full
dictionaries.

`[DRAFT OUTLINE — frames in the literature, for the co-authors.]` The
eight frames are issue-specific frames in the sense of
@devreese2005framing. The write-up should (i) map them onto the generic news
frames of @semetkovalkenburg2000 (economic consequences → productivity and
job loss; responsibility → regulation; human interest → fear and
resistance); (ii) relate them to the frames identified in earlier studies of
AI coverage [@chuan2019framing; @nguyenhekman2024; @roeperkins2023]; and
(iii) justify the threat/opportunity grouping with the negativity
literature [@soroka2015negativity; @robertson2023negativity].

Frames are not mutually exclusive: an item that discusses both displacement
and reskilling carries both. We analyse each frame's monthly prevalence, the
share of items carrying it. The dictionaries are not validated against a
human-coded sample, so per-frame precision and recall are unknown.
Measurement error that is unrelated to the launch date attenuates estimated
effects rather than creating them. It does not, however, protect against
dictionary terms whose everyday use changes over time, and several
fear-and-resistance terms (*prijetnja*, *opasnost*, *upozorenje*) are
common outside the AI context. We return to this in the robustness section.

## Outlet Classification

The outlet classification is our own; it is not an official or
vendor-supplied typology. We assign web items to named outlets by their web
domain, using an explicit list in the project configuration file (a domain
matches its subdomains, so *sportske.jutarnji.hr* belongs to Jutarnji list).
Each outlet is then assigned a type according to its editorial model, and
its owner according to @perusko2024croatia. Only web items are classified,
so an outlet's Facebook page never enters the outlet comparisons. Web items
from all other domains form the *Other web* category. @tbl-outlets lists
the named outlets.

```{r}
#| label: tbl-outlets
#| tbl-cap: "Named outlets in the corpus"

outlet_tbl <- corpus_data |>
  filter(!is.na(outlet)) |>
  group_by(outlet_type, outlet, owner) |>
  summarise(n = n(), pre = sum(post_chatgpt == 0), post = sum(post_chatgpt == 1),
            .groups = "drop") |>
  mutate(outlet_type = factor(outlet_type, levels = c("Tabloid", "Quality", "Regional",
                                                      "Public", "Tech", "Business")),
         share = 100 * n / nrow(corpus_data)) |>
  arrange(outlet_type, desc(n))

kbl_note(outlet_tbl |> transmute(Type = outlet_type, Outlet = outlet, Owner = owner,
                                 `Pre-launch` = fmtN(pre), `Post-launch` = fmtN(post),
                                 Total = fmtN(n), `Share of corpus (%)` = fmt1(share)),
         paste0("Web items only, matched by domain. Owners follow Peruško (2024) ",
                "and, for Net.hr, the 2021 takeover by RTL Hrvatska. Together the ",
                "named outlets account for ", kn$named_share, "% of all corpus items."))
```

## Descriptive Summary

@tbl-summary summarises the corpus, and @tbl-types breaks it down by outlet
type, aggregating the outlet-level statistics in Appendix E. The named
outlets supply `r kn$named_share`% of all items; most items come from other
web portals, among them technology and business sites. In every outlet type
the threat share is higher after the launch than before it, while the
opportunity share rises in some types and falls in others. Section 6.8
tests these differences formally.

```{r}
#| label: tbl-summary
#| tbl-cap: "Corpus descriptive statistics"

summary_tbl <- tibble(
  Metric = c("Items", "Distinct sources", "Period", "Months (pre / post launch)",
             "Mean word count", "Median word count",
             "Items before the launch", "Items from the launch onward"),
  Value = c(kn$N, fmtN(n_distinct(corpus_data$FROM)),
            paste(kn$start, "to", kn$end),
            paste(kn$n_pre_months, "/", kn$n_post_months),
            fmtN(mean(corpus_data$word_count, na.rm = TRUE)),
            fmtN(median(corpus_data$word_count, na.rm = TRUE)),
            fmtN(sum(corpus_data$post_chatgpt == 0)),
            fmtN(sum(corpus_data$post_chatgpt == 1)))
)
kbl_note(summary_tbl, "The launch month is December 2022. Items are deduplicated on title, date and source.")
```

```{r}
#| label: tbl-types
#| tbl-cap: "Items and frame shares by outlet type"

kbl_note(
  type_tbl |> transmute(Type, Items = fmtN(items),
                        `Threat pre (%)` = fmt1(thr_pre), `Threat post (%)` = fmt1(thr_post),
                        `Opportunity pre (%)` = fmt1(opp_pre), `Opportunity post (%)` = fmt1(opp_post)),
  paste("Shares of items carrying at least one threat frame (job loss, fear/resistance,",
        "inequality) or opportunity frame (job creation, productivity, transformation).",
        "Tabloid to Business: the named web outlets. Other web: all other web",
        "domains. Non-web: all other platforms."))
```

# Empirical Strategy

## Event Definition and Identification

The treatment date is December 1, 2022, the first full month after
ChatGPT's public release. The identifying assumption is that Croatian media
did not anticipate the release, so a sharp change in coverage at that date
can be attributed to the shock rather than to a smooth pre-existing trend.

The sample runs for `r kn$n_pre_months` months before the launch and
`r kn$n_post_months` months after it. The long post-period is what allows us
to ask whether the shift persisted, but it also contains later events that
plausibly moved coverage of AI and work: the release of GPT-4 (March 2023),
the European Parliament's adoption of the EU AI Act (March 2024), and the
Act's entry into force (August 2024). We mark these as secondary events in
the figures. They do not enter the estimation as separate treatments, so
the post-launch coefficients measure the level of coverage after December
2022 relative to before, not the effect of the release alone. For a
cleaner estimate of the launch effect we report narrow symmetric windows
around the launch in the robustness section.

## Specification

For each frame $f$ we estimate an event study on the monthly frame share,
with event time grouped into three-month bins:

$$
Y_{ft} = \alpha_f + \sum_{b \neq -1} \beta_b^f \cdot \mathbf{1}[\lfloor (t - t^*)/3 \rfloor = b] + \varepsilon_{ft}
$$

where $Y_{ft}$ is the share of items carrying frame $f$ in month $t$, $t^*$
is December 2022, and bin $b = -1$ (September–November 2022) is the
reference. The coefficient $\beta_b^f$ is the change in frame prevalence in
bin $b$ relative to the three months before the launch. We use bins rather
than single months because a monthly specification with one dummy per month
leaves no degrees of freedom for anything else in a series with one
observation per month.

For summary effects we regress each monthly series on a post-launch
indicator and a linear time trend, with Newey–West (HAC) standard errors.
For aggregate volume we also estimate an interrupted time series (ITS) model
[@bernal2017interrupted]:

$$
N_t = \beta_0 + \beta_1 t + \beta_2 \, \text{Post}_t + \beta_3 \, (t - t^*) \, \text{Post}_t + \varepsilon_t ,
$$

where $N_t$ is the monthly item count. Because the slope-change term is
measured from the launch month, $\beta_2$ is the jump in volume *at* the
launch, and $\beta_3$ is the change in the monthly trend after it.

We complement the ITS with structural-break tests that do not impose the
launch date: a Sup-F test over the central 70% of the sample, and the
Bai–Perron break-dating procedure. A placebo at December 2021 checks for
seasonal artefacts.

## Threats to Identification

**Anticipation.** We count pre-launch items that mention ChatGPT or related
terms. Although Croatian interest in ChatGPT was minimal before the launch,
we find `r kn$pre_gpt_mentions` pre-launch items in the corpus that mention
OpenAI or GPT (`r kn$pre_gpt_share`% of pre-launch items). Before the launch
these terms referred to earlier models such as GPT-3; only
`r kn$pre_chatgpt_mentions` pre-launch items mention ChatGPT itself. Anticipation is therefore not a quantitatively
important threat.

**Confounding events.** Croatia adopted the euro and joined the Schengen
Area on January 1, 2023, one month after the launch. Neither concerns AI,
and the corpus filter requires an AI term, so these events enter the corpus
only where an item also discusses AI and work. We re-estimate excluding
January 2023. The later AI-specific events (GPT-4, the EU AI Act) are part
of the post-launch environment, as discussed above.

**Measurement error in framing.** The dictionaries are not validated
against human coding. Error unrelated to the launch date attenuates
estimates, but dictionary terms whose general use changes over time can
bias them. We re-estimate the threat composite without the
fear-and-resistance dictionary, whose terms are the most generic.

**Selection into the corpus.** The corpus includes every item in the
monitoring database that passes the filter; the named outlets in
@tbl-outlets are the national portals of the main ownership groups, the
public broadcaster, and specialist titles, and together they account for
`r kn$named_share`% of the corpus. The database covers smaller regional
portals and niche forums less deeply, and it adds platforms over time
(@tbl-platforms). Because web portals supply almost all items and the
outlet-level analyses use web items from a fixed list of outlets, changes
in platform coverage cannot drive the outlet comparisons. We also report
the headline estimates on web items only.

## Inference

Monthly time-series regressions use Newey–West (HAC) standard errors. The
article-level difference-in-differences in Section 6.8 clusters on outlet
and year-month. Because the Tabloid and Quality groups together contain
only `r kn$did_G` outlets, we also report a wild-cluster bootstrap p-value
(restricted, Webb weights, 9,999 draws, clustered on outlet), following
@cameronGelbachMiller2008 and @mackinnonWebb2017; where it disagrees with the
asymptotic p-value, we follow the bootstrap. All outlets are treated on the
same date, so the staggered-adoption critiques of two-way fixed effects
[@goodmanbacon2021; @dechaisemartinDhaultfoeuille2020] do not apply.

# Results

## Volume Shock

@fig-volume-shock plots the monthly number of corpus items with the fitted
interrupted-time-series line. From `r kn$start` to November 2022 volume
averages `r kn$pre_mean` items a month and drifts slightly downward. From
December 2022 onward it averages `r kn$post_mean` items a month, a
`r kn$ratio`-fold increase, and it keeps rising: from `r kn$first_year_mean`
items a month in 2021 to `r kn$last_year_mean` in `r kn$last_year`.

```{r}
#| label: fig-volume-shock
#| fig-cap: "Monthly item volume and interrupted-time-series fit"
#| fig-height: 5

ggplot(monthly, aes(x = year_month, y = n)) +
  geom_col(aes(fill = factor(post)), show.legend = FALSE) +
  scale_fill_manual(values = c("0" = "gray80", "1" = "gray50")) +
  geom_line(aes(y = fitted), color = "black", linewidth = 0.9) +
  geom_vline(data = events_df, aes(xintercept = date),
             linetype = "dashed", color = "black", linewidth = 0.4) +
  geom_label(data = events_df, aes(x = date, y = max(monthly$n) * c(0.97, 0.85, 0.97, 0.85),
                                   label = label),
             size = 2.7, fill = "white", color = "black", label.size = 0.2) +
  scale_x_date(date_breaks = "6 months", date_labels = "%b\n%Y") +
  labs(x = NULL, y = "Items per month")
```

::: {.figure-note}
**Note:** Bars show the number of corpus items per month (light grey before
the launch, dark grey from December 2022). The black line is the fit of the
interrupted-time-series model in @tbl-its-volume. Dashed lines mark the
ChatGPT launch and the secondary events: the GPT-4 release (March 2023), the
European Parliament's adoption of the EU AI Act (March 2024), and the Act's
entry into force (August 2024).
:::

@tbl-its-volume reports the ITS estimates. The level shift, the jump in
monthly volume at December 2022, is `r kn$its_level` items and is not
statistically different from zero (p = `r kn$its_level_p`). The launch
instead changed the slope: after December 2022 monthly volume grows by an
additional `r kn$its_slope` items per month (HAC SE `r kn$its_slope_se`),
against a pre-launch trend of `r kn$its_pretrend` items per month. ChatGPT
therefore did not produce a one-off burst of coverage followed by a return
to the old level. It set coverage on a new, steeper path.

```{r}
#| label: tbl-its-volume
#| tbl-cap: "Interrupted time series for monthly item volume"

its_tbl <- tibble(
  Term = c("Intercept (level in January 2021)", "Pre-launch trend (items per month)",
           "Level shift at December 2022", "Change in trend after the launch"),
  Estimate = fmt2(its_robust[, 1]), SE = fmt2(its_robust[, 2]),
  t = fmt2(its_robust[, 3]), p = fmtP(its_robust[, 4]))
kbl_note(its_tbl, paste0(
  "Dependent variable: items per month. Model: N = b0 + b1 t + b2 Post + b3 (t - t*) Post, ",
  "where t counts months from January 2021 and t* is December 2022. Because the trend-change ",
  "term is measured from the launch month, b2 is the jump at the launch. Newey-West (HAC) ",
  "standard errors. N = ", nrow(monthly), " months."))
```

@tbl-volume-eventstudy reports the incidence-rate ratio for the post-launch
period under four count specifications. Without a trend, post-launch volume
is `r kn$irr_raw` times its pre-launch level. Once a linear trend is
included the ratio falls to `r kn$irr_trend` (p = `r kn$irr_trend_p`),
because a single trend fitted to the whole period absorbs the
post-launch acceleration. Within outlets (column 4), volume after the launch
is `r kn$irr_outlet` times its pre-launch level.

```{r}
#| label: tbl-volume-eventstudy
#| tbl-cap: "Post-launch volume under four count specifications"

.row_for <- function(model) {
  ct <- if (inherits(model, "fixest")) fixest::coeftable(model) else summary(model)$coefficients
  c(fmt3(ct["post", 1]), paste0("(", fmt3(ct["post", 2]), ")"), fmt2(exp(ct["post", 1])))
}
.rows <- rbind(.row_for(vol_m1), .row_for(vol_m2), .row_for(vol_m3), .row_for(vol_m4))
vol_es_tbl <- tibble(
  ` ` = c("Post (coefficient)", "Post (SE)", "Post (IRR)", "Linear trend",
          "Month-of-year FE", "Source FE", "Observations"),
  `(1) Post only` = c(.rows[1, ], "No", "No", "No", nobs(vol_m1)),
  `(2) + Trend`   = c(.rows[2, ], "Yes", "No", "No", nobs(vol_m2)),
  `(3) + Month FE` = c(.rows[3, ], "Yes", "Yes", "No", nobs(vol_m3)),
  `(4) Source panel` = c(.rows[4, ], "No", "Yes", "Yes", fmtN(nobs(vol_m4))))
kbl_note(vol_es_tbl, paste(
  "Dependent variable: items per month (columns 1-3) or items per source and month (column 4).",
  "Columns 1-3: quasi-Poisson, standard errors from the quasi-likelihood. Column 4: Poisson with",
  "source and month-of-year fixed effects, standard errors clustered by source. IRR = exp(coefficient)."))
```

## Structural Break Tests

The ITS imposes the launch date. The Sup-F test and the Bai–Perron
procedure do not: they search for breaks in the relationship between
monthly volume and time over the central 70% of the sample. @tbl-structural-break
reports the results, and @fig-cusum the OLS-CUSUM path.

```{r}
#| label: tbl-structural-break
#| tbl-cap: "Structural-break tests for monthly item volume"

sb_tbl <- tibble(
  Statistic = c("Sup-F statistic", "Sup-F p-value", "Sup-F break date (levels)",
                "Sup-F break date (log volume)",
                paste0("Bai-Perron break ", seq_along(bp_all$breakpoints), " (95% CI)")),
  Value = c(kn$supf, fmtP(sctest(fs_test)$p.value), kn$supf_date, kn$supf_log_date,
            paste0(ym_lab(bp_ci[, 2]), " (", ym_lab(bp_ci[, 1]), " – ", ym_lab(bp_ci[, 3]), ")")))
kbl_note(sb_tbl, paste(
  "Model: items per month on a constant and a linear trend. The Sup-F test searches the central",
  "70% of the sample for a single break; the Bai-Perron procedure chooses the number of breaks by BIC",
  "(minimum segment 15% of the sample) and reports 95% confidence intervals for the dates.",
  "The launch month is December 2022."))
```

```{r}
#| label: fig-cusum
#| fig-cap: "OLS-CUSUM test for the monthly volume series"
#| fig-height: 4

par(mar = c(4, 4, 1, 1))
plot(cusum_test, main = "", xlab = "Share of the sample", ylab = "Cumulative sum of residuals", col = "black")
```

::: {.figure-note}
**Note:** Cumulative sum of OLS residuals from a regression of monthly volume
on a constant and a linear trend. Under the null of stable parameters the
path stays within the dashed 5% bands.
:::

The Sup-F test rejects parameter stability (p < 0.001). The first break
chosen by the Bai–Perron procedure falls in `r kn$bp1`, with a 95%
confidence interval from `r kn$bp1_lo` to `r kn$bp1_hi`, within three
months of the launch. On the log scale the single most pronounced break is
in `r kn$supf_log_date`. The second Bai–Perron break, in `r kn$bp2`
(`r kn$bp2_lo`–`r kn$bp2_hi`), is the date the Sup-F test picks on the
level series, where the growth of 2024–2026 dominates. It marks a further acceleration during the months in which the EU AI Act was
adopted and entered into force, which our design does not separate from
other developments in 2024. The volume data therefore show two regime
changes, the first of which coincides with the ChatGPT launch.

## Event-Study Coefficients

@fig-event-study-frames shows the event-study coefficients for each frame:
the change in the frame's monthly share, in three-month bins, relative to
the three months before the launch.

```{r}
#| label: fig-event-study-frames
#| fig-cap: "Event-study coefficients by frame"
#| fig-height: 10
#| fig-width: 11

ggplot(es_results, aes(x = bin_start, y = estimate)) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.3) +
  geom_vline(xintercept = CHATGPT_DATE, linetype = "dotted", linewidth = 0.4) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.25, fill = "gray60") +
  geom_line(linewidth = 0.4) +
  geom_point(size = 1.3) +
  facet_wrap(~ frame, ncol = 2, scales = "free_y") +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  labs(x = NULL, y = "Change in frame share (percentage points)")
```

::: {.figure-note}
**Note:** Coefficients from a regression of the monthly share of items
carrying each frame on three-month event-time bins; the reference bin is
September–November 2022. Points are placed at the first month of each
bin; the dotted line marks December 2022. Bands are 95% confidence intervals
from Newey–West (HAC) standard errors. Appendix B reports the coefficients.
:::

The frames move in different directions and on different time profiles.
Fear and resistance rises most: by `r kn$fear_early` percentage points in
the first six months after the launch and by `r kn$fear_late` points on
average from December 2024 onward, so the shift grew rather than faded.
Skills and reskilling falls by around four points immediately and stays
there (`r kn$skills_late` points from December 2024). Productivity falls in
the first half-year (`r kn$prod_early` points) and then recovers above its
pre-launch level (`r kn$prod_late` points from December 2024). Regulation
rises after the launch and peaks at `r kn$reg_peak` points in the bin
starting `r kn$reg_peak_date`, around the EU AI Act's entry into force and
the start of its first obligations. Job loss, job creation, transformation,
and inequality move by one to two points at most.

## Pre-Trend Validation

A flat pre-launch path is a precondition for reading post-launch movements
as responses to the shock. @tbl-pretrend-tests reports, for each frame, a
joint F-test that the pre-launch bins are equal after allowing for a
linear trend, and a separate test of the pre-launch slope.

```{r}
#| label: tbl-pretrend-tests
#| tbl-cap: "Pre-trend tests by frame"

kbl_note(
  pretrend_tbl |> transmute(Frame, `F (binned)` = fmt2(F_stat), `F p-value` = fmtP(F_p),
                            `Slope (pp/month)` = fmt3(slope), `Slope p-value` = fmtP(slope_p)),
  paste("Sample: pre-launch months (January 2021 to October 2022). F-test: joint null that",
        "three-month bin dummies are zero in a regression of the monthly frame share on the bins",
        "and a linear trend (reference bin: August-October 2022). Slope: coefficient on a linear",
        "trend in a separate regression. Classical OLS standard errors."))
```

Seven of the eight frames pass the binned F-test at the 5% level; the
exception is inequality. Two frames, job creation and transformation, have
a significant negative pre-launch slope, so their post-launch coefficients,
which are measured against the three months before the launch, should be
read with that downward drift in mind. The fear-and-resistance frame, which
drives the main result, passes both tests.

## Average Post-Launch Effects

@tbl-ate-frames summarises each frame with a single number: the
coefficient on a post-launch indicator in a regression of the monthly frame
share on the indicator and a linear trend. With almost four post-launch
years, this specification is demanding: a linear trend fitted to the
whole period absorbs gradual shifts, so it isolates changes in level at the
launch rather than cumulative change. The pre- and post-launch means are
shown alongside.

```{r}
#| label: tbl-ate-frames
#| tbl-cap: "Average post-launch change in frame shares"

kbl_note(
  ate_tbl |> transmute(Frame, `Pre mean (%)` = fmt2(pre), `Post mean (%)` = fmt2(post_mean),
                       `Post coefficient (pp)` = fmt2(ate), SE = fmt2(se), p = fmtP(p)),
  paste("Dependent variable: monthly share of items carrying the frame, in percentage points.",
        "Post coefficient: coefficient on a post-launch indicator (December 2022 onward) in a",
        "regression with a linear time trend. Newey-West (HAC) standard errors.",
        "Frames are ordered by the absolute size of the coefficient."))
```

```{r}
#| label: fig-ate-forest
#| fig-cap: "Post-launch coefficients by frame"
#| fig-height: 4.5

ggplot(ate_tbl, aes(x = ate, y = reorder(Frame, ate))) +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.3) +
  geom_errorbarh(aes(xmin = ate - 1.96 * se, xmax = ate + 1.96 * se), height = 0.25, linewidth = 0.5) +
  geom_point(size = 2.4) +
  labs(x = "Change in frame share (percentage points)", y = NULL)
```

::: {.figure-note}
**Note:** Point estimates and 95% confidence intervals from @tbl-ate-frames.
:::

Only the skills frame changes significantly by this measure: it falls by
`r fmt1(abs(ate_("SKILLS", "ate")))` percentage points (p =
`r kn$skills_p`). The raw means tell a fuller story. Fear and resistance
rises from `r fmt1(ate_("FEAR_RESISTANCE", "pre"))`% of items before the
launch to `r fmt1(ate_("FEAR_RESISTANCE", "post_mean"))`% after it, and
job loss from `r fmt1(ate_("JOB_LOSS", "pre"))`% to
`r fmt1(ate_("JOB_LOSS", "post_mean"))`%. The trend-adjusted coefficients for
these frames are small because the increase builds up over the post-launch
period rather than arriving at once, as the event study showed.

## Threat and Opportunity

The three threat-leaning frames and the three opportunity-leaning frames
are combined into two composite indices: the share of items carrying at
least one frame of each kind. @fig-threat-vs-opportunity plots both over
time.

```{r}
#| label: fig-threat-vs-opportunity
#| fig-cap: "Threat and opportunity indices over time"
#| fig-height: 5

composite_long <- composite_monthly |>
  select(year_month, threat, opportunity) |>
  pivot_longer(-year_month, names_to = "index", values_to = "pct") |>
  mutate(index = factor(index, levels = c("threat", "opportunity"),
                        labels = c("Threat (job loss, fear, inequality)",
                                   "Opportunity (job creation, productivity, transformation)")))
ggplot(composite_long, aes(x = year_month, y = pct, linetype = index, group = index)) +
  geom_line(linewidth = 0.4, alpha = 0.45, color = "gray45") +
  geom_smooth(method = "loess", span = 0.3, se = FALSE, linewidth = 1, color = "black") +
  geom_vline(data = events_df, aes(xintercept = date), linetype = "dotted", linewidth = 0.4) +
  scale_linetype_manual(values = c("solid", "dashed")) +
  scale_x_date(date_breaks = "6 months", date_labels = "%b\n%Y") +
  labs(x = NULL, y = "% of items")
```

::: {.figure-note}
**Note:** Monthly shares of items carrying at least one threat frame or at
least one opportunity frame. Thin lines: monthly values; thick lines: LOESS
smooths (span 0.3). Dotted lines mark the ChatGPT launch and the secondary
events listed in @fig-volume-shock.
:::

Opportunity framing is more common than threat framing throughout the
sample, but the distance between them shrinks after the launch. The threat
share rises from `r kn$threat_pre`% to `r kn$threat_post`% and reaches
`r kn$threat_last`% in `r kn$last_year`; the opportunity share moves from
`r kn$opp_pre`% to `r kn$opp_post`%. The threat–opportunity gap therefore
narrows from `r kn$gap_pre` to `r kn$gap_post` percentage points. The
asymmetry the conceptual framework predicts is an asymmetry in *changes*:
threat framing gains and opportunity framing does not.

@tbl-asymmetry-test reports the trend-adjusted post-launch coefficients for
the two indices and for the gap. As with the individual frames, a linear
trend fitted to the whole period absorbs much of the gradual shift: the
coefficient on the gap is `r kn$gap_ate` points (p = `r kn$gap_ate_p`).
Estimated in windows around the launch, which are less exposed to later
events, the threat index is `r kn$w18` points higher within 18 months
(p = `r kn$w18_p`) and `r kn$w24` points higher within 24 months
(p = `r kn$w24_p`); see @tbl-window-robustness.

```{r}
#| label: tbl-asymmetry-test
#| tbl-cap: "Post-launch change in the threat and opportunity indices"

asym_tbl <- bind_rows(lapply(names(comp_fit), function(y) {
  r <- comp_fit[[y]]
  tibble(Index = c(threat = "Threat", opportunity = "Opportunity", gap = "Gap (threat - opportunity)")[y],
         `Pre mean (%)` = fmt1(mean(composite_monthly[[y]][composite_monthly$post == 0])),
         `Post mean (%)` = fmt1(mean(composite_monthly[[y]][composite_monthly$post == 1])),
         `Post coefficient` = fmt2(r["post", 1]), SE = fmt2(r["post", 2]), p = fmtP(r["post", 4]),
         `Trend (pp/month)` = fmt3(r["t", 1]))
}))
kbl_note(asym_tbl, paste(
  "Dependent variable: monthly share of items (percentage points). Post coefficient: coefficient",
  "on a post-launch indicator in a regression with a linear time trend. Newey-West (HAC) standard errors."))
```

## Platform Heterogeneity

Web portals dominate the corpus (@tbl-platforms). @fig-platform-volume shows
monthly volume on each platform with at least 100 items.

```{r}
#| label: fig-platform-volume
#| fig-cap: "Monthly item volume by platform"
#| fig-height: 8

platform_monthly <- corpus_data |>
  count(year_month, SOURCE_TYPE) |>
  group_by(SOURCE_TYPE) |> filter(sum(n) >= 100) |> ungroup()
ggplot(platform_monthly, aes(x = year_month, y = n)) +
  geom_col(fill = "gray65") +
  geom_vline(xintercept = CHATGPT_DATE, linetype = "dotted", linewidth = 0.4) +
  facet_wrap(~ SOURCE_TYPE, scales = "free_y", ncol = 3) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  labs(x = NULL, y = "Items per month")
```

::: {.figure-note}
**Note:** Items per month on each platform with at least 100 corpus items;
vertical scales differ across panels. The dotted line marks December 2022.
Instagram, print, radio and television items appear in the database only
from 2024.
:::

@tbl-platform-ate applies the interrupted time series to each platform with
at least 30 pre-launch items. On web portals volume both jumps at the
launch and accelerates afterwards. On Facebook, YouTube and Reddit the
level shift is not significant but the trend increases, as in the aggregate
series; forum volume shows neither. Platforms
that enter the database only after the launch are listed with their
post-launch means but cannot be used for a before–after comparison.

```{r}
#| label: tbl-platform-ate
#| tbl-cap: "Volume by platform before and after the launch"

kbl_note(
  platform_its |> transmute(Platform, Items = fmtN(total),
                            `Pre (per month)` = fmt1(pre_mo), `Post (per month)` = fmt1(post_mo),
                            `Level shift` = ifelse(is.na(level), "—", fmt1(level)),
                            `p (level)` = ifelse(is.na(level_p), "—", fmtP(level_p)),
                            `Trend change` = ifelse(is.na(slope), "—", fmt2(slope)),
                            `p (trend)` = ifelse(is.na(slope_p), "—", fmtP(slope_p))),
  paste("Pre and post: mean items per month before and from December 2022. Level shift and trend",
        "change: interrupted-time-series coefficients as in the aggregate model, with Newey-West (HAC)",
        "standard errors, estimated only for platforms with at least 30 pre-launch items (—: fewer)."))
```

## Outlet-Type Heterogeneity

Prediction 3 is that tabloids amplify threat framing more than quality
outlets. We test it with a difference-in-differences on web items from the
named tabloid and quality outlets, with the item-level threat indicator as
the outcome and year-month fixed effects absorbing shocks common to both
groups.

```{r}
#| label: tbl-did-outlet
#| tbl-cap: "Difference-in-differences: tabloid versus quality outlets"

.ct_h <- coeftable(did_hetero); .ct_2 <- coeftable(did_2way); .ct_n1 <- coeftable(did_n1)
did_tbl <- tibble(
  Term = c("Post-launch × Tabloid", "Tabloid"),
  Estimate = fmt4(.ct_h[c("post_tab", "tabloid"), 1]),
  `SE (robust)` = formatC(.ct_h[c("post_tab", "tabloid"), 2], format = "f", digits = 4),
  `SE (two-way cluster)` = formatC(.ct_2[c("post_tab", "tabloid"), 2], format = "f", digits = 4),
  `p (two-way cluster)` = fmtP(.ct_2[c("post_tab", "tabloid"), 4]),
  `p (wild bootstrap)` = c(fmtP(did_wild$p), "—"))
kbl_note(did_tbl, paste0(
  "Dependent variable: 1 if the item carries a threat frame (job loss, fear/resistance or inequality). ",
  "Sample: web items from the ", did_wild$G, " named tabloid and quality outlets (",
  fmtN(nrow(did_data)), " items). Year-month fixed effects absorb the post-launch main effect. ",
  "Two-way clustering by outlet and year-month. Wild bootstrap: restricted wild-cluster bootstrap ",
  "clustered by outlet, Webb weights, 9,999 draws. Reclassifying N1 as a tabloid gives an ",
  "interaction of ", fmt4(.ct_n1["post_tab", 1]), " (p = ", fmtP(.ct_n1["post_tab", 4]), ")."))
```

The interaction is negative: after the launch, tabloids increased threat
framing by `r abs(as.numeric(kn$did_coef))` percentage points *less* than
quality outlets (two-way clustered p = `r kn$did_p`; wild-cluster bootstrap
p = `r kn$did_p_wild`). The raw shares show why. Tabloids started from a
higher threat share (`r kn$tab_pre`%) and rose modestly to
`r kn$tab_post`%, while quality outlets rose from `r kn$qual_pre`% to
`r kn$qual_post`%. Prediction 3 is rejected: the threat shift is not a
tabloid phenomenon, and it is strongest in the quality press. @fig-tabloid-vs-quality shows the two series.

```{r}
#| label: fig-tabloid-vs-quality
#| fig-cap: "Monthly threat-frame share, tabloid and quality outlets"
#| fig-height: 5

did_panel <- did_data |>
  group_by(outlet_type, year_month) |>
  summarise(threat_share = 100 * mean(threat_i), .groups = "drop") |>
  mutate(outlet_type = factor(outlet_type, levels = c("Tabloid", "Quality")))
ggplot(did_panel, aes(x = year_month, y = threat_share)) +
  geom_line(linewidth = 0.4, alpha = 0.45, color = "gray45") +
  geom_smooth(method = "loess", span = 0.4, se = FALSE, linewidth = 1, color = "black") +
  geom_vline(xintercept = CHATGPT_DATE, linetype = "dotted", linewidth = 0.4) +
  facet_wrap(~ outlet_type, ncol = 2) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  labs(x = NULL, y = "Threat-frame share (%)")
```

::: {.figure-note}
**Note:** Share of web items from the named tabloid (24sata, Index.hr,
Net.hr) and quality (Jutarnji list, Večernji list, Dnevnik.hr, N1, Telegram,
Tportal) outlets that carry a threat frame. Thin lines: monthly values;
thick lines: LOESS smooths. The dotted line marks December 2022.
:::

@tbl-type-effects extends the comparison to all outlet types, estimating the
trend-adjusted post-launch change in each type's monthly threat and
opportunity shares.

```{r}
#| label: tbl-type-effects
#| tbl-cap: "Post-launch change in threat and opportunity shares by outlet type"

kbl_note(
  type_tbl |> transmute(Type, Items = fmtN(items),
                        `Threat coefficient` = fmt2(thr_ate), `p (threat)` = fmtP(thr_p),
                        `Opportunity coefficient` = fmt2(opp_ate), `p (opportunity)` = fmtP(opp_p)),
  paste("For each outlet type: coefficient on a post-launch indicator in a regression of the monthly",
        "threat (or opportunity) share on the indicator and a linear trend, with Newey-West (HAC)",
        "standard errors. Pre- and post-launch means are in the descriptive table by outlet type."))
```

Within outlet types, the monthly series are short and noisy, and none of
the trend-adjusted threat coefficients is significant. The opportunity
share falls significantly for tabloids and for non-web platforms. Together
with the difference-in-differences, this points to a broad-based shift in
coverage rather than one concentrated in a single segment of the market.

## Mechanisms

We consider three channels through which the threat shift could arise.

**Wire copy.** If Croatian outlets imported the threat tilt by
republishing international agency copy, near-identical headlines would
appear in several outlets within a few days, and the share of such items
would rise after the launch. We flag an item as a near-duplicate if its
title shares at least 70% of its words (Jaccard similarity on words of
four or more letters) with an item from another source within three days.

```{r}
#| label: tbl-wire-share
#| tbl-cap: "Near-duplicate items before and after the launch"

kbl_note(
  wire_tbl |> transmute(Period, Items = fmtN(items), `Near-duplicates` = fmtN(dups),
                        `Share (%)` = fmt1(share),
                        `Threat share, near-duplicates (%)` = fmt1(threat_dup),
                        `Threat share, other items (%)` = fmt1(threat_nodup)),
  paste("Near-duplicate: title-word Jaccard similarity of at least 0.7 with an item from a",
        "different source published within three days. Candidate pairs are restricted to items",
        "sharing one of their two rarest title words. Items with fewer than three title words are excluded."))
```

The near-duplicate share falls from `r kn$wire_pre`% before the launch to
`r kn$wire_post`% after it, and near-duplicates carry *less* threat framing
than other items in both periods. Copied material is therefore not the
source of the threat shift; it is concentrated in original coverage.

**Editorial selection.** The difference-in-differences shows that the
threat shift is not driven by tabloids: quality outlets moved further. Any
editorial-selection explanation therefore has to operate across the market
rather than in its most attention-driven segment.

**Reader demand.** The monitoring database records reader interactions
(reactions, comments and shares) for web articles, which lets us test
whether threat-framed articles attract more engagement. @tbl-engagement
regresses log interactions on the two frame indicators with outlet and
year-month fixed effects, so the comparison is between articles from the
same outlet in the same month.

```{r}
#| label: tbl-engagement
#| tbl-cap: "Reader interactions and framing, web articles"

.ce <- coeftable(eng_all); .cp <- coeftable(eng_pos)
eng_tbl <- tibble(
  ` ` = c("Threat frame", "", "Opportunity frame", "", "Observations"),
  `(1) log(1 + interactions)` = c(fmt3(.ce["threatTRUE", 1]), paste0("(", fmt3(.ce["threatTRUE", 2]), ")"),
                                  fmt3(.ce["opportunityTRUE", 1]), paste0("(", fmt3(.ce["opportunityTRUE", 2]), ")"),
                                  fmtN(nobs(eng_all))),
  `(2) log(interactions), interactions > 0` = c(fmt3(.cp["threatTRUE", 1]), paste0("(", fmt3(.cp["threatTRUE", 2]), ")"),
                                                fmt3(.cp["opportunityTRUE", 1]), paste0("(", fmt3(.cp["opportunityTRUE", 2]), ")"),
                                                fmtN(nobs(eng_pos))))
kbl_note(eng_tbl, paste(
  "Web items with a recorded interaction count. Interactions: the monitoring vendor's total of",
  "reactions, comments and shares. Both columns include source (outlet) and year-month fixed effects;",
  "standard errors clustered by source in parentheses. Column 2 drops items with zero interactions,",
  "which may reflect missing measurement rather than no engagement."))
```

Threat-framed articles receive about `r kn$eng_threat`% more interactions
than other articles from the same outlet and month (p = `r kn$eng_threat_p`),
and opportunity-framed articles about `r abs(as.numeric(kn$eng_opp))`% fewer
(p = `r kn$eng_opp_p`). The estimates are correlations, not causal effects
of framing on engagement, and the vendor's interaction measure is
undocumented. They are, however, what the demand-side mechanism predicts,
and they give editors a reason to supply threat framing that does not depend
on outlet type.

## Persistence

@fig-persistence plots three-month moving averages of the two indices, and
@tbl-decay tests whether each series trends after the launch.

```{r}
#| label: fig-persistence
#| fig-cap: "Threat and opportunity indices, three-month moving averages"
#| fig-height: 5

persistence_long <- composite_monthly |>
  arrange(year_month) |>
  mutate(Threat = zoo::rollmean(threat, 3, fill = NA, align = "right"),
         Opportunity = zoo::rollmean(opportunity, 3, fill = NA, align = "right")) |>
  select(year_month, Threat, Opportunity) |>
  pivot_longer(-year_month, names_to = "index", values_to = "pct")
ggplot(persistence_long, aes(x = year_month, y = pct, linetype = index)) +
  geom_line(linewidth = 0.8) +
  geom_vline(xintercept = CHATGPT_DATE, linetype = "dotted", linewidth = 0.4) +
  scale_linetype_manual(values = c(Threat = "solid", Opportunity = "dashed")) +
  scale_x_date(date_breaks = "6 months", date_labels = "%b\n%Y") +
  labs(x = NULL, y = "% of items (three-month moving average)")
```

::: {.figure-note}
**Note:** Right-aligned three-month moving averages of the monthly threat
and opportunity shares. The dotted line marks December 2022.
:::

```{r}
#| label: tbl-decay
#| tbl-cap: "Post-launch trends in frame shares"

.pattern <- function(s, p) ifelse(p >= 0.10, "No significant trend", ifelse(s > 0, "Rising", "Falling"))
kbl_note(
  decay_tbl |> mutate(Series = recode(Series, threat = "Threat index", opportunity = "Opportunity index",
                                      gap = "Gap (threat - opportunity)")) |>
    transmute(Series, `Level at launch (%)` = fmt2(level0), `Slope (pp/month)` = fmt3(slope),
              SE = fmt3(se), p = fmtP(p), Pattern = .pattern(slope, p)),
  paste("Sample: post-launch months only (December 2022 onward). Regression of each monthly share on",
        "months since the launch; the intercept is the fitted share in December 2022.",
        "Newey-West (HAC) standard errors. Pattern uses the 10% level."))
```

Nothing that rose after the launch fades. Fear and resistance, productivity
and regulation keep rising, skills keeps drifting down, and both composite
indices trend upward, the threat index faster than the opportunity index,
although the difference between their slopes is not significant. The
post-launch shift in coverage is not a novelty spike but a change of
direction that continued for almost four years.

# Robustness

## Event Windows

The full-sample estimates compare almost four post-launch years with less
than two pre-launch years, and a linear trend fitted to the whole period
absorbs gradual change. @tbl-window-robustness re-estimates the
trend-adjusted change in the threat index in symmetric windows around the
launch.

```{r}
#| label: tbl-window-robustness
#| tbl-cap: "Threat-index change in symmetric windows around the launch"

kbl_note(window_tbl |> transmute(Window, `Post coefficient (pp)` = fmt2(est), SE = fmt2(se), p = fmtP(p)),
         paste("Dependent variable: monthly share of items carrying a threat frame (percentage points).",
               "Each row uses the months within the stated distance of December 2022. Regression on a",
               "post-launch indicator and a linear trend; Newey-West (HAC) standard errors."))
```

In the narrowest windows the estimate is imprecise, because six or nine
months on each side leave too few monthly observations to separate a level
change from a trend. From 12 months onward the coefficient is positive, at
four to six percentage points within 12 to 24 months of the launch, and
significant at the 10% level at 18 months and at the 5% level at 24 months.
In the ±36-month window, which contains the whole pre-launch period and
therefore the same comparison as the full sample, the coefficient shrinks
again as the trend absorbs the rise.

## Sample, Dictionary and Outlet Perturbations

@tbl-robustness-summary reports the trend-adjusted change in the threat index
under alternative samples and dictionaries. None of these perturbations
changes the conclusion drawn from the full sample: restricting the sample to
web items, dropping January 2023 (the month of euro adoption and Schengen
entry), or dropping any single tabloid leaves the estimate small and
imprecise, so over the full period the shift is gradual rather than a level
break at the launch. Dropping the
fear-and-resistance dictionary, whose terms are the most generic, leaves a
threat index built from job loss and inequality alone, with a coefficient of
`r kn$nofear` points (p = `r kn$nofear_p`).

```{r}
#| label: tbl-robustness-summary
#| tbl-cap: "Threat-index change under alternative samples and dictionaries"

kbl_note(robust_tbl |> transmute(Specification, `Post coefficient (pp)` = fmt2(est), SE = fmt2(se), p = fmtP(p)),
         paste("Dependent variable: monthly share of items carrying a threat frame (percentage points),",
               "full sample period. Regression on a post-launch indicator and a linear trend; Newey-West (HAC)",
               "standard errors. 'Threat without fear/resistance' counts only the job-loss and inequality frames."))
```

## Placebo Test

If the change in December 2022 were a seasonal artefact, a comparable
change should appear at other year-ends. We apply the level-shift
specification to December 2021, using only months before the actual
launch.

```{r}
#| label: tbl-placebo
#| tbl-cap: "Placebo launch date: December 2021"

kbl_note(tibble(Term = c("Intercept", "Placebo post (December 2021 onward)", "Time trend"),
                Estimate = fmt2(placebo_fit[, 1]), SE = fmt2(placebo_fit[, 2]), p = fmtP(placebo_fit[, 4])),
         paste("Dependent variable: items per month. Sample: January 2021 to November 2022.",
               "Newey-West (HAC) standard errors."))
```

The placebo coefficient is `r kn$placebo` items (p = `r kn$placebo_p`).
There is no change at the placebo date.

## Count Models

@tbl-poisson reports the quasi-Poisson volume model with a linear trend,
and Appendix C a negative binomial alternative. Both give the same reading
as @tbl-volume-eventstudy: large post-launch volume relative to the
pre-launch level, most of which a single linear trend over the whole period
absorbs.

```{r}
#| label: tbl-poisson
#| tbl-cap: "Quasi-Poisson model of monthly item volume"

.qp <- broom::tidy(vol_m2)
kbl_note(bind_rows(
  .qp |> transmute(Term = term, Estimate = fmt3(estimate), SE = fmt3(std.error), p = fmtP(p.value),
                   IRR = fmt2(exp(estimate))),
  tibble(Term = "Dispersion parameter", Estimate = fmt2(summary(vol_m2)$dispersion), SE = "", p = "", IRR = "")),
  paste("Dependent variable: items per month. Quasi-Poisson regression on a post-launch indicator and a",
        "linear trend (t, months since January 2021). IRR = exp(coefficient). A dispersion parameter",
        "well above one indicates overdispersion relative to a Poisson model."))
```

# Discussion

Croatian-language coverage of AI and the labour market changed direction at
the ChatGPT launch. Volume did not jump in December 2022; its trend did, by
`r kn$its_slope` items per month, and the first data-driven break falls in
`r kn$bp1`. A second acceleration followed in `r kn$bp2`. By `r kn$last_year`
coverage ran at `r kn$last_year_mean` items a month, against
`r kn$pre_mean` before the launch.

The tone shifted with the volume. The threat share rose from
`r kn$threat_pre`% to `r kn$threat_post`% while the opportunity share stayed
flat; fear-and-resistance framing kept growing for more than three years;
and skills and reskilling framing fell and did not recover. Opportunity
framing remains more common than threat framing in levels, so the
information environment is not dominated by threat. What changed is the
balance: generative AI brought a rise in threat framing and a decline in
adjustment framing, not a symmetric increase in both.

Two results qualify the conceptual framework. First, the shift was gradual.
A model that allows only a level change at the launch and a common trend
finds little, while windows of 18 to 24 months around the launch and the
event-study coefficients show a clear rise. Second, quality outlets, not
tabloids, moved furthest. The tabloidisation prediction fails; the
reader-engagement result suggests one explanation. Threat framing attracts
more interactions across outlets, not only in tabloids, so the incentive to
supply it is market-wide. Why quality outlets responded most, for instance
because they publish more AI and economy stories in which to use it, is a
question for further work.

The decline of skills framing is the result with the clearest policy
relevance. Workers who learn about generative AI from the media receive
more signals about risk and fewer about adjustment than before the launch,
precisely when reskilling is the most direct response available to them.
For Croatia, where firm-level data on AI adoption are sparse, that balance
of signals is a measurable input for active labour-market policy and adult
education. `[AUTHOR INPUT REQUIRED: specific Croatian levers — HZZ training
measures, lifelong-learning programmes, national implementation of the EU
AI Act — to sharpen this paragraph.]`

The evidence has limits. The dictionaries are not validated against human
coding. The corpus measures supply, not exposure or belief. The long
post-period contains later events, including GPT-4 and the EU AI Act, that
our design does not separate from the launch. And the interaction measure
behind the demand-side result is a vendor metric of undocumented
construction.

# Conclusion

ChatGPT's release set Croatian coverage of AI and the labour market on a
new path. Coverage accelerated rather than jumped; threat framing, led by
fear and resistance, rose and kept rising for almost four years; and
skills framing fell. The shift was broad-based: quality outlets moved
further than tabloids, and threat-framed articles attract more reader
engagement across the market.

Two extensions follow. The first is comparative: applying the same design
to other small European language markets would show whether this pattern
is specific to Croatia. The second links supply to demand: combining
coverage with survey data on workers' expectations would show whether the
shift in framing moves beliefs about job security and the willingness to
retrain.

# Appendix

## A. Dictionaries and Validation

```{r}
#| label: tbl-app-dictionaries
#| tbl-cap: "Frame dictionaries"

dict_long <- tibble(Frame = rep(names(frame_dictionaries), lengths(frame_dictionaries)),
                    Keyword = unlist(frame_dictionaries, use.names = FALSE)) |>
  group_by(Frame) |>
  summarise(`N keywords` = n(), Keywords = paste(Keyword, collapse = ", "), .groups = "drop") |>
  arrange(match(Frame, c("JOB_LOSS", "FEAR_RESISTANCE", "INEQUALITY", "JOB_CREATION",
                         "PRODUCTIVITY", "TRANSFORMATION", "SKILLS", "REGULATION")))
kbl_note(dict_long, paste("An item carries a frame if any keyword occurs in its lower-cased title or text.",
                          "Threat-leaning frames first, opportunity-leaning frames second, the other two last.",
                          "Dictionaries are defined in config.yml."))
```

**Validation.** The chunk below reads `data/validation/human_coded.csv` if
present (columns `article_id`, `frame`, `human_label`, `machine_label`) and
reports agreement, precision, recall and Cohen's κ.

```{r}
#| label: tbl-app-icr
#| tbl-cap: "Agreement with human coding"

VAL_PATH <- file.path(PROJECT_ROOT, "data", "validation", "human_coded.csv")
if (file.exists(VAL_PATH)) {
  val <- utils::read.csv(VAL_PATH)
  val |>
    group_by(frame) |>
    summarise(n = n(), agreement = mean(human_label == machine_label),
              precision = sum(human_label == 1 & machine_label == 1) / max(sum(machine_label == 1), 1),
              recall = sum(human_label == 1 & machine_label == 1) / max(sum(human_label == 1), 1),
              pe = mean(human_label) * mean(machine_label) + (1 - mean(human_label)) * (1 - mean(machine_label)),
              kappa = (agreement - pe) / (1 - pe), .groups = "drop") |>
    select(-pe) |>
    kbl_note("Human labels from a stratified random sample of corpus items.", digits = 3)
} else {
  cat("[AUTHOR INPUT REQUIRED] Human coding of a stratified random sample of about 200 items is needed.",
      "Place it at data/validation/human_coded.csv with columns article_id, frame, human_label,",
      "machine_label; this table then reports agreement, precision, recall and Cohen's kappa.\n")
}
```

## B. Event-Study Coefficients

```{r}
#| label: tbl-app-eventstudy-coefs
#| tbl-cap: "Event-study coefficients at selected bins"

es_results |>
  filter(bin %in% c(-4, -2, 0, 2, 4, 8, 12, max(bin))) |>
  transmute(Frame = frame, `Bin starts` = format(bin_start, "%b %Y"),
            Estimate = fmt2(estimate), SE = fmt2(se), p = fmtP(p)) |>
  kbl_note(paste("Dependent variable: monthly frame share (percentage points). Coefficients on three-month",
                 "event-time bins; reference bin September-November 2022. Newey-West (HAC) standard errors.",
                 "These coefficients are plotted in the event-study figure."))
```

## C. Alternative Specifications

```{r}
#| label: tbl-app-altspec-nb
#| tbl-cap: "Negative binomial model of monthly item volume"

broom::tidy(vol_nb) |>
  transmute(Term = term, Estimate = fmt3(estimate), SE = fmt3(std.error), p = fmtP(p.value),
            IRR = fmt2(exp(estimate))) |>
  bind_rows(tibble(Term = "Theta", Estimate = fmt2(vol_nb$theta), SE = "", p = "", IRR = "")) |>
  kbl_note("Dependent variable: items per month. Regression on a post-launch indicator and a linear trend. Smaller theta means more overdispersion.")
```

```{r}
#| label: tbl-app-altspec-lpm
#| tbl-cap: "Item-level linear probability model for the threat indicator"

.cl <- coeftable(lpm_fit)
tibble(Term = c("Post-launch", "Linear trend (months)"),
       Estimate = fmt4(.cl[, 1]), SE = fmt4(.cl[, 2]), p = fmtP(.cl[, 4])) |>
  kbl_note(paste0("Dependent variable: 1 if the item carries a threat frame. Source and month-of-year fixed ",
                  "effects (year-month fixed effects would absorb the post-launch indicator). Standard errors ",
                  "clustered by source. N = ", fmtN(nobs(lpm_fit)), "."))
```

## D. Sample Excerpts by Frame

```{r}
#| label: tbl-app-excerpts
#| tbl-cap: "Sample excerpts by frame"

bind_rows(lapply(names(frame_dictionaries), function(f) {
  pat <- paste(frame_dictionaries[[f]], collapse = "|")
  corpus_data |>
    filter(.data[[paste0("frame_", f)]], !is.na(FULL_TEXT), SOURCE_TYPE == "web") |>
    mutate(hits = stri_count_regex(.text_lower, pat)) |>
    arrange(desc(hits)) |> slice_head(n = 3) |>
    transmute(Frame = f, Outlet = FROM, Date = format(DATE, "%Y-%m-%d"),
              Excerpt = stri_sub(FULL_TEXT, 1, 200))
})) |>
  kbl_note("Three web items per frame with the most dictionary matches; first 200 characters of the text.")
```

## E. Outlet-Level Statistics

```{r}
#| label: tbl-app-outlet-stats
#| tbl-cap: "Web sources with at least 50 items"

corpus_data |>
  filter(SOURCE_TYPE == "web") |>
  group_by(Source = FROM, Type = as.character(outlet_type)) |>
  summarise(Total = n(), Pre = sum(post_chatgpt == 0), Post = sum(post_chatgpt == 1),
            `Threat (%)` = fmt1(100 * mean(threat)), `Opportunity (%)` = fmt1(100 * mean(opportunity)),
            .groups = "drop") |>
  filter(Total >= 50) |>
  arrange(desc(Total)) |>
  kbl_note("Web sources (domains) with at least 50 corpus items. Pre and post split at December 1, 2022. Type as in the outlet classification; Other web for unlisted domains.")
```

## F. Replication

The corpus (`data/raw/ai_labour_corpus_2021_2026.rds`) is produced by
`R/01_extract_corpus.R` from the merged Determ database, which is
proprietary and not distributed. This document renders from
`R/04_paper1_chatgpt_natural_experiment.qmd`. Frame dictionaries, the SQL and
regular-expression filters, the outlet list, and the event dates are defined
in `config.yml`; shared functions, including the outlet classification and
the wild-cluster bootstrap, are in `R/00_helpers.R`.

```{r}
#| label: session-info

sessionInfo()
```
