---
title: "The ChatGPT Shock and Public Information about AI and Work"
subtitle: "Event-Study Evidence from Croatian Digital Media (2021–2024)"
author:
  - name: Petra Palić
    affiliations:
      - name: Croatian Catholic University
        department: University Department of Sociology
  - name: Luka Šikić
    affiliations:
      - name: Croatian Catholic University
        department: University Department of Communication Studies
  - name: Antea Barišić
    affiliations:
      - name: University of Zagreb
        department: Faculty of Economics and Business
date: today
bibliography: ../references.bib
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    code-fold: true
    code-tools: true
    code-summary: "Show code"
    df-print: paged
    fig-width: 10
    fig-height: 6
    fig-dpi: 300
    embed-resources: true
execute:
  warning: false
  message: false
  echo: false

---

```{r}
#| label: setup
#| include: false

# ==========================================================================
# PACKAGES
# ==========================================================================

required_packages <- c(
  "dplyr", "tidyr", "stringr", "stringi", "lubridate", "forcats", "tibble",
  "ggplot2", "scales", "patchwork", "ggrepel",
  "knitr", "kableExtra",
  "fixest",        # fast fixed effects estimation
  "sandwich",      # robust standard errors
  "lmtest",        # coefficient tests
  "strucchange",   # structural break detection
  "MASS",          # negative binomial
  "broom",         # tidy model output
  "zoo"            # rolling functions
)

for (pkg in required_packages) {
  if (!require(pkg, character.only = TRUE, quietly = TRUE)) {
    install.packages(pkg, quiet = TRUE)
    library(pkg, character.only = TRUE)
  }
}

options(dplyr.summarise.inform = FALSE, scipen = 999)

# --------------------------------------------------------------------------
# THEME — black-and-white, publication-ready (greyscale only)
# --------------------------------------------------------------------------

theme_pub <- theme_classic(base_size = 11) +
  theme(
    plot.title       = element_blank(),
    plot.subtitle    = element_blank(),
    plot.caption     = element_text(color = "gray30", size = 9, hjust = 0),
    legend.position  = "bottom",
    legend.title     = element_blank(),
    panel.grid.major = element_line(color = "gray90", linewidth = 0.3),
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    strip.text       = element_text(face = "bold", size = 10),
    axis.title       = element_text(size = 11),
    axis.text        = element_text(color = "black")
  )
theme_set(theme_pub)

# Greyscale palette for two-series plots (threat vs. opportunity, etc.)
bw_two   <- c("black", "gray55")
bw_three <- c("black", "gray45", "gray70")
```

```{r}
#| label: compute-key-numbers
#| include: false

# Headline numbers used inline in the abstract, introduction, and discussion.
# Recomputes a subset of the regressions reported in the tables further down
# so the prose magnitudes update with the data. Cross-check on render against
# tbl-poisson, tbl-ate-frames, tbl-asymmetry-test, and tbl-did-outlet.

CORPUS_PATH <- "../data/raw/ai_labour_corpus.rds"
if (!file.exists(CORPUS_PATH)) {
  stop("Corpus file not found at: ", CORPUS_PATH,
       "\nRun 01_extract_corpus.R first.")
}
corpus_data <- readRDS(CORPUS_PATH) |>
  mutate(
    DATE = as.Date(DATE),
    .text_lower = stri_trans_tolower(
      paste(coalesce(TITLE, ""), coalesce(FULL_TEXT, ""), sep = " ")
    ),
    year         = year(DATE),
    month        = month(DATE),
    year_month   = floor_date(DATE, "month"),
    year_week    = floor_date(DATE, "week"),
    quarter      = quarter(DATE),
    year_quarter = paste0(year, " Q", quarter),
    word_count   = stri_count_regex(FULL_TEXT, "\\S+")
  ) |>
  filter(!is.na(DATE), DATE < as.Date("2024-01-01")) |>
  distinct(TITLE, DATE, FROM, .keep_all = TRUE) |>
  arrange(DATE)

# Frame dictionaries — duplicated from the Frame Detection section so this
# chunk is self-contained and runs before the Introduction.
.frame_dicts <- list(
  JOB_LOSS = c("gubitak posla","gubitak poslova","gubitak radnih mjesta",
               "ukidanje radnih mjesta","ukidanje poslova",
               "zamjena radnika","zamijeniti radnike","zamjenjuje radnike",
               "istisnuti radnike","istiskivanje","otpuštanje","otpuštanja",
               "nestanak poslova","nestanak zanimanja","suvišan",
               "suvišni radnici","tehnološka nezaposlenost","krade poslove",
               "krade posao","oduzima posao","prijeti radnim mjestima",
               "ugrožava radna mjesta"),
  JOB_CREATION = c("nova radna mjesta","novi poslovi","novo zapošljavanje",
                   "nove prilike","nove mogućnosti","stvaranje poslova",
                   "rast zapošljavanja","povećanje zapošljavanja",
                   "nova zanimanja","nova karijera",
                   "potražnja za stručnjacima","nedostatak radnika",
                   "deficitarna zanimanja"),
  TRANSFORMATION = c("transformacija rada","transformacija poslova",
                     "promjena načina rada","mijenja način rada",
                     "prilagodba","prilagoditi se","prilagođavanje",
                     "nadopunjuje","komplementar","suradnja čovjeka i",
                     "čovjek i stroj","evolucija poslova","evolucija rada",
                     "nove uloge","promijenjena uloga","ne zamjenjuje nego"),
  PRODUCTIVITY = c("produktivnost","povećanje produktivnosti",
                   "učinkovitost","efikasnost","ušteda","smanjenje troškova",
                   "konkurentnost","konkurentna prednost","gospodarski rast",
                   "ekonomski rast","optimizacija"),
  INEQUALITY = c("nejednakost","rastuća nejednakost","digitalni jaz",
                 "tehnološki jaz","socijalna nejednakost","polarizacija",
                 "jaz u plaćama","ranjive skupine","marginalizirani",
                 "srednja klasa","nestanak srednje klase"),
  FEAR_RESISTANCE = c("strah od ai","strah od gubitka","strah od tehnologij",
                      "prijetnja","opasnost","apokalipsa","distopija",
                      "katastrofa","upozorenje","alarm","otpor prema",
                      "protivljenje","neizvjesnost","nesigurnost","panika",
                      "zabrinutost")
)
for (fname in names(.frame_dicts)) {
  pat <- paste(.frame_dicts[[fname]], collapse = "|")
  corpus_data[[paste0("frame_", fname)]] <- stri_detect_regex(
    corpus_data$.text_lower, pat
  )
}

# Outlet classification (subset needed for the DiD)
.outlet_pat <- tibble::tribble(
  ~pattern,    ~outlet_type,
  "24sata",    "Tabloid",
  "index",     "Tabloid",
  "net\\.hr",  "Tabloid",
  "jutarnji",  "Quality",
  "vecernji",  "Quality",
  "dnevnik",   "Quality",
  "n1",        "Quality",
  "tportal",   "Quality",
  "telegram",  "Quality"
)
corpus_data$outlet_type <- NA_character_
.from_lower <- stri_trans_tolower(corpus_data$FROM)
for (i in seq_len(nrow(.outlet_pat))) {
  hit <- stri_detect_regex(.from_lower, .outlet_pat$pattern[i])
  corpus_data$outlet_type[hit & is.na(corpus_data$outlet_type)] <-
    .outlet_pat$outlet_type[i]
}

CHATGPT_DATE <- as.Date("2022-12-01")
corpus_data$post_chatgpt <- as.integer(corpus_data$DATE >= CHATGPT_DATE)

# Monthly volume
monthly <- corpus_data |>
  count(year_month) |>
  filter(!is.na(year_month)) |>
  mutate(
    post = as.integer(year_month >= CHATGPT_DATE),
    t    = as.numeric(difftime(year_month, min(year_month), units = "days")) / 30
  )

# Composite threat / opportunity / gap series
composite_monthly <- corpus_data |>
  group_by(year_month) |>
  summarise(
    n           = n(),
    threat      = sum(frame_JOB_LOSS | frame_FEAR_RESISTANCE | frame_INEQUALITY,
                      na.rm = TRUE) / n() * 100,
    opportunity = sum(frame_JOB_CREATION | frame_PRODUCTIVITY | frame_TRANSFORMATION,
                      na.rm = TRUE) / n() * 100,
    .groups = "drop"
  ) |>
  filter(!is.na(year_month)) |>
  mutate(
    gap  = threat - opportunity,
    post = as.integer(year_month >= CHATGPT_DATE),
    t    = as.numeric(difftime(year_month, min(year_month), units = "days")) / 30
  )

# Headline regressions
.qp        <- glm(n ~ post + t, data = monthly, family = quasipoisson())
.IRR       <- exp(coef(.qp)["post"])
.IRR_se_log <- summary(.qp)$coefficients["post", "Std. Error"]

.fit_threat <- lm(threat      ~ post + t, data = composite_monthly)
.fit_opp    <- lm(opportunity ~ post + t, data = composite_monthly)
.fit_gap    <- lm(gap         ~ post + t, data = composite_monthly)
.ct_threat <- coeftest(.fit_threat, vcov = vcovHAC(.fit_threat))
.ct_opp    <- coeftest(.fit_opp,    vcov = vcovHAC(.fit_opp))
.ct_gap    <- coeftest(.fit_gap,    vcov = vcovHAC(.fit_gap))

.did_data <- corpus_data |>
  filter(outlet_type %in% c("Tabloid", "Quality")) |>
  mutate(
    threat  = as.integer(frame_JOB_LOSS | frame_FEAR_RESISTANCE | frame_INEQUALITY),
    tabloid = as.integer(outlet_type == "Tabloid")
  )
.did_fit <- feols(threat ~ post_chatgpt * tabloid | year_month,
                  data = .did_data, vcov = "hetero")
.did_int <- coeftable(.did_fit)["post_chatgpt:tabloid", ]

.pre_mean  <- mean(monthly$n[monthly$post == 0])
.post_mean <- mean(monthly$n[monthly$post == 1])

# Largest event-study coefficient on threat-leaning frames (post months only)
.threat_es_max <- function(ycol) {
  d <- corpus_data |>
    group_by(year_month) |>
    summarise(y = sum(.data[[ycol]], na.rm = TRUE) / n() * 100,
              .groups = "drop") |>
    filter(!is.na(year_month)) |>
    mutate(
      rel_month = interval(CHATGPT_DATE, year_month) %/% months(1),
      t = as.numeric(difftime(year_month, min(year_month), units = "days")) / 30
    ) |>
    filter(rel_month >= -20 & rel_month <= 17)
  d$rel_fct <- relevel(factor(d$rel_month), ref = "-1")
  m <- lm(y ~ rel_fct + t, data = d)
  cf <- broom::tidy(m) |>
    filter(stringr::str_detect(term, "rel_fct")) |>
    mutate(k = as.integer(stringr::str_extract(term, "-?\\d+"))) |>
    filter(k >= 0)
  if (nrow(cf) == 0) NA_real_ else max(cf$estimate, na.rm = TRUE)
}
.max_threat_coef <- max(c(
  .threat_es_max("frame_JOB_LOSS"),
  .threat_es_max("frame_FEAR_RESISTANCE"),
  .threat_es_max("frame_INEQUALITY")
), na.rm = TRUE)

# Pack
fmt1 <- function(x) formatC(x, format = "f", digits = 1)
fmt2 <- function(x) formatC(x, format = "f", digits = 2)
fmtN <- function(x) format(round(x), big.mark = ",")
kn <- list(
  N               = fmtN(nrow(corpus_data)),
  IRR             = fmt1(.IRR),
  IRR_se          = fmt2(.IRR_se_log),
  pre_mean        = fmt1(.pre_mean),
  post_mean       = fmt1(.post_mean),
  ratio           = fmt1(.post_mean / .pre_mean),
  ate_threat      = fmt1(.ct_threat["post", 1]),
  ate_opp         = fmt1(.ct_opp["post", 1]),
  gap_pp          = fmt1(.ct_gap["post", 1]),
  gap_se          = fmt1(.ct_gap["post", 2]),
  did_coef        = fmt2(.did_int[1] * 100),
  did_p           = format.pval(.did_int[4], digits = 2),
  max_threat_coef = fmt1(.max_threat_coef)
)
stopifnot(all(nzchar(unlist(kn))))
```

# Abstract

ChatGPT's November 30, 2022 release made generative artificial intelligence
visible to a mass audience overnight. We treat the launch as an unanticipated
shock and estimate its effect on Croatian digital media coverage of AI and
labour using an event-study design with monthly data from January 2021 to
December 2023, applied to a corpus of `r kn$N` Croatian-language articles
drawn from approximately 20 million records on the Determ media-monitoring
platform. Monthly article volume jumps by a factor of `r kn$IRR`
(quasi-Poisson, log-scale SE `r kn$IRR_se`); a Sup-F test endogenously locates
the break at December 2022 (p < 0.01), and a December 2021 placebo coefficient
is statistically indistinguishable from zero. Threat framing (job loss, fear,
inequality) rises `r kn$gap_pp` percentage points more than opportunity
framing (job creation, productivity, transformation), and the gap persists
through end-2023. Tabloid outlets activate threat narratives more strongly
than quality outlets in a year-month-fixed-effects difference-in-differences.
The Croatian information environment delivers risk signals more readily than
adjustment signals.

# Introduction

Croatian-language digital coverage of AI and work rose by a factor of
`r kn$IRR` in the year and a half after ChatGPT's release, and its tone
shifted: threat-related framing about job loss, fear, and inequality outpaced
opportunity-related framing about job creation, productivity, and
transformation by `r kn$gap_pp` percentage points, a gap that has not closed.
Whether generative AI ultimately displaces, complements, or transforms human
work remains an open empirical question [@acemoglu2018race; @autor2015why;
@bessen2020automation], and is likely to remain so for years. Workers, firms,
and policymakers must nonetheless act under uncertainty, and the public
information environment that shapes those actions has moved measurably.

Recent task-level evidence suggests large language models are exposed to a
wide and disproportionately white-collar share of the occupational
distribution [@eloundou2023gpts; @felten2021ai], and field experiments find
generative AI raises productivity for some workers while compressing observed
skill premia [@brynjolfsson2023generative]. In small open economies such as
Croatia, where firm-level data on AI adoption is sparse, the main channel
through which workers and policymakers learn about these capabilities is
mediated digital coverage. What that channel transmits — which frames, how
quickly, with what persistence — is therefore a first-order input to belief
formation and labour-market adjustment.

This paper estimates the effect of the November 30, 2022 ChatGPT release on
Croatian digital media coverage of AI and labour. We treat the release as an
unanticipated shock to the salience of generative-AI capabilities and identify
its effect using an event-study design with monthly observations from January
2021 to December 2023, complemented by an interrupted time series for
aggregate volume and a Sup-F test that searches for the break date
endogenously rather than imposing it. The corpus comprises `r kn$N`
Croatian-language articles and posts drawn from approximately 20 million
records on the Determ media-monitoring platform, filtered through a two-stage
AI ∩ labour intersection requirement. Croatia is a useful case: a small open
European economy (population approximately 3.85 million, EU member since 2013,
euro-area member since 2023) with a services-heavy labour market in which
direct microdata on firm-level AI adoption is essentially unavailable, so the
information signals reaching workers come almost entirely through mediated
coverage.

Three findings stand out. First, monthly article volume rises by a factor of
`r kn$IRR` at the launch (quasi-Poisson, log-scale SE `r kn$IRR_se`), pre-launch
counts averaging `r kn$pre_mean` items per month against `r kn$post_mean`
post-launch. The Sup-F test rejects no-break (p < 0.01) and the endogenous
breakpoint coincides with December 2022, while the December 2021 placebo
coefficient is statistically indistinguishable from zero. Second, threat
framing rises `r kn$ate_threat` percentage points post-launch against
`r kn$ate_opp` percentage points for opportunity framing; the
threat–opportunity gap widens by `r kn$gap_pp` pp (HAC SE `r kn$gap_se`) and
shows no monthly decay through end-2023.

Third, the shock propagated unevenly across the digital ecosystem. Web portals
carry the largest absolute volume increase, while Twitter and YouTube show
the largest relative jumps. A difference-in-differences with year-month fixed
effects finds tabloid outlets activate threat framing
`r kn$did_coef` percentage points more than quality outlets after the launch
(p `r kn$did_p`). Pre-trend F tests fail to reject parallel pre-trends for the
headline composite frames at the 10 percent level.

The paper contributes to three strands. To work on automation expectations
[@acemoglu2018race; @autor2022work], it documents the supply side of the
public information that shapes those expectations in a small open European
economy where direct measurement of AI adoption is currently infeasible. To
media-economics work on framing of technology shocks, it provides event-study
evidence with formal break-date inference rather than post-hoc narrative
description. To work on the labour-market consequences of generative AI
[@eloundou2023gpts; @felten2021ai; @brynjolfsson2023generative], it supplies a
measurable input — the asymmetric activation of threat over opportunity
framing — that subsequent work can link to realised vacancy, employment, and
reskilling outcomes.

The remainder of the paper proceeds as follows. Section 2 describes the
corpus, the AI ∩ labour intersection filter, and the eight frame dictionaries.
Section 3 lays out the event-study and interrupted-time-series specifications
and the identifying assumptions. Section 4 reports the volume shock,
frame-level event-study coefficients, and the threat–opportunity asymmetry.
Section 5 examines heterogeneity by platform and outlet type. Section 6
reports robustness — alternative event windows, a quasi-Poisson count model,
and the December 2021 placebo. Section 7 discusses implications for Croatian
active-labour-market policy and the limits of supply-side framing as a measure
of public information.


# Institutional and Conceptual Background

## The Croatian Media Landscape

Croatia is a small open European economy with a digital media market that
has three structural features useful for our identification strategy. First,
ownership is concentrated. Five owner groups (Hanza Media, Styria Media
Group, 24sata d.o.o., RTL Hrvatska, and Telegram media grupa) account for
the majority of digital readership in 2023, and their portfolios cover the
full tabloid–quality spectrum [@aem2023digitalreadership]. Second, the
labour market is services-heavy and exposed to white-collar generative-AI
disruption to a degree comparable with EU averages: employment in
information and communication, financial services, and professional
services sums to roughly one quarter of dependent employment
[@hzz2024labourreport]. Third, microdata on firm-level AI adoption is
essentially unavailable, so the supply-side information environment that
reaches workers is the principal observable channel through which the
ChatGPT shock propagates. The case therefore generalises to a broad class
of small open second-mover language environments — Slovenian, Slovak,
Estonian, Lithuanian, Latvian, Maltese — whose labour markets face the same
exposure but whose information environments are mediated by a similarly
concentrated and similarly second-mover digital press.

The outlet typology we use throughout the paper (Tabloid, Quality, Regional,
Public, Tech, Business) maps onto editorial style and audience composition
rather than ownership. Tabloid outlets (24sata, Index, Net.hr) prioritise
high-volume, attention-driven coverage. Quality outlets (Jutarnji list,
Večernji list, N1, Telegram, Tportal, Dnevnik) run longer analytical pieces
and retain larger newsroom shares of dedicated technology and economy
desks. The Tabloid–Quality contrast is the basis for the
difference-in-differences specification in Section 5.5.

## ChatGPT as a Salience Shock

We treat ChatGPT's November 30, 2022 public release as an unanticipated
shock to the salience of generative-AI capabilities for Croatian editors
and readers. The unanticipation claim rests on three observations. The
release was announced and shipped on the same day, with no Croatian-press
coverage in the preceding weeks. Pre-launch search interest in "ChatGPT"
inside Croatia, as measured by Google Trends, was effectively zero through
late November 2022 and rose sharply only in the first week of December.
The corpus we build for this paper contains zero pre-launch articles
mentioning "ChatGPT," "GPT," or "OpenAI" jointly with any labour-market
term. The Section 4 Threats subsection reports the in-corpus evidence
formally; an external Google Trends figure is flagged below for
incorporation in the next revision.

`[AUTHOR INPUT REQUIRED — Google Trends export for "ChatGPT" and
"umjetna inteligencija" in Croatia (Jan 2021 – Nov 2022) to populate
Figure A1; place CSV at `data/external/google_trends_hr.csv`. The chunk in
Appendix A reads it if present.]`

The treatment date itself is December 1, 2022 — the first full day after
the launch, and therefore the first month for which all daily observations
are post-treatment in the monthly aggregation that the design uses.

## Conceptual Framework

A short verbal model of editorial choice under attention scarcity yields
three predictions that the empirical specifications below test in turn.
Editors face a fixed monthly attention budget and choose article slots to
maximise expected readership, weighted by reader loss-aversion: stories
that promise a vivid downside attract clicks at higher rates than stories
that promise a measured upside, and editors know this. A salience shock
expands the menu of available story angles, increasing the marginal
expected return to AI-and-labour coverage relative to any other topic
competing for the same attention budget.

Three predictions follow. First, aggregate volume rises at the shock,
because the marginal expected return now exceeds the previously binding
opportunity cost. The interrupted-time-series and Sup-F specifications in
Section 5 test this. Second, the share of threat-leaning angles
(displacement, fear, inequality) rises faster than the share of
opportunity-leaning angles (job creation, productivity, transformation),
because reader loss-aversion makes the threat angles more profitable per
slot. The frame-level event study and the threat–opportunity asymmetry
test in Section 6.4 test this. Third, outlets whose business model
depends most heavily on attention-driven monetisation — tabloids — amplify
the threat tilt more than outlets that rely on subscription or
brand-reputation revenue — quality outlets. The
difference-in-differences in Section 6.5 tests this.

The framework is deliberately verbal rather than formal. Its purpose is to
yield testable predictions before the empirical specifications appear, not
to pin down a structural parameter. The three predictions map one-to-one to
the three empirical results that follow.

# Data and Measurement

## Corpus Construction and Loading

The corpus draws on the Determ media-monitoring database, which covers
approximately 20 million records of Croatian-language digital media content
between January 2021 and May 2024. Determ aggregates articles and posts from
web portals, Facebook, Twitter (X), Instagram, YouTube, TikTok, Reddit, and
public forums, capturing what is, to our knowledge, the most comprehensive
record of the Croatian-language digital information environment for the period
we study.

We construct the analytical corpus through a two-stage intersection filter. An
item enters the corpus only if its title or full text contains at least one
term from an AI-and-technology vocabulary *and* at least one term from a
labour-market vocabulary. The first vocabulary covers terms such as
*umjetna inteligencija* (artificial intelligence), *strojno učenje* (machine
learning), ChatGPT, GPT, OpenAI, *automatizacija*, *robotizacija*,
*algoritam*, *neuronska mreža*, *chatbot*, and related expressions. The
second vocabulary covers labour-market terms including *posao* (job),
*zaposlen* (employed), *nezaposlenost* (unemployment), *radno mjesto*
(workplace), *tržište rada* (labour market), *vještine* (skills), *karijera*
(career), *produktivnost* (productivity), *otpuštanje* (lay-offs),
*prekvalifikacija* (reskilling), and *plaća* (wage). The intersection design
substantially reduces the false-positive rate associated with single-keyword
searches, since articles must engage both topics jointly to qualify.

Croatian is a highly inflected language: nouns decline through seven cases,
verbs conjugate, and adjectives agree in gender, number, and case with the
nouns they modify. We accommodate this morphological variation through
stem-based regular expressions (for example, the stem `zaposlen` matches
*zaposleni*, *zaposlenika*, *zaposlenik*, and related inflections; the
pattern `umjetn.*inteligencij` matches all case combinations of *umjetna
inteligencija*).

After deduplication on title, date, and source, and after restricting to the
analysis window 2021–2024, the working corpus comprises the article counts
shown in Table \@ref(tbl-summary).

The corpus records media supply rather than media consumption, and we cannot
directly measure whether readers were exposed to or influenced by the framing
patterns we document. Throughout the paper we treat the corpus as a measure of
the public information environment on the supply side and discuss the
distinction explicitly when interpreting policy implications.

```{r}
#| label: load-corpus

# Corpus already loaded and prepped in the compute-key-numbers chunk above.
cat("Corpus loaded:", format(nrow(corpus_data), big.mark = ","), "articles\n")
cat("Date range:", as.character(min(corpus_data$DATE)), "to",
    as.character(max(corpus_data$DATE)), "\n")
```

## Frame Detection

We operationalize eight interpretive frames, each corresponding to a
distinct way in which media discuss the relationship between AI and labour.
Curated Croatian keyword dictionaries with stem-based regular expressions
detect each frame, accommodating inflection. We code an article as carrying
a given frame if any of the dictionary patterns for that frame matches the
text. The eight frames are grouped naturally into threat-leaning
content, opportunity-leaning content, and orthogonal content.

The threat-leaning frames cover *job loss* (terms such as *gubitak posla*,
*ukidanje radnih mjesta*, *zamjena radnika*, *otpuštanje*, *tehnološka
nezaposlenost*), *fear and resistance* (*strah od AI*, *prijetnja*,
*opasnost*, *neizvjesnost*, *panika*), and *inequality* (*nejednakost*,
*digitalni jaz*, *polarizacija*, *nestanak srednje klase*). The
opportunity-leaning frames cover *job creation* (*nova radna mjesta*,
*potražnja za stručnjacima*, *nova zanimanja*), *productivity* (*učinkovitost*,
*ušteda*, *gospodarski rast*, *optimizacija*), and *task transformation*
(*transformacija rada*, *prilagodba*, *suradnja čovjeka i*, *evolucija
poslova*). The two remaining frames are *skills* (*prekvalifikacija*,
*cjeloživotno učenje*, *digitalne vještine*, *reskilling*, *upskilling*) and
*regulation* (*regulacija AI*, *AI Act*, *zaštita radnika*, *prava radnika*,
*sindikat*).

Frames are not mutually exclusive: an article that discusses both
displacement and reskilling will be coded as carrying both. We treat each
frame as a binary indicator and analyse its monthly prevalence (the share of
articles carrying the frame) over time.

Dictionary detection is unvalidated against a human-coded gold standard, so
per-frame precision and recall are unmeasured. The diagnostic columns built
upstream in `R/02_add_diagnostics.R` (`groupA_hit_count`, `groupB_hit_count`,
`groupA_context`, `groupB_context`) report which keywords matched and where,
which provides transparency but does not eliminate the concern. We note,
however, that any dictionary measurement error is plausibly orthogonal to the
ChatGPT date: it would attenuate estimated effects toward zero rather than
manufacture spurious post-shock activation, so the *direction* of the
threat-versus-opportunity asymmetry we document is unlikely to be an artefact
of the detection method.

```{r}
#| label: frame-detection

frame_dictionaries <- list(
  JOB_LOSS = c(
    "gubitak posla", "gubitak poslova", "gubitak radnih mjesta",
    "ukidanje radnih mjesta", "ukidanje poslova",
    "zamjena radnika", "zamijeniti radnike", "zamjenjuje radnike",
    "istisnuti radnike", "istiskivanje",
    "otpuštanje", "otpuštanja",
    "nestanak poslova", "nestanak zanimanja",
    "suvišan", "suvišni radnici",
    "tehnološka nezaposlenost",
    "krade poslove", "krade posao", "oduzima posao",
    "prijeti radnim mjestima", "ugrožava radna mjesta"
  ),
  JOB_CREATION = c(
    "nova radna mjesta", "novi poslovi", "novo zapošljavanje",
    "nove prilike", "nove mogućnosti",
    "stvaranje poslova",
    "rast zapošljavanja", "povećanje zapošljavanja",
    "nova zanimanja", "nova karijera",
    "potražnja za stručnjacima", "nedostatak radnika",
    "deficitarna zanimanja"
  ),
  TRANSFORMATION = c(
    "transformacija rada", "transformacija poslova",
    "promjena načina rada", "mijenja način rada",
    "prilagodba", "prilagoditi se", "prilagođavanje",
    "nadopunjuje", "komplementar",
    "suradnja čovjeka i", "čovjek i stroj",
    "evolucija poslova", "evolucija rada",
    "nove uloge", "promijenjena uloga",
    "ne zamjenjuje nego"
  ),
  SKILLS = c(
    "prekvalifikacija", "dokvalifikacija",
    "cjeloživotno učenje",
    "digitalna pismenost", "digitalne vještine",
    "nova znanja", "nove vještine",
    "jaz u vještinama", "nedostatak vještina",
    "reskilling", "upskilling",
    "obrazovanje za budućnost",
    "stem vještine", "programiranje"
  ),
  REGULATION = c(
    "regulacija ai", "regulativa",
    "zakon o ai", "zakonski okvir",
    "eu regulativa", "ai act",
    "etička pitanja", "etika ai",
    "sindikat", "sindikalni",
    "zaštita radnika", "prava radnika",
    "socijalna zaštita"
  ),
  PRODUCTIVITY = c(
    "produktivnost", "povećanje produktivnosti",
    "učinkovitost", "efikasnost",
    "ušteda", "smanjenje troškova",
    "konkurentnost", "konkurentna prednost",
    "gospodarski rast", "ekonomski rast",
    "optimizacija"
  ),
  INEQUALITY = c(
    "nejednakost", "rastuća nejednakost",
    "digitalni jaz", "tehnološki jaz",
    "socijalna nejednakost",
    "polarizacija",
    "jaz u plaćama",
    "ranjive skupine", "marginalizirani",
    "srednja klasa", "nestanak srednje klase"
  ),
  FEAR_RESISTANCE = c(
    "strah od ai", "strah od gubitka", "strah od tehnologij",
    "prijetnja", "opasnost",
    "apokalipsa", "distopija", "katastrofa",
    "upozorenje", "alarm",
    "otpor prema", "protivljenje",
    "neizvjesnost", "nesigurnost",
    "panika", "zabrinutost"
  )
)

# Detect frames
for (fname in names(frame_dictionaries)) {
  pattern <- paste(frame_dictionaries[[fname]], collapse = "|")
  corpus_data[[paste0("frame_", fname)]] <- stri_detect_regex(
    corpus_data$.text_lower, pattern
  )
}

frame_cols <- paste0("frame_", names(frame_dictionaries))
```

## Outlet Classification

For the outlet-type heterogeneity analysis, we classify Croatian web sources
into six categories based on the source name in the database. *Tabloid*
outlets (24sata, Index, Net.hr) emphasise high-volume, sensational coverage.
*Quality* outlets (Jutarnji list, Večernji list, N1, Telegram, Tportal,
Dnevnik) typically run longer and more analytical pieces. *Regional* outlets
(Slobodna Dalmacija, Novi list) cover local perspectives. *Public*
broadcasting (HRT) operates under regulatory obligations of balanced
coverage. *Tech* outlets (Bug, Netokracija) target a technologically
literate audience. *Business* outlets (Poslovni dnevnik, Lider, Forbes
Hrvatska) frame stories economically. Sources that do not match any of these
patterns are grouped as *Other*.

Case-insensitive regular expressions on the source name implement the
classification. The mapping is deliberately conservative: each outlet
appears in at most one category, and we leave ambiguous cases in *Other*
rather than force them into a category.

```{r}
#| label: outlet-classification

outlet_patterns <- tribble(
  ~pattern,              ~outlet_type,
  "24sata",              "Tabloid",
  "index",               "Tabloid",
  "net\\.hr",            "Tabloid",
  "jutarnji",            "Quality",
  "vecernji",            "Quality",
  "dnevnik",             "Quality",
  "n1",                  "Quality",
  "tportal",             "Quality",
  "telegram",            "Quality",
  "slobodna.*dalmacija", "Regional",
  "novi.*list",          "Regional",
  "hrt",                 "Public",
  "bug",                 "Tech",
  "netokracija",         "Tech",
  "poslovni",            "Business",
  "lider",               "Business",
  "forbes.*hr",          "Business"
)

corpus_data$outlet_type <- "Other"
for (i in seq_len(nrow(outlet_patterns))) {
  hit <- stri_detect_regex(
    stri_trans_tolower(corpus_data$FROM), outlet_patterns$pattern[i]
  )
  corpus_data$outlet_type[hit] <- outlet_patterns$outlet_type[i]
}

corpus_data$outlet_type <- factor(
  corpus_data$outlet_type,
  levels = c("Tabloid", "Quality", "Regional", "Public",
             "Tech", "Business", "Other")
)
```

## Descriptive Summary

Table \@ref(tbl-summary) summarises the working corpus. The total number of
items, the share covered before and after the ChatGPT launch in December
2022, and the typical article length give a first sense of the data on
which the event-study analysis rests. The pre-ChatGPT and post-ChatGPT
splits in the final two rows already hint at the asymmetric volume shift
that the formal analysis confirms below.

```{r}
#| label: tbl-summary
#| tbl-cap: "Corpus descriptive statistics"

summary_tbl <- tibble(
  Metric = c(
    "Total articles",
    "Unique sources",
    "Date range",
    "Mean word count",
    "Median word count",
    "Pre ChatGPT articles (before Dec 2022)",
    "Post ChatGPT articles (Dec 2022 onward)"
  ),
  Value = c(
    format(nrow(corpus_data), big.mark = ","),
    format(n_distinct(corpus_data$FROM), big.mark = ","),
    paste(min(corpus_data$DATE), "to", max(corpus_data$DATE)),
    round(mean(corpus_data$word_count, na.rm = TRUE)),
    round(median(corpus_data$word_count, na.rm = TRUE)),
    format(sum(corpus_data$DATE < as.Date("2022-12-01")), big.mark = ","),
    format(sum(corpus_data$DATE >= as.Date("2022-12-01")), big.mark = ",")
  )
)

kable(summary_tbl) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

# Empirical Strategy

## Event Definition and Identification

We define the treatment date as December 1, 2022, the first full day after
ChatGPT's public release. The identifying assumption is that the timing of the
ChatGPT release was not anticipated by Croatian media outlets, so a sharp
break in coverage at that date can be attributed to the shock rather than to
a smooth pre-existing trend. Two secondary events are used for robustness
purposes only: the GPT-4 release on March 14, 2023 and the EU AI Act adoption
on March 13, 2024.

```{r}
#| label: event-setup

# Event dates
CHATGPT_DATE  <- as.Date("2022-12-01")
GPT4_DATE     <- as.Date("2023-03-14")
EU_AI_ACT     <- as.Date("2024-03-13")

# Treatment indicator
corpus_data$post_chatgpt <- as.integer(corpus_data$DATE >= CHATGPT_DATE)

# Relative time in months (for event study)
corpus_data$rel_month <- interval(CHATGPT_DATE, corpus_data$year_month) %/% months(1)
```

## Specification

We estimate the following event study specification at the monthly level for
each frame $f$:

$$
Y_{ft} = \alpha_f + \sum_{k \neq -1} \beta_k^f \cdot \mathbf{1}[t - t^* = k] + \gamma_f \cdot t + \varepsilon_{ft}
$$

where $Y_{ft}$ is the share of articles containing frame $f$ in month $t$,
$t^*$ is December 2022, and $k$ indexes months relative to the event. We
normalize to $k = -1$ (November 2022). The coefficient $\beta_k^f$ captures
the change in frame $f$ prevalence in month $k$ relative to the last
pre-treatment month, conditional on a linear time trend.

For aggregate attention we additionally estimate an interrupted time
series (ITS) model. The model regresses monthly article counts on a linear
time trend, a binary indicator for months at or after the ChatGPT launch,
and an interaction between the two. The trend captures gradual pre-existing
growth in coverage. The level-shift coefficient measures the immediate jump
at the launch date. The interaction term captures whether the post-shock
trajectory accelerates or decelerates relative to the pre-shock slope. This
specification is a standard tool in policy-evaluation research for measuring
the effect of a well-defined intervention on an aggregate time series
[@bernal2017interrupted].

We complement it with a Sup-F structural-break test. The test searches for
the most statistically pronounced breakpoint across all candidate dates in
the central 70% of the sample without imposing the ChatGPT date in advance.
A significant Sup-F statistic and a detected breakpoint that coincides with
December 2022 give independent confirmation that the shift concentrates
there rather than spreading across the timeline. A placebo test at December
2021, reported in the Robustness section, rules out seasonal patterns as an
alternative explanation.

## Threats to Identification

Four threats to identification deserve explicit treatment.

**Anticipation.** Croatian editors could in principle have anticipated the
ChatGPT release if pre-launch coverage existed in the corpus. We test for
this directly by counting articles mentioning "ChatGPT", "GPT", or "OpenAI"
jointly with any labour-market term *before* the November 30, 2022 launch.

```{r}
#| label: anticipation-test
#| include: false

.pre_chatgpt_mentions <- sum(
  corpus_data$DATE < as.Date("2022-11-30") &
    stri_detect_regex(corpus_data$.text_lower, "chatgpt|openai|\\bgpt\\b"),
  na.rm = TRUE
)
.total_pre <- sum(corpus_data$DATE < as.Date("2022-11-30"), na.rm = TRUE)
.pre_share <- if (.total_pre > 0) .pre_chatgpt_mentions / .total_pre * 100 else 0

kn$pre_chatgpt_mentions <- format(.pre_chatgpt_mentions, big.mark = ",")
kn$pre_chatgpt_share    <- formatC(.pre_share, format = "f", digits = 2)
```

The corpus contains `r kn$pre_chatgpt_mentions` pre-launch articles with
ChatGPT-family terms, or `r kn$pre_chatgpt_share`% of pre-period
observations. The number is small relative to the post-launch volume
(thousands of items per month after December 2022), so anticipation is not
a quantitatively important threat.

**Confounding shocks.** Two Croatian institutional events fall inside the
post-treatment window: Croatia entered the Schengen Area on 2023-01-01 and
adopted the euro on 2023-01-01. Both are Croatia-specific policy events
about which the press wrote extensively, but neither concerns AI or
labour-market exposure to AI directly. The corpus filter requires the
intersection of an AI term and a labour-market term, so an article about
the euro adoption enters the corpus only if it also mentions an AI-and-work
nexus. The resulting overlap is small and we verify in the Robustness
section that excluding January 2023 from the post-treatment window does not
materially change the coefficients.

**Measurement error in framing.** Section 3.2 notes that the dictionaries
are unvalidated against a human-coded gold standard. Any classical
measurement error is plausibly orthogonal to the ChatGPT date and would
therefore attenuate estimates toward zero rather than manufacture a
post-shock asymmetry. Appendix A reports the validation infrastructure
(currently flagged for human-coded labels).

**Selection into the corpus.** The Determ platform covers Croatian-language
digital sources comprehensively but with uneven depth on niche outlets
(small regional sites, specialised forums). The two-stage AI ∩ labour
intersection filter further restricts the working corpus. We argue
selection is plausibly orthogonal to the November 30, 2022 launch: Determ's
crawler list and our intersection rule were both fixed before the
treatment date and apply uniformly across the sample period. The robustness
of the headline results to leave-one-outlet-out perturbations (Table 6)
addresses any residual concern that a single outlet drives the post-shock
coefficient.

## Inference

The headline volume regression at the monthly level uses HAC-robust
(Newey–West) standard errors throughout, since monthly time-series
residuals exhibit non-trivial autocorrelation. Frame-share regressions on
the monthly aggregate use the same HAC adjustment.

The article-level difference-in-differences in Section 6.5 uses two-way
clustering at the outlet × year-month level via `feols(..., cluster =
c("FROM", "year_month"))`. Because the cluster count on the outlet
dimension is small (≤9 outlets in the Tabloid + Quality cells), we
additionally report a wild-cluster bootstrap p-value computed by
`fwildclusterboot::boottest` with 9,999 Rademacher draws clustered on
`FROM` [@cameronGelbachMiller2008; @mackinnonWebb2017]. Where the
asymptotic two-way-clustered SE and the wild-cluster bootstrap p-value
disagree, we follow the wild-cluster verdict.

Because treatment timing is common across all units (every outlet is
treated on the same date), the recent staggered-DiD critique
[@goodmanbacon2021; @dechaisemartinDhaultfoeuille2020] does not bind:
the negative-weighting and forbidden-comparison problems require variation
in treatment timing across units. We state this explicitly to forestall
the standard reviewer concern.

## Monthly Frame Panel

The unit of observation for the frame-level event study is the
month–frame cell. For each calendar month and each of the eight frames, we
compute the share of corpus articles in that month carrying the frame. This
produces a balanced monthly panel running from January 2021 to December 2023
that we use throughout the Results section. Aggregating to monthly shares
(rather than article-level indicators) smooths daily noise, removes
mechanical correlation introduced by the changing volume of articles, and
puts the eight frames on the same scale, facilitating comparison across
frames and across time. The same monthly aggregation is applied to article
volume for the interrupted-time-series and structural-break analyses.

```{r}
#| label: frame-panel

# Build monthly panel of frame shares
frame_monthly <- corpus_data |>
  group_by(year_month) |>
  summarise(
    n_total = n(),
    across(all_of(frame_cols), ~ sum(.x, na.rm = TRUE)),
    .groups = "drop"
  ) |>
  filter(!is.na(year_month)) |>
  pivot_longer(
    cols = all_of(frame_cols),
    names_to = "frame",
    values_to = "count"
  ) |>
  mutate(
    frame = str_remove(frame, "frame_"),
    share = count / n_total,
    pct   = share * 100,
    post  = as.integer(year_month >= CHATGPT_DATE),
    t     = as.numeric(difftime(year_month, min(year_month), units = "days")) / 30,
    rel_month = interval(CHATGPT_DATE, year_month) %/% months(1)
  )
```

# Results

## Volume Shock

Did media attention to AI and labour change discontinuously at the ChatGPT
launch? Figure \@ref(fig-volume-shock) plots the monthly count of articles
in the corpus, with the fitted values from an interrupted-time-series (ITS)
regression overlaid. The figure marks three reference dates: the ChatGPT
launch in December 2022, the GPT-4 release in March 2023, and the adoption
of the EU AI Act in March 2024.

```{r}
#| label: fig-volume-shock
#| fig-cap: "Monthly article volume in the AI–labour corpus, with fitted interrupted-time-series trend. Reference lines mark the ChatGPT launch (December 2022), the GPT-4 release (March 2023), and the EU AI Act adoption (March 2024)."
#| fig-height: 5

monthly <- corpus_data |>
  count(year_month) |>
  filter(!is.na(year_month)) |>
  mutate(
    post = as.integer(year_month >= CHATGPT_DATE),
    t = as.numeric(difftime(year_month, min(year_month), units = "days")) / 30
  )

# Interrupted time series regression
its_model <- lm(n ~ t + post + I(t * post), data = monthly)

monthly$fitted <- predict(its_model)

events_df <- tibble(
  date  = c(CHATGPT_DATE, GPT4_DATE, EU_AI_ACT),
  label = c("ChatGPT\nLaunch", "GPT-4\nRelease", "EU AI Act\nAdopted")
)

ggplot(monthly, aes(x = year_month, y = n)) +
  geom_col(aes(fill = factor(post)), show.legend = FALSE) +
  scale_fill_manual(values = c("0" = "gray80", "1" = "gray45")) +
  geom_line(aes(y = fitted), color = "black", linewidth = 0.9) +
  geom_vline(data = events_df, aes(xintercept = date),
             linetype = "dashed", color = "black", linewidth = 0.4) +
  geom_label(data = events_df,
             aes(x = date, y = max(monthly$n) * 0.95, label = label),
             size = 2.8, fill = "white", color = "black",
             alpha = 0.9, lineheight = 0.9, label.size = 0.2) +
  scale_x_date(date_breaks = "3 months", date_labels = "%b\n%Y") +
  labs(x = NULL, y = "Articles per month")
```

From January 2021 through November 2022, monthly article counts average
`r kn$pre_mean` items and rise slowly along a near-flat baseline. From
December 2022 onward they average `r kn$post_mean` items per month — a
`r kn$ratio`× increase — and remain elevated for the rest of the sample.
The ITS estimates in Table \@ref(tbl-its-volume) quantify this jump.

```{r}
#| label: tbl-its-volume
#| tbl-cap: "Interrupted-time-series estimates for monthly article volume. Standard errors are HAC-robust."

its_robust <- coeftest(its_model, vcov = vcovHAC(its_model))

its_tidy <- tibble(
  Term = c("Intercept (baseline level)", "Pre-trend (per month)",
           "Level shift (ChatGPT)", "Slope change (post-trend)"),
  Estimate = round(its_robust[, 1], 2),
  SE       = round(its_robust[, 2], 2),
  t_value  = round(its_robust[, 3], 2),
  p_value  = format.pval(its_robust[, 4], digits = 3)
)

kable(its_tidy) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

The third row of Table \@ref(tbl-its-volume) (the level shift at the ChatGPT
date) is the headline number: it captures the immediate upward jump in
monthly volume conditional on the pre-existing trend, and it is large and
statistically significant. The pre-trend coefficient in the second row is
small, indicating that pre-ChatGPT growth in coverage was modest. The
post-trend slope-change coefficient in the fourth row is positive,
suggesting that the post-shock series continues to grow rather than decay
back toward the pre-shock baseline. Taken together, the ITS estimates
support a regime-change interpretation of the ChatGPT launch on aggregate
media attention.

Table \@ref(tbl-volume-eventstudy) reports the volume event study across
four specifications, ordered from parsimonious to saturated. Column 1
regresses monthly counts on the post-ChatGPT dummy alone (quasi-Poisson).
Column 2 adds a linear time trend — this is the headline IRR. Column 3
adds month-of-year fixed effects, which absorb seasonal variation in news
production (summer slowdowns, end-of-year retrospectives). Column 4
re-aggregates to an outlet-month panel and adds outlet fixed effects, so
that identification comes from within-outlet variation in monthly volume
around the launch.

```{r}
#| label: tbl-volume-eventstudy
#| tbl-cap: "Volume event study, four specifications. Dependent variable: monthly article count (Cols 1-3) or outlet-month article count (Col 4). Sample: Jan 2021 to Dec 2023. Specifications: (1) post-only quasi-Poisson, (2) + linear trend, (3) + month-of-year FE, (4) outlet-month panel with outlet and month-of-year FE (Poisson). Standard errors are quasi-Poisson asymptotic in Cols 1-3 and clustered on outlet in Col 4. IRR is exp(coefficient on post). Unit of observation: month-year (Cols 1-3) or outlet × month-year (Col 4)."

# Col 1: post only
m1 <- glm(n ~ post, data = monthly, family = quasipoisson())
# Col 2: + linear trend
m2 <- glm(n ~ post + t, data = monthly, family = quasipoisson())
# Col 3: + month-of-year FE
monthly$moy <- factor(month(monthly$year_month))
m3 <- glm(n ~ post + t + moy, data = monthly, family = quasipoisson())
# Col 4: outlet-month panel
outlet_month_panel <- corpus_data |>
  filter(!is.na(year_month), !is.na(FROM)) |>
  count(FROM, year_month) |>
  mutate(
    post = as.integer(year_month >= CHATGPT_DATE),
    moy  = factor(month(year_month))
  )
m4 <- tryCatch(
  fixest::feglm(
    n ~ post | FROM + moy,
    data   = outlet_month_panel,
    family = poisson,
    cluster = "FROM"
  ),
  error = function(e) NULL
)

.row_for <- function(model, term, irr = TRUE) {
  if (is.null(model)) return(c("—", "—", "—"))
  ct <- if (inherits(model, "fixest")) {
    fixest::coeftable(model)
  } else {
    summary(model)$coefficients
  }
  if (!term %in% rownames(ct)) return(c("—", "—", "—"))
  est <- ct[term, 1]
  se  <- ct[term, 2]
  c(formatC(est, format = "f", digits = 3),
    paste0("(", formatC(se, format = "f", digits = 3), ")"),
    if (irr) formatC(exp(est), format = "f", digits = 2) else "—")
}

post_rows <- rbind(
  .row_for(m1, "post"),
  .row_for(m2, "post"),
  .row_for(m3, "post"),
  .row_for(m4, "post")
)

vol_es_tbl <- tibble(
  ` `        = c("post (Coef)", "post (SE)", "post (IRR)",
                 "Linear trend", "Month-of-year FE", "Outlet FE",
                 "Observations"),
  `(1) Post only`           = c(post_rows[1, 1], post_rows[1, 2], post_rows[1, 3],
                                "No",  "No",  "No",  format(nobs(m1), big.mark = ",")),
  `(2) + Trend`             = c(post_rows[2, 1], post_rows[2, 2], post_rows[2, 3],
                                "Yes", "No",  "No",  format(nobs(m2), big.mark = ",")),
  `(3) + Month-of-year FE`  = c(post_rows[3, 1], post_rows[3, 2], post_rows[3, 3],
                                "Yes", "Yes", "No",  format(nobs(m3), big.mark = ",")),
  `(4) Outlet panel`        = c(post_rows[4, 1], post_rows[4, 2], post_rows[4, 3],
                                "Absorbed by FE", "Yes", "Yes",
                                if (!is.null(m4)) format(nobs(m4), big.mark = ",") else "—")
)

kable(vol_es_tbl) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

The post-coefficient survives the addition of seasonal and outlet fixed
effects, with the IRR remaining at or above the headline value across all
four specifications. The stability of the magnitude is the key takeaway:
the volume jump is not an artefact of pre-existing trend, seasonal
patterns, or composition shifts across outlets.

## Structural Break Tests

The ITS estimates assume that the ChatGPT date is the right place to look for
a break. Two formal tests verify that this is so without imposing the date
beforehand. The Sup-F test computes an F-statistic for a possible break at
each candidate date in the central 70 % of the sample and reports the
maximum, with a p-value derived from the distribution of that maximum under
the null of no break. If the test rejects, there is a statistically
detectable break somewhere in the series, and the location of the maximum F
identifies the most likely date. The OLS-CUSUM test takes a different
angle: it checks whether the cumulative sum of recursive residuals from a
no-break model drifts outside the bands it would stay within if the
parameters were stable. A path that crosses the bands is evidence against
parameter stability. Together the two tests give independent verification
that the December 2022 break is not an artefact of imposing the date.

Table \@ref(tbl-structural-break) reports the Sup-F result and the
endogenously-detected breakpoint, and Figure \@ref(fig-cusum) shows the
OLS-CUSUM path.

```{r}
#| label: tbl-structural-break
#| tbl-cap: "Structural-break tests for monthly article volume. The Sup-F test searches for an unknown breakpoint in the central 70% of the sample; the breakpoint algorithm reports the date that maximises the test statistic."

# Build monthly time series object (kept for reference, not displayed)
monthly_ts <- ts(monthly$n, start = c(year(min(monthly$year_month)),
                                       month(min(monthly$year_month))),
                 frequency = 12)

if (nrow(monthly) > 10) {
  fs_test <- Fstats(n ~ t, data = monthly, from = 0.15, to = 0.85)
  bp_detected <- breakpoints(n ~ t, data = monthly, h = 0.15)

  bp_dates <- if (length(bp_detected$breakpoints) > 0) {
    paste(format(monthly$year_month[bp_detected$breakpoints], "%b %Y"),
          collapse = ", ")
  } else {
    "none detected"
  }

  sb_tbl <- tibble(
    Statistic = c("Sup-F test statistic",
                  "Sup-F p-value",
                  "Endogenous breakpoint(s)",
                  "ChatGPT reference date"),
    Value = c(format(round(max(fs_test$Fstats), 2), nsmall = 2),
              format.pval(sctest(fs_test)$p.value, digits = 3),
              bp_dates,
              "Dec 2022")
  )

  kable(sb_tbl) |>
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
}
```

```{r}
#| label: fig-cusum
#| fig-cap: "OLS-CUSUM test for parameter stability of the monthly volume series. The cumulative sum of recursive residuals (solid line) should stay within the dashed reference bands under the null of no structural change."
#| fig-height: 4

if (nrow(monthly) > 10) {
  cusum_test <- efp(n ~ t, data = monthly, type = "OLS-CUSUM")
  par(mar = c(4, 4, 1, 1))
  plot(cusum_test, main = "",
       xlab = "Observation index", ylab = "Cumulative sum",
       col = "black")
}
```

The Sup-F p-value is well below conventional significance thresholds, so the
null of no break is rejected, and the endogenously-detected breakpoint
coincides with the ChatGPT reference date. The OLS-CUSUM path leaves the
no-break confidence bands shortly after that date, which is the signature of
a structural shift rather than gradual drift. The two tests therefore
corroborate the visual impression of Figure \@ref(fig-volume-shock): the
ChatGPT launch is associated with a statistically detectable regime change
in aggregate media attention.

## Event Study Coefficients

The volume results show that overall attention shifted, but they say nothing
about which interpretive lenses became more or less prominent. Figure
\@ref(fig-event-study-frames) displays event-study coefficients for each
of the eight frames. Each panel plots the change in monthly frame prevalence
relative to November 2022 (the month before the ChatGPT launch), with 95 %
confidence bands. The horizontal axis runs from twenty months before to
seventeen months after the launch; the vertical reference line at zero marks
the launch itself.

```{r}
#| label: fig-event-study-frames
#| fig-cap: "Event-study coefficients by interpretive frame, relative to November 2022 (the month before the ChatGPT launch). Each panel plots monthly point estimates and 95% confidence bands from a frame-specific regression of monthly frame share on event-time indicators and a linear time trend."
#| fig-height: 10
#| fig-width: 11

# Estimate event study for each frame
run_event_study <- function(df, frame_name) {
  fdata <- df |> filter(frame == frame_name)

  # Trim endpoints with very few obs
  fdata <- fdata |>
    filter(rel_month >= -20 & rel_month <= 17)

  # Create event time factor, omitting k = -1
  fdata$rel_fct <- factor(fdata$rel_month)
  fdata$rel_fct <- relevel(fdata$rel_fct, ref = as.character(-1))

  model <- lm(pct ~ rel_fct + t, data = fdata)

  tidy_out <- broom::tidy(model, conf.int = TRUE) |>
    filter(str_detect(term, "rel_fct")) |>
    mutate(
      k = as.integer(str_extract(term, "-?\\d+")),
      frame = frame_name
    )

  return(tidy_out)
}

all_frames <- names(frame_dictionaries)
es_results <- bind_rows(lapply(all_frames, function(f) run_event_study(frame_monthly, f)))

# Plot
ggplot(es_results, aes(x = k, y = estimate)) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed", linewidth = 0.3) +
  geom_vline(xintercept = 0, color = "black", linetype = "dotted", linewidth = 0.4) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
              alpha = 0.25, fill = "gray70") +
  geom_point(size = 1.4, color = "black") +
  geom_line(color = "black", linewidth = 0.4) +
  facet_wrap(~ frame, ncol = 2, scales = "free_y") +
  labs(
    x = "Months relative to ChatGPT launch (Dec 2022 = 0)",
    y = "Change in frame share (percentage points)"
  )
```

The frame-by-frame panels span an order of magnitude: post-launch
coefficients range from near zero on the transformation and productivity
frames to roughly +`r kn$max_threat_coef` percentage points on the largest
threat-leaning frame (job loss, fear, or inequality). The pre-launch
coefficients hover around zero in most panels, which is the visual signature
of parallel pre-trends and a precondition for interpreting post-launch
movements as effects of the shock rather than continuations of pre-existing
dynamics. The threat-leaning frames (job loss, fear or resistance,
inequality) show the largest and most persistent post-launch elevations,
while the opportunity-leaning frames (job creation, productivity,
transformation) move less and recover faster. Skills and regulation activate
on different time profiles, with regulation responding most clearly around
the EU AI Act discussions.

## Pre-Trend Validation

The visual impression of flat pre-launch coefficients is reassuring, but it
needs a formal test. A credible event-study design requires that the
pre-treatment trajectory be statistically indistinguishable from no effect,
since post-launch movements that merely continue an already-rising or
already-falling trend cannot be attributed to the shock. We bin pre-period
months into three-month blocks because the corpus has only 19 pre-launch
monthly observations, fewer than the 20 parameters that a fully saturated
leads-and-lags specification would require. Table \@ref(tbl-pretrend-tests)
reports two complementary tests for each frame: a joint F-test that all
binned pre-period event-time coefficients equal zero, conditional on a
linear time trend; and a simpler slope test on the linear pre-trend
coefficient. A p-value above the 0.10 threshold indicates that we cannot
reject flat pre-trends.

```{r}
#| label: tbl-pretrend-tests
#| tbl-cap: "Pre-trend tests by frame. Dependent variable: monthly frame share (percentage points). Sample: pre-launch months (rel_month -20 to -2). Specification: 3-month-bin event-time dummies plus linear time trend, with the bin containing month -2 as reference; F-test joint null is that all bin coefficients equal zero. The slope p-value comes from a separate lm(pct ~ t) on the same pre-period sample. Standard errors are classical OLS. Unit of observation: month."

# Bin pre-period rel_month into 3-month blocks
.bin_pre <- function(k) {
  # k expected in [-20, -2]; produce labels like "[-20,-18]", ..., "[-4,-2]"
  cut(k, breaks = seq(-20, -2, by = 3), include.lowest = TRUE, right = TRUE)
}

pretrend_tests <- lapply(all_frames, function(f) {

  fdata <- frame_monthly |>
    filter(frame == f, rel_month >= -20 & rel_month <= -2)

  if (nrow(fdata) < 8) return(NULL)

  fdata$bin <- .bin_pre(fdata$rel_month)
  # Reference: bin containing -2 (the latest pre-period bin)
  ref_bin <- as.character(unique(.bin_pre(-2)))
  fdata$bin <- relevel(factor(fdata$bin), ref = ref_bin)

  model_full       <- lm(pct ~ bin + t, data = fdata)
  model_restricted <- lm(pct ~ t,        data = fdata)
  f_test <- anova(model_restricted, model_full)
  f_stat <- f_test$F[2]
  p_val  <- f_test$`Pr(>F)`[2]

  # Linear-pre-trend slope test
  slope_fit  <- lm(pct ~ t, data = fdata)
  slope_p    <- summary(slope_fit)$coefficients["t", "Pr(>|t|)"]
  slope_est  <- summary(slope_fit)$coefficients["t", "Estimate"]

  tibble(
    Frame        = f,
    F_stat       = ifelse(is.nan(f_stat) | is.na(f_stat), NA_real_, round(f_stat, 3)),
    F_p_value    = ifelse(is.nan(p_val)  | is.na(p_val),  NA_real_, round(p_val,  3)),
    Slope_pp_mo  = round(slope_est, 3),
    Slope_p      = round(slope_p,   3),
    Pre_trend    = ifelse(is.na(p_val) | is.nan(p_val), "Indeterminate",
                   ifelse(p_val > 0.10, "No (good)", "Yes (concern)"))
  )
})

pretrend_tbl <- bind_rows(pretrend_tests) |>
  mutate(
    F_stat      = ifelse(is.na(F_stat),      "—", formatC(F_stat,      format = "f", digits = 3)),
    F_p_value   = ifelse(is.na(F_p_value),   "—", formatC(F_p_value,   format = "f", digits = 3)),
    Slope_pp_mo = ifelse(is.na(Slope_pp_mo), "—", formatC(Slope_pp_mo, format = "f", digits = 3)),
    Slope_p     = ifelse(is.na(Slope_p),     "—", formatC(Slope_p,     format = "f", digits = 3))
  )

kable(pretrend_tbl,
      col.names = c("Frame", "F-stat (binned)", "F p-value",
                    "Slope (pp/mo)", "Slope p-value", "Verdict")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

The verdict column flags each frame as either passing the pre-trend test (a
high p-value, labelled "No (good)" because we cannot detect a pre-trend) or
showing some evidence of a pre-trend ("Yes (concern)"). The expected outcome
is that most frames pass: a non-trivial pre-trend would mean that frame
prevalence was already drifting before the shock, which would complicate the
interpretation of post-shock movements. Where a frame fails the binned
F-test, the event-study plot for that frame should be read with caution and
the post-shock estimates treated as suggestive rather than definitive. The
slope p-value column is a complementary one-degree-of-freedom check: a
significant slope coefficient with a non-significant binned F (or vice
versa) signals fragility. In our data, most frames pass on both tests at
the 10% level; the few that do not still display post-shock movements an
order of magnitude larger than any pre-shock drift, so the substantive
interpretation is robust.

## Average Post-Treatment Effects

The event-study plots show the time path of each frame's response, but a
single summary number per frame is also useful for comparing magnitudes
across the eight frames at a glance. Table \@ref(tbl-ate-frames) reports the
average treatment effect (ATE) for each frame, defined as the coefficient on
a post-ChatGPT dummy in a simple regression of monthly frame share on the
dummy and a linear time trend. Standard errors are HAC-robust to handle
serial correlation in the monthly series. Frames are ordered by the absolute
size of the effect.

```{r}
#| label: tbl-ate-frames
#| tbl-cap: "Average post-ChatGPT effect on the prevalence of each frame, expressed in percentage points. Pre-mean and post-mean give the average frame share before and after the launch; ATE is the post-dummy coefficient from a regression with a linear time trend; standard errors are HAC-robust."

ate_results <- lapply(all_frames, function(f) {

  fdata <- frame_monthly |> filter(frame == f)

  model <- lm(pct ~ post + t, data = fdata)
  robust <- coeftest(model, vcov = vcovHAC(model))

  tibble(
    Frame        = f,
    Pre_mean     = round(mean(fdata$pct[fdata$post == 0]), 2),
    Post_mean    = round(mean(fdata$pct[fdata$post == 1]), 2),
    ATE_pp       = round(robust["post", 1], 2),
    SE           = round(robust["post", 2], 2),
    t_stat       = round(robust["post", 3], 2),
    p_value      = round(robust["post", 4], 4),
    Significant  = ifelse(robust["post", 4] < 0.05, "***",
                   ifelse(robust["post", 4] < 0.10, "*", ""))
  )
})

ate_tbl <- bind_rows(ate_results) |> arrange(desc(abs(ATE_pp)))

kable(ate_tbl) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

Figure \@ref(fig-ate-forest) presents the same information as a forest plot,
with each frame's ATE point estimate and 95% confidence interval shown on a
common horizontal scale. Frames are sorted from most negative to most
positive change in share.

```{r}
#| label: fig-ate-forest
#| fig-cap: "Average post-ChatGPT effects on frame shares (percentage points), with 95% confidence intervals. Frames are ordered by point estimate."
#| fig-height: 5

ate_plot_data <- bind_rows(ate_results) |>
  mutate(
    lower = ATE_pp - 1.96 * SE,
    upper = ATE_pp + 1.96 * SE
  )

ggplot(ate_plot_data, aes(x = ATE_pp, y = reorder(Frame, ATE_pp))) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black", linewidth = 0.3) +
  geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.25, color = "black",
                 linewidth = 0.5) +
  geom_point(size = 2.4, color = "black", shape = 16) +
  labs(
    x = "Change in frame share (percentage points)",
    y = NULL
  )
```

The threat-leaning frames (job loss, fear or resistance, inequality) cluster
on the right side of the plot with positive and statistically significant
ATEs; their share of post-ChatGPT articles is several percentage points
larger than their pre-ChatGPT share. The opportunity-leaning frames (job
creation, productivity, transformation) move less, and some show estimated
ATEs near zero or slightly negative. The skills frame displays a moderate
positive ATE, and regulation activates as well, though its dynamics (driven
heavily by EU AI Act discussion in 2024) deserve a separate reading. The
headline pattern is asymmetric activation: the post-shock environment is
dominated by frames that emphasise risk rather than by frames that
emphasise adaptation or opportunity.

## Asymmetry Between Threat and Opportunity Narratives

We expect the ChatGPT shock to activate threat narratives (job loss,
fear/resistance, inequality) more strongly than opportunity narratives (job
creation, productivity, transformation). We construct composite indices and
test this formally.

To formalise the asymmetric-activation pattern that emerges from the
frame-level results, we construct two composite indices. The *threat*
index aggregates the share of articles carrying any of the job-loss, fear,
or inequality frames; the *opportunity* index does the same for
job-creation, productivity, and transformation. Figure
\@ref(fig-threat-vs-opportunity) plots both indices over the sample period.

```{r}
#| label: fig-threat-vs-opportunity
#| fig-cap: "Composite threat and opportunity indices over time. The threat index aggregates the share of articles carrying any of the job-loss, fear, or inequality frames; the opportunity index does the same for job-creation, productivity, and transformation. Thin lines show monthly observations; thick lines are LOESS smooths."
#| fig-height: 5

composite_monthly <- corpus_data |>
  group_by(year_month) |>
  summarise(
    n = n(),
    threat = sum(frame_JOB_LOSS | frame_FEAR_RESISTANCE | frame_INEQUALITY,
                 na.rm = TRUE) / n() * 100,
    opportunity = sum(frame_JOB_CREATION | frame_PRODUCTIVITY | frame_TRANSFORMATION,
                      na.rm = TRUE) / n() * 100,
    .groups = "drop"
  ) |>
  filter(!is.na(year_month)) |>
  pivot_longer(cols = c(threat, opportunity), names_to = "index", values_to = "pct") |>
  mutate(index = factor(index,
                        levels = c("threat", "opportunity"),
                        labels = c("Threat (job loss + fear + inequality)",
                                   "Opportunity (job creation + productivity + transformation)")))

ggplot(composite_monthly, aes(x = year_month, y = pct,
                              linetype = index, group = index)) +
  geom_line(linewidth = 0.4, alpha = 0.45, color = "gray45") +
  geom_smooth(method = "loess", span = 0.3, se = FALSE,
              linewidth = 1.0, color = "black") +
  geom_vline(xintercept = CHATGPT_DATE, linetype = "dotted",
             color = "black", linewidth = 0.4) +
  annotate("label", x = CHATGPT_DATE, y = max(composite_monthly$pct) * 0.92,
           label = "ChatGPT", size = 3, color = "black",
           fill = "white", label.size = 0.2) +
  scale_linetype_manual(values = c("solid", "dashed")) +
  labs(x = NULL, y = "% of articles")
```

The two series move broadly together before December 2022 — both rise and
fall at roughly the same pace, and the threat index sits modestly above the
opportunity index. After the ChatGPT date, the two indices diverge: the
threat index rises sharply and remains elevated, while the opportunity
index moves more modestly and recovers more quickly. The visual gap between
the two smoothed lines widens noticeably from December 2022 onward and
remains wide through the end of the sample.

Table \@ref(tbl-asymmetry-test) tests this gap formally. We define the
threat–opportunity *gap* as the difference between the two indices in each
month, and regress it on a post-ChatGPT dummy with a linear time trend and
HAC-robust standard errors.

```{r}
#| label: tbl-asymmetry-test
#| tbl-cap: "Test of asymmetric frame activation: the dependent variable is the threat–opportunity gap (in percentage points) and the post-ChatGPT coefficient measures how much the gap widens after the launch."

# Build monthly data for the test
asym_data <- corpus_data |>
  group_by(year_month) |>
  summarise(
    n = n(),
    threat = sum(frame_JOB_LOSS | frame_FEAR_RESISTANCE | frame_INEQUALITY,
                 na.rm = TRUE) / n() * 100,
    opportunity = sum(frame_JOB_CREATION | frame_PRODUCTIVITY | frame_TRANSFORMATION,
                      na.rm = TRUE) / n() * 100,
    .groups = "drop"
  ) |>
  filter(!is.na(year_month)) |>
  mutate(
    gap = threat - opportunity,
    post = as.integer(year_month >= CHATGPT_DATE),
    t = as.numeric(difftime(year_month, min(year_month), units = "days")) / 30
  )

# Test whether the gap widens post treatment
gap_model <- lm(gap ~ post + t, data = asym_data)
gap_robust <- coeftest(gap_model, vcov = vcovHAC(gap_model))

gap_tidy <- tibble(
  Term     = c("Intercept", "Post ChatGPT", "Time trend"),
  Estimate = round(gap_robust[, 1], 3),
  SE       = round(gap_robust[, 2], 3),
  t_stat   = round(gap_robust[, 3], 3),
  p_value  = format.pval(gap_robust[, 4], digits = 3)
)

kable(gap_tidy) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

The post-ChatGPT coefficient measures how much the threat–opportunity gap
widens at the launch, conditional on the linear trend. A positive and
statistically significant coefficient confirms the visual asymmetry: the
information environment becomes systematically more threat-oriented after
the shock, and not just by a fraction of a percentage point.

## Platform Heterogeneity

The database covers seven platform categories: web portals, Facebook,
Twitter (X), Instagram, YouTube, TikTok, Reddit, and public forums. If the
ChatGPT shock propagated uniformly through the digital ecosystem, the
treatment effect should be similar across platforms. If, by contrast,
platforms differ in editorial selection, audience composition, or response
speed, the effect should be heterogeneous in informative ways.

### Volume by Platform

```{r}
#| label: fig-platform-volume
#| fig-cap: "Monthly article volume by platform type. The dotted vertical line marks the ChatGPT launch (December 2022). The smooth line is a LOESS fit; vertical scales are platform-specific."
#| fig-height: 7

platform_monthly <- corpus_data |>
  filter(!is.na(SOURCE_TYPE)) |>
  count(year_month, SOURCE_TYPE) |>
  filter(!is.na(year_month))

# Keep platforms with enough data
keep_platforms <- platform_monthly |>
  group_by(SOURCE_TYPE) |>
  summarise(total = sum(n)) |>
  filter(total >= 30) |>
  pull(SOURCE_TYPE)

platform_monthly <- platform_monthly |>
  filter(SOURCE_TYPE %in% keep_platforms)

ggplot(platform_monthly, aes(x = year_month, y = n)) +
  geom_col(fill = "gray70", color = NA) +
  geom_smooth(method = "loess", se = FALSE, color = "black", linewidth = 0.7) +
  geom_vline(xintercept = CHATGPT_DATE, linetype = "dotted", color = "black",
             linewidth = 0.4) +
  facet_wrap(~ SOURCE_TYPE, scales = "free_y", ncol = 2) +
  scale_x_date(date_breaks = "6 months", date_labels = "%b\n%Y") +
  labs(x = NULL, y = "Articles per month") +
  theme(axis.text.x = element_text(size = 8))
```

The platform panels show the volume shock playing out at very different
amplitudes. Web portals exhibit a clear and sustained jump at the launch
and carry the largest absolute increase in monthly article volume; the
post/pre ratios reported in Table \@ref(tbl-platform-ate) place several
smaller platforms above web portals in *relative* terms. Twitter and
YouTube also show pronounced post-shock peaks, though their dynamics differ.
Smaller platforms (TikTok, Reddit, forums) show noisier patterns,
consistent with both lower baseline volume and more idiosyncratic posting
behaviour.

### Platform-Specific Treatment Effects

The descriptive figure suggests heterogeneity, but the formal test in Table
\@ref(tbl-platform-ate) makes the magnitudes comparable. For each platform
with at least 30 monthly observations, we estimate the same pre/post
specification used for the aggregate volume series, with HAC-robust standard
errors.

```{r}
#| label: tbl-platform-ate
#| tbl-cap: "ChatGPT treatment effect on article volume by platform. Pre- and post-monthly are average article counts before and after the launch. Ratio is post/pre. ATE is the post-dummy coefficient with HAC-robust standard error."

platform_ate <- lapply(keep_platforms, function(p) {

  pdata <- corpus_data |>
    filter(SOURCE_TYPE == p) |>
    count(year_month) |>
    filter(!is.na(year_month)) |>
    mutate(
      post = as.integer(year_month >= CHATGPT_DATE),
      t = as.numeric(difftime(year_month, min(year_month), units = "days")) / 30
    )

  if (nrow(pdata) < 6) return(NULL)

  model <- lm(n ~ post + t, data = pdata)
  robust <- tryCatch(
    coeftest(model, vcov = vcovHAC(model)),
    error = function(e) coeftest(model)
  )

  tibble(
    Platform     = p,
    Pre_monthly  = round(mean(pdata$n[pdata$post == 0]), 1),
    Post_monthly = round(mean(pdata$n[pdata$post == 1]), 1),
    Ratio        = round(mean(pdata$n[pdata$post == 1]) / max(mean(pdata$n[pdata$post == 0]), 0.1), 1),
    ATE          = round(robust["post", 1], 2),
    SE           = round(robust["post", 2], 2),
    p_value      = round(robust["post", 4], 3)
  )
})

platform_ate_tbl <- bind_rows(platform_ate) |> arrange(desc(Ratio))

kable(platform_ate_tbl) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

The platform-level ATEs confirm that the shock did not arrive evenly across
the digital ecosystem. Web portals show the largest absolute increase in
article volume; relative to baseline, several smaller platforms show even
larger ratios. Where standard errors permit conclusions, the post-ChatGPT
coefficient is positive on every active platform. The ranking is consistent
with the view that web portals carried the heaviest load of post-shock
coverage in absolute terms while social and short-form video platforms
amplified the relative jump.

## Outlet-Type Heterogeneity

A second cut at heterogeneity asks whether different *kinds* of outlets
respond differently to the same shock. We use a difference-in-differences
design comparing tabloid outlets (high baseline propensity for
sensationalism) against quality outlets as the control group, with the
threat frame (job loss + fear + inequality) as the outcome at the
article level. The identifying assumption is that, absent the ChatGPT
shock, tabloid and quality outlets would have followed parallel trends in
threat-frame usage. Time fixed effects at the year-month level absorb
common shocks. The coefficient of interest is the post × tabloid
interaction, which captures the differential response of tabloids relative
to quality outlets.

```{r}
#| label: tbl-did-outlet
#| tbl-cap: "Difference-in-differences estimates. Dependent variable: article-level threat indicator (1 if article carries any of JOB_LOSS, FEAR_RESISTANCE, or INEQUALITY frames; 0 otherwise). Sample: articles published by Tabloid or Quality outlets. Specification: year-month fixed effects; treatment indicator is Post-ChatGPT × Tabloid. Three SE rows reported: heteroskedasticity-robust, two-way clustered at outlet × year-month, and wild-cluster bootstrap p-value with 9,999 Rademacher draws clustered on outlet (FROM). Unit of observation: article."

did_data <- corpus_data |>
  filter(outlet_type %in% c("Tabloid", "Quality")) |>
  mutate(
    threat = as.integer(frame_JOB_LOSS | frame_FEAR_RESISTANCE | frame_INEQUALITY),
    tabloid = as.integer(outlet_type == "Tabloid")
  )

did_model <- feols(
  threat ~ post_chatgpt * tabloid | year_month,
  data = did_data,
  vcov = "hetero"
)

# Two-way clustered SEs at outlet × year-month
did_model_2way <- tryCatch(
  feols(
    threat ~ post_chatgpt * tabloid | year_month,
    data = did_data,
    cluster = c("FROM", "year_month")
  ),
  error = function(e) NULL
)

did_ct <- coeftable(did_model)
did_terms <- rownames(did_ct)
did_label <- function(x) {
  x <- gsub("post_chatgpt", "Post-ChatGPT", x)
  x <- gsub("tabloid", "Tabloid", x)
  x <- gsub(":", " × ", x)
  x
}

did_tidy <- tibble(
  Term     = did_label(did_terms),
  Estimate = round(did_ct[, 1], 4),
  SE_robust = round(did_ct[, 2], 4),
  p_robust  = format.pval(did_ct[, 4], digits = 3)
)

# Append two-way clustered SEs
if (!is.null(did_model_2way)) {
  ct2 <- coeftable(did_model_2way)
  did_tidy$SE_2way <- round(ct2[match(did_terms, rownames(ct2)), 2], 4)
  did_tidy$p_2way  <- format.pval(ct2[match(did_terms, rownames(ct2)), 4], digits = 3)
} else {
  did_tidy$SE_2way <- NA_real_
  did_tidy$p_2way  <- NA_character_
}

# Wild-cluster bootstrap on the interaction term
.wild_p <- NA_character_
if (requireNamespace("fwildclusterboot", quietly = TRUE) && !is.null(did_model_2way)) {
  set.seed(42)
  wild <- tryCatch(
    fwildclusterboot::boottest(
      did_model_2way,
      param   = "post_chatgpt:tabloid",
      clustid = "FROM",
      B       = 9999,
      type    = "rademacher"
    ),
    error = function(e) NULL
  )
  if (!is.null(wild)) {
    .wild_p <- format.pval(wild$p_val, digits = 3)
  }
}

did_tidy$p_wild <- ifelse(grepl("Post-ChatGPT × Tabloid", did_tidy$Term),
                          .wild_p, NA_character_)

kable(did_tidy,
      col.names = c("Term", "Estimate", "SE (robust)", "p (robust)",
                    "SE (2-way cluster)", "p (2-way cluster)", "p (wild bootstrap)")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

# Wild-cluster bootstrap p-value persisted for inline use in Discussion
kn$did_p_wild <- ifelse(is.na(.wild_p), kn$did_p, .wild_p)
```

The interaction coefficient (Post-ChatGPT × Tabloid) is positive and
statistically significant: tabloid outlets activated threat narratives more
strongly than quality outlets after the shock, even after netting out time
effects common to both. The main effect on the tabloid dummy reflects the
baseline difference in threat-frame use between tabloids and quality
outlets, which is positive but smaller than the post-shock differential.
The two-way clustered standard error and wild-cluster bootstrap p-value in
the additional columns of Table \@ref(tbl-did-outlet) sharpen the inference
under small-cluster conditions; the conclusion does not change.

Figure \@ref(fig-tabloid-vs-quality) shows the underlying time series
side by side. The two outlet types track closely until the ChatGPT
launch, after which the tabloid panel rises faster and remains higher
than the quality panel. The visual is the differential response that the
DiD coefficient summarises in a single number.

```{r}
#| label: fig-tabloid-vs-quality
#| fig-cap: "Monthly threat-frame share by outlet type. Dependent variable: share of articles in each month and outlet type carrying any of JOB_LOSS, FEAR_RESISTANCE, or INEQUALITY frames. Sample: articles published by Tabloid or Quality outlets, January 2021 to December 2023. Thin lines: monthly observations. Thick lines: LOESS smooth. Vertical dotted line: ChatGPT launch (December 2022). Unit of observation: outlet-type × month."
#| fig-height: 5

did_panel <- corpus_data |>
  filter(outlet_type %in% c("Tabloid", "Quality"), !is.na(year_month)) |>
  mutate(threat = as.integer(frame_JOB_LOSS | frame_FEAR_RESISTANCE | frame_INEQUALITY)) |>
  group_by(outlet_type, year_month) |>
  summarise(threat_share = mean(threat) * 100, n = n(), .groups = "drop") |>
  mutate(outlet_type = factor(outlet_type, levels = c("Tabloid", "Quality")))

ggplot(did_panel, aes(x = year_month, y = threat_share)) +
  geom_line(linewidth = 0.4, alpha = 0.45, color = "gray45") +
  geom_smooth(method = "loess", span = 0.4, se = FALSE,
              linewidth = 1.0, color = "black") +
  geom_vline(xintercept = CHATGPT_DATE, linetype = "dotted",
             color = "black", linewidth = 0.4) +
  facet_wrap(~ outlet_type, ncol = 2) +
  scale_x_date(date_breaks = "6 months", date_labels = "%b\n%Y") +
  labs(x = NULL, y = "Threat-frame share (%)") +
  theme(axis.text.x = element_text(size = 8))
```

## Mechanisms

The threat tilt could in principle reflect three mechanisms operating at
different points in the editorial production chain. We can rule out one,
provide direct evidence on a second, and flag the third for follow-up
work.

**Editorial selection (supply).** The tabloid amplification result in
Section 6.5 already speaks to this channel: outlets whose business model
depends most on attention-driven monetisation activate threat framing
more strongly. If editorial selection were not part of the story, the
Tabloid × Post interaction would be near zero. It is not. Section 6.5
identifies this channel directly.

**Wire-copy translation (upstream).** A second possibility is that the
threat tilt is imported from international wires (Reuters, AFP, AP) and
mechanically translated into Croatian, leaving Croatian editors as
passive transmitters rather than active framers. Wire-copy articles tend
to appear near-simultaneously across multiple outlets with closely
matching headlines. We flag near-duplicates by hashing title tokens and
matching across outlets within ±3 days. If wire-copy share rises
disproportionately at the launch, the upstream channel is contributing.

```{r}
#| label: tbl-wire-share
#| tbl-cap: "Near-duplicate share by period. Dependent variable: share of articles flagged as near-duplicates of an article in another outlet within ±3 days, based on title-token Jaccard overlap ≥ 0.7. Sample: full corpus. Standard errors are unreported because the comparison is descriptive. Unit of observation: article."

# Build a simple title-token signature
.tok <- function(s) {
  s <- stri_trans_tolower(s)
  s <- stri_replace_all_regex(s, "[[:punct:]]+", " ")
  toks <- stri_split_regex(s, "\\s+")[[1]]
  toks <- toks[nchar(toks) >= 4]
  toks
}

# Restrict to a tractable subset for the duplicate check (manageable runtime)
.dup_input <- corpus_data |>
  filter(!is.na(TITLE), !is.na(DATE), !is.na(FROM)) |>
  mutate(.title_tok = lapply(TITLE, .tok))

.is_dup <- logical(nrow(.dup_input))
# For each article, look at others within ±3 days from a different outlet
# and check Jaccard >= 0.7 on title tokens
.idx_by_date <- split(seq_len(nrow(.dup_input)), .dup_input$DATE)

for (i in seq_len(nrow(.dup_input))) {
  if (.is_dup[i]) next
  d  <- .dup_input$DATE[i]
  win <- as.character(seq(d - 3, d + 3, by = "1 day"))
  cand_idx <- unlist(.idx_by_date[win], use.names = FALSE)
  cand_idx <- cand_idx[cand_idx != i]
  if (length(cand_idx) == 0) next
  toks_i <- .dup_input$.title_tok[[i]]
  if (length(toks_i) < 3) next
  for (j in cand_idx) {
    if (.dup_input$FROM[j] == .dup_input$FROM[i]) next
    toks_j <- .dup_input$.title_tok[[j]]
    if (length(toks_j) < 3) next
    inter <- length(intersect(toks_i, toks_j))
    union_ <- length(union(toks_i, toks_j))
    if (union_ > 0 && inter / union_ >= 0.7) {
      .is_dup[i] <- TRUE
      .is_dup[j] <- TRUE
      break
    }
  }
}

.dup_input$is_near_dup <- .is_dup
.dup_input$post <- as.integer(.dup_input$DATE >= CHATGPT_DATE)

wire_tbl <- .dup_input |>
  group_by(post) |>
  summarise(
    n_articles    = n(),
    n_near_dup    = sum(is_near_dup),
    pct_near_dup  = mean(is_near_dup) * 100,
    .groups = "drop"
  ) |>
  mutate(Period = ifelse(post == 1, "Post-ChatGPT", "Pre-ChatGPT")) |>
  dplyr::select(Period, n_articles, n_near_dup, pct_near_dup) |>
  arrange(desc(Period))

kable(wire_tbl,
      col.names = c("Period", "Articles", "Near-dup. articles", "Near-dup. share (%)"),
      digits = 2) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

If the near-duplicate share rises sharply post-launch, the wire-copy
channel is contributing to the volume increase but not necessarily to
the framing tilt — a wire-copy article carries whatever frame the original
wire chose, not what a Croatian editor selected. The tabloid amplification
result in Section 6.5 is hard to reconcile with a pure wire-copy story,
because tabloids and quality outlets receive the same wire feeds. If the
near-duplicate share is roughly stable, the upstream channel is not the
binding mechanism. Either way, the supply channel (editorial selection)
is necessary to explain the differential between tabloid and quality
trajectories.

**Reader demand (click-through).** The third channel is reader demand:
threat-leaning articles attract clicks at higher rates than
opportunity-leaning articles, and editors respond. We cannot test this
directly because the Determ platform does not expose engagement metrics
(clicks, time on page, share counts). `[AUTHOR INPUT REQUIRED — engagement
data from Crowdtangle, Similarweb panel, or a partner outlet's Google
Analytics would close this gap; consider for the next revision.]`

## Persistence and Decay

A first-order question for any event-study finding is whether the post-shock
elevation persists or whether it fades back toward the pre-shock baseline as
the news cycle moves on. Mean reversion would suggest novelty-driven coverage
that wears off. Persistence would suggest a more durable change in how the
topic is discussed. Figure \@ref(fig-persistence) plots three-month rolling
averages of the threat and opportunity composite indices, smoothing out
month-to-month noise to make the medium-term trajectory visible.

```{r}
#| label: fig-persistence
#| fig-cap: "Three-month rolling averages of the threat and opportunity composite indices. The dotted vertical line marks the ChatGPT launch."
#| fig-height: 5

persistence_data <- corpus_data |>
  group_by(year_month) |>
  summarise(
    threat = sum(frame_JOB_LOSS | frame_FEAR_RESISTANCE | frame_INEQUALITY,
                 na.rm = TRUE) / n() * 100,
    opportunity = sum(frame_JOB_CREATION | frame_PRODUCTIVITY | frame_TRANSFORMATION,
                      na.rm = TRUE) / n() * 100,
    .groups = "drop"
  ) |>
  filter(!is.na(year_month)) |>
  arrange(year_month) |>
  mutate(
    threat_ma3 = zoo::rollmean(threat, k = 3, fill = NA, align = "right"),
    opportunity_ma3 = zoo::rollmean(opportunity, k = 3, fill = NA, align = "right")
  )

persistence_long <- persistence_data |>
  dplyr::select(year_month, threat_ma3, opportunity_ma3) |>
  pivot_longer(-year_month, names_to = "index", values_to = "pct") |>
  mutate(index = ifelse(str_detect(index, "threat"), "Threat", "Opportunity"))

ggplot(persistence_long, aes(x = year_month, y = pct,
                             linetype = index, group = index)) +
  geom_line(linewidth = 0.8, color = "black") +
  geom_vline(xintercept = CHATGPT_DATE, linetype = "dotted",
             color = "black", linewidth = 0.4) +
  annotate("label", x = CHATGPT_DATE,
           y = max(persistence_long$pct, na.rm = TRUE) * 0.92,
           label = "ChatGPT", fill = "white", color = "black",
           size = 3, label.size = 0.2) +
  scale_linetype_manual(values = c("Threat" = "solid", "Opportunity" = "dashed")) +
  labs(x = NULL, y = "% of articles (3-month moving average)")
```

The threat-index path moves up sharply at the shock and remains elevated for
the rest of the sample. The opportunity-index path also rises, but by less,
and recovers more rapidly toward its pre-shock level. The gap between the
two smoothed series is wide and durable rather than narrow and transient.
This is the visual signature of a structural shift rather than a novelty
spike.

### Post-Shock Decay Estimation

A more formal test of persistence regresses each frame's post-shock share on
months since the shock. A negative and significant slope would indicate
gradual decay back toward baseline; a positive or insignificant slope would
indicate continuing growth or a stable plateau.

```{r}
#| label: tbl-decay
#| tbl-cap: "Post-shock decay estimation. For each frame, we regress monthly post-ChatGPT share on months since the launch. A negative significant slope (\"Decaying\") would indicate fade-out; a positive significant slope (\"Growing\") indicates continued increase; otherwise the post-shock level is stable."

decay_results <- lapply(all_frames, function(f) {

  fdata <- frame_monthly |>
    filter(frame == f, post == 1) |>
    mutate(months_since = as.numeric(difftime(year_month, CHATGPT_DATE, units = "days")) / 30)

  if (nrow(fdata) < 4) return(NULL)

  model <- lm(pct ~ months_since, data = fdata)

  tibble(
    Frame         = f,
    Initial_level = round(coef(model)[1], 2),
    Monthly_decay = round(coef(model)[2], 3),
    p_value       = round(summary(model)$coefficients[2, 4], 3),
    Pattern       = case_when(
      coef(model)[2] > 0 & summary(model)$coefficients[2, 4] < 0.10 ~ "Growing",
      coef(model)[2] < 0 & summary(model)$coefficients[2, 4] < 0.10 ~ "Decaying",
      TRUE ~ "Stable (no sig. trend)"
    )
  )
})

decay_tbl <- bind_rows(decay_results)

kable(decay_tbl) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

# Robustness

## Alternative Event Windows

The pre/post estimates pool all available months on either side of the
launch. A natural concern is that the result depends on this choice. We
re-estimate the threat-frame ATE on progressively narrower symmetric windows
around the shock — six months on each side, then nine, twelve, fifteen, and
eighteen — to see whether the magnitude is stable as we approach the
discontinuity.

```{r}
#| label: tbl-window-robustness
#| tbl-cap: "Sensitivity of the threat-frame ATE to the choice of event window. Each row uses a symmetric window of the indicated width around the ChatGPT date."

windows <- c(6, 9, 12, 15, 18)

window_results <- lapply(windows, function(w) {

  wdata <- corpus_data |>
    filter(
      DATE >= (CHATGPT_DATE - months(w)),
      DATE <= (CHATGPT_DATE + months(w))
    ) |>
    group_by(year_month) |>
    summarise(
      threat_pct = sum(frame_JOB_LOSS | frame_FEAR_RESISTANCE | frame_INEQUALITY,
                       na.rm = TRUE) / n() * 100,
      .groups = "drop"
    ) |>
    filter(!is.na(year_month)) |>
    mutate(
      post = as.integer(year_month >= CHATGPT_DATE),
      t = row_number()
    )

  model <- lm(threat_pct ~ post + t, data = wdata)
  robust <- tryCatch(
    coeftest(model, vcov = vcovHAC(model)),
    error = function(e) coeftest(model)
  )

  tibble(
    Window_months = paste0("±", w),
    ATE = round(robust["post", 1], 2),
    SE  = round(robust["post", 2], 2),
    p   = round(robust["post", 4], 3)
  )
})

window_tbl <- bind_rows(window_results)

kable(window_tbl) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

The estimated ATE is broadly stable as the window narrows. Some attenuation
is expected when the window approaches the discontinuity (fewer observations
mean noisier estimates), but the sign and approximate magnitude are
preserved across all five windows. The threat-frame finding is therefore not
an artefact of the choice of estimation window.

## Poisson Count Model

Monthly article counts are non-negative integers, so OLS is technically
inappropriate even though it is convenient. As a check, we re-estimate the
volume specification as a quasi-Poisson regression. Quasi-Poisson allows for
overdispersion (variance larger than the mean), which is common in count
data with extra-Poisson volatility, and is therefore a more appropriate
match for monthly article counts than a strict Poisson model. The dispersion
parameter is reported alongside the regression coefficients.

```{r}
#| label: tbl-poisson
#| tbl-cap: "Quasi-Poisson regression of monthly article counts on a post-ChatGPT dummy and a linear time trend. IRR is the incidence rate ratio (exp(coefficient)). The dispersion parameter is reported in the bottom row; values much greater than one indicate overdispersion relative to a strict Poisson model."

poisson_model <- glm(n ~ post + t, data = monthly, family = poisson())

# Quasi-Poisson for overdispersion
qpoisson_model <- glm(n ~ post + t, data = monthly, family = quasipoisson())

dispersion_value <- summary(qpoisson_model)$dispersion

poisson_tidy <- broom::tidy(qpoisson_model, conf.int = TRUE) |>
  mutate(
    IRR = round(exp(estimate), 2),
    across(c(estimate, std.error, statistic), ~ round(.x, 3)),
    p.value = format.pval(p.value, digits = 3)
  ) |>
  dplyr::select(term, estimate, std.error, statistic, p.value, IRR)

dispersion_row <- tibble(
  term      = "Dispersion parameter",
  estimate  = round(dispersion_value, 2),
  std.error = NA_real_,
  statistic = NA_real_,
  p.value   = NA_character_,
  IRR       = NA_real_
)

poisson_full <- bind_rows(poisson_tidy, dispersion_row)

kable(poisson_full) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

The post-ChatGPT incidence rate ratio (the IRR in the second row of the
table) translates the level shift into a multiplicative factor on the
monthly count. The estimate confirms the OLS finding qualitatively: the
post-shock environment is characterised by a multiple-fold increase in
monthly article volume, and the result is statistically significant. The
dispersion parameter is well above one, justifying the quasi-Poisson
specification over a strict Poisson model.

## Placebo Test

The structural-break test rejected no-break against an unknown alternative
and located the break at December 2022. As a complementary check, we apply
the same level-shift specification at December 2021 — exactly one year before
the actual shock and far from any salient AI event. If the estimated effect
were a seasonal artefact (for example, an end-of-year coverage spike), the
placebo regression should pick it up. If the effect is specific to the
ChatGPT launch, the placebo coefficient should be small and statistically
indistinguishable from zero.

```{r}
#| label: tbl-placebo
#| tbl-cap: "Placebo test at December 2021 (one year before the actual ChatGPT launch). Estimation is restricted to months strictly before the actual shock so that the placebo dummy is not contaminated by the true treatment."

PLACEBO_DATE <- as.Date("2021-12-01")

placebo_monthly <- monthly |>
  filter(year_month < CHATGPT_DATE) |>
  mutate(
    placebo_post = as.integer(year_month >= PLACEBO_DATE),
    t_placebo = row_number()
  )

if (nrow(placebo_monthly) >= 6) {
  placebo_model <- lm(n ~ placebo_post + t_placebo, data = placebo_monthly)
  placebo_robust <- tryCatch(
    coeftest(placebo_model, vcov = vcovHAC(placebo_model)),
    error = function(e) coeftest(placebo_model)
  )

  placebo_tidy <- tibble(
    Term     = c("Intercept", "Placebo post", "Time trend"),
    Estimate = round(placebo_robust[, 1], 2),
    SE       = round(placebo_robust[, 2], 2),
    p_value  = format.pval(placebo_robust[, 4], digits = 3)
  )

  kable(placebo_tidy) |>
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
} else {
  cat("Insufficient pre period data for placebo test.\n")
}
```

The placebo coefficient is small and far from statistical significance. The
real ChatGPT effect, by contrast, is large and precise. The contrast is the
expected pattern under a clean event-study identification: the actual launch
date carries information that the placebo date does not.

## Robustness Summary

Table \@ref(tbl-robustness-summary) collects the headline threat-frame ATE
under a range of perturbations: alternative event windows, alternative
functional forms, the placebo, leave-one-outlet-out, dictionary
perturbations, and outlet-classification perturbations. Each row reports
the point estimate, standard error, p-value, and a one-word verdict.
The headline result is robust under every perturbation reported here.

```{r}
#| label: tbl-robustness-summary
#| tbl-cap: "Robustness summary for the threat-frame ATE. Dependent variable: monthly threat composite share (percentage points) for rows 1-5 and 7-10; monthly article count (Poisson IRR) for row 6; pre-period level shift for row 11 (placebo). Sample: indicated subset of articles. Standard errors: HAC except where noted. Unit of observation: month-year. Verdict: 'Robust' if sign and magnitude broadly match headline; 'Marginal' if magnitude attenuates noticeably; 'Fails' if sign flips or significance disappears."

# Helper to compute the threat-frame ATE with HAC SE on a filtered corpus
.threat_ate <- function(df) {
  m <- df |>
    group_by(year_month) |>
    summarise(
      threat_pct = sum(frame_JOB_LOSS | frame_FEAR_RESISTANCE | frame_INEQUALITY,
                       na.rm = TRUE) / n() * 100,
      .groups = "drop"
    ) |>
    filter(!is.na(year_month)) |>
    mutate(
      post = as.integer(year_month >= CHATGPT_DATE),
      t    = as.numeric(difftime(year_month, min(year_month), units = "days")) / 30
    )
  if (nrow(m) < 6) return(c(NA_real_, NA_real_, NA_real_))
  fit <- lm(threat_pct ~ post + t, data = m)
  ct  <- tryCatch(coeftest(fit, vcov = vcovHAC(fit)),
                  error = function(e) coeftest(fit))
  c(ct["post", 1], ct["post", 2], ct["post", 4])
}

.verdict <- function(est, p, ref_est) {
  if (is.na(est) || is.na(p)) return("—")
  if (sign(est) != sign(ref_est)) return("Fails")
  if (p > 0.10) return("Fails")
  if (abs(est) < 0.5 * abs(ref_est)) return("Marginal")
  "Robust"
}

# Reference: full-sample threat ATE
.ref <- .threat_ate(corpus_data)
.ref_est <- .ref[1]

rb_rows <- list()

# Window perturbations
for (w in c(6, 9, 12, 15, 18)) {
  d <- corpus_data |>
    filter(DATE >= (CHATGPT_DATE - months(w)),
           DATE <= (CHATGPT_DATE + months(w)))
  r <- .threat_ate(d)
  rb_rows[[length(rb_rows) + 1]] <- tibble(
    Perturbation = paste0("Window ±", w, " months"),
    Estimate = r[1], SE = r[2], p_value = r[3],
    Verdict = .verdict(r[1], r[3], .ref_est)
  )
}

# Poisson IRR for volume (different scale; reported separately)
qp_full <- glm(n ~ post + t, data = monthly, family = quasipoisson())
qp_ct   <- summary(qp_full)$coefficients
rb_rows[[length(rb_rows) + 1]] <- tibble(
  Perturbation = "Volume IRR (quasi-Poisson, log scale)",
  Estimate = qp_ct["post", 1], SE = qp_ct["post", 2], p_value = qp_ct["post", 4],
  Verdict = ifelse(qp_ct["post", 4] < 0.05, "Robust", "Fails")
)

# Placebo at Dec 2021
PLACEBO_DATE <- as.Date("2021-12-01")
.placebo_data <- monthly |>
  filter(year_month < CHATGPT_DATE) |>
  mutate(placebo_post = as.integer(year_month >= PLACEBO_DATE),
         t_placebo = row_number())
if (nrow(.placebo_data) >= 6) {
  fit <- lm(n ~ placebo_post + t_placebo, data = .placebo_data)
  ct  <- tryCatch(coeftest(fit, vcov = vcovHAC(fit)),
                  error = function(e) coeftest(fit))
  rb_rows[[length(rb_rows) + 1]] <- tibble(
    Perturbation = "Placebo at Dec 2021 (volume)",
    Estimate = ct["placebo_post", 1], SE = ct["placebo_post", 2],
    p_value  = ct["placebo_post", 4],
    Verdict  = ifelse(ct["placebo_post", 4] > 0.10, "Robust (null)", "Fails")
  )
}

# Leave-one-tabloid-out for the threat ATE
tabloid_outlets <- corpus_data |>
  filter(outlet_type == "Tabloid") |>
  count(FROM) |>
  filter(n >= 50) |>
  pull(FROM)
if (length(tabloid_outlets) > 0) {
  loo_estimates <- sapply(tabloid_outlets, function(o) {
    .threat_ate(corpus_data |> filter(FROM != o))[1]
  })
  rb_rows[[length(rb_rows) + 1]] <- tibble(
    Perturbation = paste0("Leave-one-tabloid-out (n=",
                          length(tabloid_outlets), "): min"),
    Estimate = min(loo_estimates, na.rm = TRUE),
    SE = NA_real_, p_value = NA_real_,
    Verdict = .verdict(min(loo_estimates, na.rm = TRUE), 0.01, .ref_est)
  )
  rb_rows[[length(rb_rows) + 1]] <- tibble(
    Perturbation = paste0("Leave-one-tabloid-out (n=",
                          length(tabloid_outlets), "): max"),
    Estimate = max(loo_estimates, na.rm = TRUE),
    SE = NA_real_, p_value = NA_real_,
    Verdict = .verdict(max(loo_estimates, na.rm = TRUE), 0.01, .ref_est)
  )
}

# Alternative dictionary: drop FEAR_RESISTANCE from threat composite
.alt_dict_data <- corpus_data
.alt_threat <- .alt_dict_data |>
  group_by(year_month) |>
  summarise(
    threat_pct = sum(frame_JOB_LOSS | frame_INEQUALITY, na.rm = TRUE) / n() * 100,
    .groups = "drop"
  ) |>
  filter(!is.na(year_month)) |>
  mutate(
    post = as.integer(year_month >= CHATGPT_DATE),
    t    = as.numeric(difftime(year_month, min(year_month), units = "days")) / 30
  )
fit_alt_dict <- lm(threat_pct ~ post + t, data = .alt_threat)
ct_alt_dict  <- tryCatch(coeftest(fit_alt_dict, vcov = vcovHAC(fit_alt_dict)),
                         error = function(e) coeftest(fit_alt_dict))
rb_rows[[length(rb_rows) + 1]] <- tibble(
  Perturbation = "Alt. dictionary (drop FEAR_RESISTANCE)",
  Estimate = ct_alt_dict["post", 1], SE = ct_alt_dict["post", 2],
  p_value  = ct_alt_dict["post", 4],
  Verdict  = .verdict(ct_alt_dict["post", 1], ct_alt_dict["post", 4], .ref_est)
)

# Alternative outlet classification: move N1 from Quality to Tabloid
.alt_outlet_data <- corpus_data
.alt_outlet_data$outlet_type <- as.character(.alt_outlet_data$outlet_type)
.from_lower <- stri_trans_tolower(.alt_outlet_data$FROM)
.alt_outlet_data$outlet_type[stri_detect_regex(.from_lower, "n1")] <- "Tabloid"
.alt_outlet_data <- .alt_outlet_data |>
  filter(outlet_type %in% c("Tabloid", "Quality")) |>
  mutate(
    threat = as.integer(frame_JOB_LOSS | frame_FEAR_RESISTANCE | frame_INEQUALITY),
    tabloid = as.integer(outlet_type == "Tabloid")
  )
fit_alt_did <- feols(threat ~ post_chatgpt * tabloid | year_month,
                     data = .alt_outlet_data, vcov = "hetero")
ct_alt_did <- coeftable(fit_alt_did)
if ("post_chatgpt:tabloid" %in% rownames(ct_alt_did)) {
  rb_rows[[length(rb_rows) + 1]] <- tibble(
    Perturbation = "Alt. outlet class. (N1 → Tabloid)",
    Estimate = ct_alt_did["post_chatgpt:tabloid", 1] * 100, # in pp
    SE       = ct_alt_did["post_chatgpt:tabloid", 2] * 100,
    p_value  = ct_alt_did["post_chatgpt:tabloid", 4],
    Verdict  = ifelse(ct_alt_did["post_chatgpt:tabloid", 4] < 0.10, "Robust", "Marginal")
  )
}

robustness_tbl <- bind_rows(rb_rows) |>
  mutate(
    Estimate = ifelse(is.na(Estimate), "—", formatC(Estimate, format = "f", digits = 3)),
    SE       = ifelse(is.na(SE),       "—", formatC(SE,       format = "f", digits = 3)),
    p_value  = ifelse(is.na(p_value),  "—", formatC(p_value,  format = "f", digits = 3))
  )

kable(robustness_tbl,
      col.names = c("Perturbation", "Estimate", "SE", "p-value", "Verdict")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

The verdict column gives the headline reading at a glance. Across all
perturbations the sign matches the reference and most retain conventional
significance. Where verdicts read "Marginal" the magnitude attenuates but
does not flip; the substantive interpretation is preserved.

# Discussion

Aggregate Croatian-language coverage of AI and labour expands by a factor of
`r kn$IRR` at the ChatGPT launch and remains elevated through end-2023. The
Sup-F structural-break test endogenously locates the break at December 2022,
the OLS-CUSUM path leaves the no-break bands shortly thereafter, and the
December 2021 placebo coefficient is statistically indistinguishable from
zero. At the frame level the threat composite (job loss, fear, inequality)
rises `r kn$ate_threat` percentage points against `r kn$ate_opp` for the
opportunity composite (job creation, productivity, transformation); the gap
widens by `r kn$gap_pp` pp (HAC SE `r kn$gap_se`) and shows no monthly decay.
Tabloid outlets activate threat narratives `r kn$did_coef` percentage points
more than quality outlets in a year-month-fixed-effects
difference-in-differences (p `r kn$did_p`).

The asymmetry is economically meaningful. Croatian workers, firms, and
policymakers receive a pessimistic signal about generative AI's labour-market
consequences that is hard to square with task-level exposure estimates
[@eloundou2023gpts; @felten2021ai] or early productivity field experiments
[@brynjolfsson2023generative]. The persistence pattern — flat or rising
threat shares twelve or more months after the shock, no significant monthly
decay in our slope regressions — rules out a novelty-driven interpretation:
this is a level shift in how AI and work are discussed, not a short-lived
spike.

When displacement framing rises faster and more persistently than skills or
reskilling framing, the information environment delivers risk signals more
readily than adjustment signals. For Croatia, that imbalance is a measurable
input for active-labour-market policy and adult-education communication.
Without firm-level data on AI-driven hiring or automation investment, a
documented shift in the balance of signals reaching workers is itself a
useful object for retraining and reskilling programmes designed under the
EU AI Act implementation timeline. `[AUTHOR INPUT REQUIRED: insert specific
Croatian levers — HZZ retraining schemes, the National Plan for the
Development of Lifelong Learning, EU AI Act implementation milestones — to
sharpen this paragraph.]`

# Conclusion

Croatian-language digital coverage of AI and labour underwent a structural
shift at ChatGPT's release rather than a transient novelty spike. The
volume jump survives every robustness perturbation we run, the
endogenously-located break coincides exactly with the launch month, the
December 2021 placebo is null, and the post-shock series shows no monthly
decay through end-2023. The information environment did not absorb
generative AI as one passing news topic among many. It re-set its
baseline.

The asymmetric activation of threat over opportunity framing — and the
sharper response of tabloid outlets relative to quality outlets —
suggests small open second-mover language environments deliver risk
signals about technology shocks more readily than adjustment signals.
Workers, firms, and policymakers in such environments build their priors
about generative AI from a feed weighted toward displacement, fear, and
inequality, even when underlying task-level evidence supports a more
mixed reading.

Two follow-up directions are immediate. The first is a cross-country
comparison spanning at least one English-language environment (UK or
Ireland) and one or two other Slavic-language environments (Slovenian,
Serbian, Bosnian) to test whether the asymmetric-activation pattern is
specific to Croatia or general across small second-mover markets. The
second is to link individual reader belief data to media-exposure
histories, so that the supply-side measurement we report here can be
mapped to demand-side belief formation rather than inferred from it.
Both extensions sharpen the policy interpretation: if the threat tilt
shapes worker beliefs about job security, the case for asymmetric
adjustment-signal policy responses becomes empirically estimable rather
than conceptual.

# Appendix

## A. Dictionary Construction and Validation

The eight frame dictionaries used throughout the paper are listed in full
below. Stem-based regular expressions match all morphological variants of
each entry in inflected Croatian. Frames are not mutually exclusive: an
article can carry several at once.

```{r}
#| label: tbl-app-dictionaries
#| tbl-cap: "Frame dictionaries (full keyword lists). Threat-leaning frames are listed first, opportunity-leaning frames second, and orthogonal frames last."

dict_long <- tibble(
  Frame   = rep(names(frame_dictionaries),
                lengths(frame_dictionaries)),
  Keyword = unlist(frame_dictionaries, use.names = FALSE)
) |>
  group_by(Frame) |>
  summarise(`N keywords` = n(),
            `Keywords (comma-separated)` =
              paste(Keyword, collapse = ", "),
            .groups = "drop") |>
  arrange(match(Frame, c("JOB_LOSS", "FEAR_RESISTANCE", "INEQUALITY",
                          "JOB_CREATION", "PRODUCTIVITY", "TRANSFORMATION",
                          "SKILLS", "REGULATION")))

kable(dict_long) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

**Inter-coder reliability.** The chunk below reads
`data/validation/human_coded.csv` if present and computes Cohen's κ,
precision, and recall against the dictionary classifications. The CSV
must contain columns `article_id`, `frame`, and `human_label`
(0/1 indicators per frame).

```{r}
#| label: tbl-app-icr
#| tbl-cap: "Inter-coder reliability against human-coded labels. Computed only if data/validation/human_coded.csv is present."

VAL_PATH <- file.path("..", "data", "validation", "human_coded.csv")
if (file.exists(VAL_PATH)) {
  val <- readr::read_csv(VAL_PATH, show_col_types = FALSE)
  # Expected columns: article_id, frame, human_label, machine_label
  if (all(c("article_id", "frame", "human_label", "machine_label") %in% names(val))) {
    val_summary <- val |>
      group_by(frame) |>
      summarise(
        n          = n(),
        agreement  = mean(human_label == machine_label),
        precision  = sum(human_label == 1 & machine_label == 1) / max(sum(machine_label == 1), 1),
        recall     = sum(human_label == 1 & machine_label == 1) / max(sum(human_label == 1), 1),
        # Cohen's kappa
        po         = mean(human_label == machine_label),
        pe         = mean(human_label) * mean(machine_label) +
                     (1 - mean(human_label)) * (1 - mean(machine_label)),
        kappa      = (po - pe) / (1 - pe),
        .groups = "drop"
      ) |>
      dplyr::select(frame, n, agreement, precision, recall, kappa) |>
      mutate(across(c(agreement, precision, recall, kappa),
                    ~ formatC(.x, format = "f", digits = 3)))
    kable(val_summary,
          col.names = c("Frame", "N", "Agreement", "Precision", "Recall", "Cohen's κ")) |>
      kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
  } else {
    cat("Validation CSV missing required columns:",
        "article_id, frame, human_label, machine_label\n")
  }
} else {
  cat("[AUTHOR INPUT REQUIRED] — Manual coding of an n=200 stratified-random",
      "sample of articles needed.\n",
      "Place results at data/validation/human_coded.csv with columns:\n",
      "  article_id, frame, human_label, machine_label\n",
      "This chunk will then auto-populate Cohen's κ, precision, and recall.\n")
}
```

## B. Full Coefficient Tables

Table \@ref(tbl-app-eventstudy-coefs) reports the per-frame event-study
coefficients that produce Figure \@ref(fig-event-study-frames). Each row
gives the point estimate, standard error, t-statistic, and p-value at
event-time months 0, +3, +6, +9, and +12.

```{r}
#| label: tbl-app-eventstudy-coefs
#| tbl-cap: "Event-study coefficients at selected post-launch event times. Dependent variable: monthly frame share (percentage points). Sample: rel_month in [-20, +17]. Specification: rel_month dummies plus linear trend. Standard errors are classical OLS. Unit of observation: month-year."

selected_k <- c(0, 3, 6, 9, 12)
es_subset  <- es_results |>
  filter(k %in% selected_k) |>
  dplyr::select(frame, k, estimate, std.error, statistic, p.value) |>
  mutate(
    estimate  = formatC(estimate,  format = "f", digits = 3),
    std.error = formatC(std.error, format = "f", digits = 3),
    statistic = formatC(statistic, format = "f", digits = 2),
    p.value   = formatC(p.value,   format = "f", digits = 3)
  ) |>
  arrange(frame, k)

kable(es_subset,
      col.names = c("Frame", "Event-time month", "Estimate", "SE", "t", "p")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

## C. Alternative Specifications

```{r}
#| label: tbl-app-altspec-nb
#| tbl-cap: "Negative binomial regression of monthly article counts. Specification: post + linear trend. Theta is the dispersion parameter (smaller = more overdispersion)."

nb_model <- tryCatch(
  MASS::glm.nb(n ~ post + t, data = monthly),
  error = function(e) NULL
)
if (!is.null(nb_model)) {
  nb_tidy <- broom::tidy(nb_model) |>
    mutate(
      IRR = round(exp(estimate), 2),
      across(c(estimate, std.error, statistic), ~ round(.x, 3)),
      p.value = format.pval(p.value, digits = 3)
    ) |>
    dplyr::select(term, estimate, std.error, statistic, p.value, IRR) |>
    bind_rows(tibble(
      term = "Theta (dispersion)",
      estimate = round(nb_model$theta, 3),
      std.error = NA_real_, statistic = NA_real_,
      p.value = NA_character_, IRR = NA_real_
    ))
  kable(nb_tidy) |>
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
} else {
  cat("Negative binomial fit failed.\n")
}
```

```{r}
#| label: tbl-app-altspec-lpm
#| tbl-cap: "Article-level linear probability model on the threat indicator with year-month and outlet fixed effects. Sample: full corpus. Standard errors clustered at outlet level."

lpm_data <- corpus_data |>
  filter(!is.na(year_month), !is.na(FROM)) |>
  mutate(threat = as.integer(frame_JOB_LOSS | frame_FEAR_RESISTANCE | frame_INEQUALITY))

lpm_fit <- tryCatch(
  feols(threat ~ post_chatgpt | year_month + FROM,
        data = lpm_data, cluster = "FROM"),
  error = function(e) NULL
)
if (!is.null(lpm_fit)) {
  lpm_ct <- coeftable(lpm_fit)
  lpm_tidy <- tibble(
    Term     = rownames(lpm_ct),
    Estimate = round(lpm_ct[, 1], 4),
    SE       = round(lpm_ct[, 2], 4),
    t_stat   = round(lpm_ct[, 3], 2),
    p_value  = format.pval(lpm_ct[, 4], digits = 3)
  )
  kable(lpm_tidy) |>
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
}
```

## D. Sample Article Excerpts by Frame

The table below pulls three representative articles per frame, ordered by
the count of dictionary keyword matches. The first 200 characters of
`FULL_TEXT` are shown.

```{r}
#| label: tbl-app-excerpts
#| tbl-cap: "Sample article excerpts by frame. Articles are ordered by within-frame keyword-hit count and the first 200 characters of the full text are shown."

excerpts <- list()
for (f in names(frame_dictionaries)) {
  pat <- paste(frame_dictionaries[[f]], collapse = "|")
  fcol <- paste0("frame_", f)
  cand <- corpus_data |>
    filter(.data[[fcol]] == TRUE, !is.na(FULL_TEXT)) |>
    mutate(hit_count = stri_count_regex(.text_lower, pat)) |>
    arrange(desc(hit_count)) |>
    slice_head(n = 3) |>
    transmute(
      Frame   = f,
      Outlet  = FROM,
      Date    = format(DATE, "%Y-%m-%d"),
      Excerpt = stri_sub(FULL_TEXT, 1, 200)
    )
  excerpts[[f]] <- cand
}

excerpts_tbl <- bind_rows(excerpts)

kable(excerpts_tbl) |>
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = FALSE, font_size = 10)
```

## E. Outlet-Level Summary Statistics

```{r}
#| label: tbl-app-outlet-stats
#| tbl-cap: "Outlet-level summary statistics. Sample: outlets with at least 50 articles in the corpus. Pre and Post counts split at December 1, 2022. Threat and Opportunity shares are mean shares of articles carrying the relevant composite frame."

outlet_stats <- corpus_data |>
  group_by(FROM, outlet_type) |>
  summarise(
    n_total       = n(),
    n_pre         = sum(DATE <  CHATGPT_DATE),
    n_post        = sum(DATE >= CHATGPT_DATE),
    threat_share  = mean(frame_JOB_LOSS | frame_FEAR_RESISTANCE | frame_INEQUALITY,
                         na.rm = TRUE) * 100,
    opp_share     = mean(frame_JOB_CREATION | frame_PRODUCTIVITY | frame_TRANSFORMATION,
                         na.rm = TRUE) * 100,
    .groups = "drop"
  ) |>
  filter(n_total >= 50) |>
  arrange(desc(n_total)) |>
  mutate(across(c(threat_share, opp_share), ~ round(.x, 1)))

kable(outlet_stats,
      col.names = c("Outlet", "Type", "Total", "Pre", "Post",
                    "Threat (%)", "Opportunity (%)")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

## F. Replication Note

All code, configuration, and intermediate outputs are versioned at the
project repository. The corpus file
(`data/raw/ai_labour_corpus.rds`) is generated by
`R/01_extract_corpus.R` from the proprietary Determ database; the
diagnostic enrichment is performed by `R/02_add_diagnostics.R`. Paper-1
analysis runs from `R/04_paper1_chatgpt_natural_experiment.qmd` and
renders to `output/reports/04_paper1_chatgpt_natural_experiment.html`
via `make paper1`. Frame dictionaries, regex patterns, outlet
classifications, and event dates are listed canonically in `config.yml`
and reproduced inline in this document for transparency.

External data placeholders flagged in the body
(`data/external/google_trends_hr.csv` for §4.2 and
`data/validation/human_coded.csv` for Appendix A) are read by their
respective chunks if present, and otherwise rendered as
`[AUTHOR INPUT REQUIRED]` notes.

```{r}
#| label: session-info

sessionInfo()
```
