---
title: "The ChatGPT Shock and Public Information about AI and Work"
subtitle: "Event-Study Evidence from Croatian Digital Media (2021–2024)"
author: "Media Analysis Research"
date: today
bibliography: ../references.bib
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    code-fold: true
    code-tools: true
    code-summary: "Show code"
    df-print: paged
    fig-width: 10
    fig-height: 6
    fig-dpi: 300
    embed-resources: true
execute:
  warning: false
  message: false
  echo: false

---

```{r}
#| label: setup
#| include: false

# ==========================================================================
# PACKAGES
# ==========================================================================

required_packages <- c(
  "dplyr", "tidyr", "stringr", "stringi", "lubridate", "forcats", "tibble",
  "ggplot2", "scales", "patchwork", "ggrepel",
  "knitr", "kableExtra",
  "fixest",        # fast fixed effects estimation
  "sandwich",      # robust standard errors
  "lmtest",        # coefficient tests
  "strucchange",   # structural break detection
  "MASS",          # negative binomial
  "broom",         # tidy model output
  "zoo"            # rolling functions
)

for (pkg in required_packages) {
  if (!require(pkg, character.only = TRUE, quietly = TRUE)) {
    install.packages(pkg, quiet = TRUE)
    library(pkg, character.only = TRUE)
  }
}

options(dplyr.summarise.inform = FALSE, scipen = 999)

# --------------------------------------------------------------------------
# THEME — black-and-white, publication-ready (greyscale only)
# --------------------------------------------------------------------------

theme_pub <- theme_classic(base_size = 11) +
  theme(
    plot.title       = element_blank(),
    plot.subtitle    = element_blank(),
    plot.caption     = element_text(color = "gray30", size = 9, hjust = 0),
    legend.position  = "bottom",
    legend.title     = element_blank(),
    panel.grid.major = element_line(color = "gray90", linewidth = 0.3),
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    strip.text       = element_text(face = "bold", size = 10),
    axis.title       = element_text(size = 11),
    axis.text        = element_text(color = "black")
  )
theme_set(theme_pub)

# Greyscale palette for two-series plots (threat vs. opportunity, etc.)
bw_two   <- c("black", "gray55")
bw_three <- c("black", "gray45", "gray70")
```

# Introduction

A central question in the economics of automation is whether new technologies
displace, complement, or transform human work [@acemoglu2018race; @autor2015why;
@bessen2020automation]. The release of ChatGPT on November 30, 2022 sharpened
this question by making generative artificial intelligence (AI) visible to a
mass audience essentially overnight. Recent task-level evidence suggests that
large language models are exposed to a wide and disproportionately white-collar
share of the occupational distribution [@eloundou2023gpts; @felten2021ai] and
that, in deployed settings, generative AI raises productivity for some workers
while compressing observed skill premia [@brynjolfsson2023generative]. Whether
these capabilities ultimately produce displacement, reallocation, or
augmentation of labour remains an open empirical question, and will likely
remain so for years to come.

Workers, firms, and policymakers must nonetheless form beliefs and take decisions
about reskilling, hiring, occupational choice, and regulation under substantial
uncertainty about AI capabilities and their economic consequences. In small open
economies such as Croatia, where systematic firm-level data on AI adoption is
sparse, the main channel through which workers and policymakers learn about
AI's labour-market implications is mediated coverage in digital news and social
media. Knowing what signals that channel transmits, and how they shift in
response to major capability announcements, is thus a prerequisite for
understanding how beliefs form and how adjustment unfolds.

This paper exploits the ChatGPT launch as a natural experiment on the public
information environment about AI and work. We treat November 30, 2022 as an
exogenous shock to the salience of generative-AI capabilities and estimate its
effect on Croatian digital media coverage of the AI-labour intersection. As in
standard event-study designs, the identifying assumption is that Croatian media
did not anticipate the ChatGPT release, so a sharp break in coverage at that
date reflects the shock rather than a pre-existing trend. We support this
assumption with formal structural-break tests and a placebo exercise at
December 2021.

We use a corpus of Croatian-language articles and posts from 2021 to 2024
extracted from the Determ media-monitoring platform (approximately 20 million
underlying records across web portals and seven social platforms). Within this
corpus we measure the prevalence of eight interpretive frames covering the
main channels through which the literature links AI to labour-market outcomes,
including job displacement, job creation, task transformation, skills and
reskilling, regulation, productivity, inequality, and fear or resistance. We
address four empirical questions: how large is the ChatGPT shock on aggregate
media attention; which frames activate and how quickly; whether the effect
differs across platforms and outlet types; and whether post-shock framing shifts
persist or fade over time.

The paper makes three contributions. First, it provides event-study evidence
on how a generative-AI capability shock reshapes public information about
labour-market consequences in a small open European economy where direct
measurement of AI exposure or adoption is currently infeasible and where worker
expectations and policy responses are still forming. Second, threat-related
frames (job loss, fear, inequality) activate faster and persist longer than
opportunity-related frames (job creation, productivity, transformation),
suggesting that the public information environment overweights displacement
risk relative to the adjustment margins stressed in the labour-economics
literature [@acemoglu2018race; @autor2022work]. Third, the shock propagated
unevenly across media platforms and outlet types, reflecting differences in
audience and editorial orientation. The results map the supply side of public
information about AI and work and offer a measurable input to subsequent
research connecting media framing to realised labour-market outcomes.

<!--
COLLABORATOR NOTE: literature engagement here is intentionally light.
Natural directions to deepen, in order of priority:
  (i)  Automation-vs-augmentation debate: Acemoglu-Restrepo task framework,
       Autor's reinstatement effect, Bessen on automation and employment.
  (ii) Expectations and beliefs: Coibion-Gorodnichenko on expectation
       formation, Roth-Wohlfart on news shocks and beliefs, Bursztyn et al.
       on information provision.
  (iii) Croatian and CEE labour-market context: HZZ data, EU AI Act
        implementation, services-heavy occupational structure.
  (iv) Media-economics on supply and framing: Gentzkow-Shapiro on slant,
       Mullainathan-Shleifer on media bias, agenda-setting tradition.
  (v)  Generative-AI productivity field experiments: Noy-Zhang, Peng et al.
       (Copilot), Dell'Acqua et al. (BCG).
-->


# Data and Measurement

## Setting

Croatia is a small open European economy (population approximately 3.85
million, EU member since 2013 and euro-area member since 2023). The Croatian
labour market is services-heavy, has a sizeable public-sector employment
share, and operates under structurally tight labour supply driven by
emigration and demographic decline. Direct microdata on firm-level AI
adoption or AI-driven hiring is essentially unavailable for Croatia, so the
public information environment about generative AI is shaped almost entirely
by media coverage, much of it translated or paraphrased from international
sources. Croatia is therefore a useful case for studying what signals the media
environment transmits to workers, firms, and policymakers, and what happens
to those signals when a major AI capability becomes publicly available.

The Croatian Employment Service (HZZ) maintains monthly vacancy and
registered-unemployment series at the occupational level. We treat these as a
complementary data layer for follow-up work linking the framing patterns
documented here to realised labour-market outcomes (job search, vacancy
posting, reskilling enrolment), and leave that linkage outside the scope of
the present paper.

<!--
COLLABORATOR NOTE: verify the headline figures (population, EU/euro-area
accession dates) against the latest Eurostat/CBS (Croatian Bureau of
Statistics) releases at the time of submission. Two or three sentences on
Croatian sectoral structure (services share, tourism, public sector) and
on the recent CEE labour-market literature would strengthen this paragraph.
-->

## Corpus Construction and Loading

The corpus draws on the Determ media-monitoring database, which covers
approximately 20 million records of Croatian-language digital media content
between January 2021 and May 2024. Determ aggregates articles and posts from
web portals, Facebook, Twitter (X), Instagram, YouTube, TikTok, Reddit, and
public forums, capturing what is, to our knowledge, the most comprehensive
record of the Croatian-language digital information environment for the period
we study.

We construct the analytical corpus through a two-stage intersection filter. An
item enters the corpus only if its title or full text contains at least one
term from an AI-and-technology vocabulary *and* at least one term from a
labour-market vocabulary. The first vocabulary covers terms such as
*umjetna inteligencija* (artificial intelligence), *strojno učenje* (machine
learning), ChatGPT, GPT, OpenAI, *automatizacija*, *robotizacija*,
*algoritam*, *neuronska mreža*, *chatbot*, and related expressions. The
second vocabulary covers labour-market terms including *posao* (job),
*zaposlen* (employed), *nezaposlenost* (unemployment), *radno mjesto*
(workplace), *tržište rada* (labour market), *vještine* (skills), *karijera*
(career), *produktivnost* (productivity), *otpuštanje* (lay-offs),
*prekvalifikacija* (reskilling), and *plaća* (wage). The intersection design
substantially reduces the false-positive rate associated with single-keyword
searches, since articles must engage both topics jointly to qualify.

Croatian is a highly inflected language: nouns decline through seven cases,
verbs conjugate, and adjectives agree in gender, number, and case with the
nouns they modify. We accommodate this morphological variation through
stem-based regular expressions (for example, the stem `zaposlen` matches
*zaposleni*, *zaposlenika*, *zaposlenik*, and related inflections; the
pattern `umjetn.*inteligencij` matches all case combinations of *umjetna
inteligencija*). Filtering proceeds in two passes for performance: an
in-database SQL `LIKE` filter narrows the candidate set, and a regular-
expression refinement in R applies the precise morphological patterns. The
extraction script and full keyword tables are reproducible from the
companion repository.

After deduplication on title, date, and source, and after restricting to the
analysis window 2021–2024, the working corpus comprises the article counts
shown in Table \@ref(tbl-summary).

```{r}
#| label: load-corpus

CORPUS_PATH <- "../data/raw/ai_labour_corpus.rds"

if (!file.exists(CORPUS_PATH)) {
  stop("Corpus file not found at: ", CORPUS_PATH,
       "\nRun 01_extract_corpus.R first.")
}

corpus_raw <- readRDS(CORPUS_PATH)

# --------------------------------------------------------------------------
# Prepare text and temporal variables
# --------------------------------------------------------------------------

corpus_data <- corpus_raw |>
  mutate(
    DATE = as.Date(DATE),
    .text_lower = stri_trans_tolower(
      paste(coalesce(TITLE, ""), coalesce(FULL_TEXT, ""), sep = " ")
    ),
    year        = year(DATE),
    month       = month(DATE),
    year_month  = floor_date(DATE, "month"),
    year_week   = floor_date(DATE, "week"),
    quarter     = quarter(DATE),
    year_quarter = paste0(year, " Q", quarter),
    word_count  = stri_count_regex(FULL_TEXT, "\\S+")
  ) |>
  filter(!is.na(DATE), DATE < as.Date("2024-01-01")) |>
  distinct(TITLE, DATE, FROM, .keep_all = TRUE) |>
  arrange(DATE)

cat("Corpus loaded:", format(nrow(corpus_data), big.mark = ","), "articles\n")
cat("Date range:", as.character(min(corpus_data$DATE)), "to",
    as.character(max(corpus_data$DATE)), "\n")
```

## Frame Detection

We operationalize eight interpretive frames, each corresponding to a distinct
way in which media discuss the relationship between AI and labour. Frames are
detected through curated Croatian keyword dictionaries with stem-based
regular expressions, again to accommodate inflection. An article is coded as
carrying a given frame if any of the dictionary patterns for that frame
matches the text. The eight frames are grouped naturally into threat-leaning
content, opportunity-leaning content, and orthogonal content.

The threat-leaning frames cover *job loss* (terms such as *gubitak posla*,
*ukidanje radnih mjesta*, *zamjena radnika*, *otpuštanje*, *tehnološka
nezaposlenost*), *fear and resistance* (*strah od AI*, *prijetnja*,
*opasnost*, *neizvjesnost*, *panika*), and *inequality* (*nejednakost*,
*digitalni jaz*, *polarizacija*, *nestanak srednje klase*). The
opportunity-leaning frames cover *job creation* (*nova radna mjesta*,
*potražnja za stručnjacima*, *nova zanimanja*), *productivity* (*učinkovitost*,
*ušteda*, *gospodarski rast*, *optimizacija*), and *task transformation*
(*transformacija rada*, *prilagodba*, *suradnja čovjeka i*, *evolucija
poslova*). The two remaining frames are *skills* (*prekvalifikacija*,
*cjeloživotno učenje*, *digitalne vještine*, *reskilling*, *upskilling*) and
*regulation* (*regulacija AI*, *AI Act*, *zaštita radnika*, *prava radnika*,
*sindikat*).

Frames are not mutually exclusive: an article that discusses both
displacement and reskilling will be coded as carrying both. We treat each
frame as a binary indicator and analyse its monthly prevalence (the share of
articles carrying the frame) over time.

```{r}
#| label: frame-detection

frame_dictionaries <- list(
  JOB_LOSS = c(
    "gubitak posla", "gubitak poslova", "gubitak radnih mjesta",
    "ukidanje radnih mjesta", "ukidanje poslova",
    "zamjena radnika", "zamijeniti radnike", "zamjenjuje radnike",
    "istisnuti radnike", "istiskivanje",
    "otpuštanje", "otpuštanja",
    "nestanak poslova", "nestanak zanimanja",
    "suvišan", "suvišni radnici",
    "tehnološka nezaposlenost",
    "krade poslove", "krade posao", "oduzima posao",
    "prijeti radnim mjestima", "ugrožava radna mjesta"
  ),
  JOB_CREATION = c(
    "nova radna mjesta", "novi poslovi", "novo zapošljavanje",
    "nove prilike", "nove mogućnosti",
    "stvaranje poslova",
    "rast zapošljavanja", "povećanje zapošljavanja",
    "nova zanimanja", "nova karijera",
    "potražnja za stručnjacima", "nedostatak radnika",
    "deficitarna zanimanja"
  ),
  TRANSFORMATION = c(
    "transformacija rada", "transformacija poslova",
    "promjena načina rada", "mijenja način rada",
    "prilagodba", "prilagoditi se", "prilagođavanje",
    "nadopunjuje", "komplementar",
    "suradnja čovjeka i", "čovjek i stroj",
    "evolucija poslova", "evolucija rada",
    "nove uloge", "promijenjena uloga",
    "ne zamjenjuje nego"
  ),
  SKILLS = c(
    "prekvalifikacija", "dokvalifikacija",
    "cjeloživotno učenje",
    "digitalna pismenost", "digitalne vještine",
    "nova znanja", "nove vještine",
    "jaz u vještinama", "nedostatak vještina",
    "reskilling", "upskilling",
    "obrazovanje za budućnost",
    "stem vještine", "programiranje"
  ),
  REGULATION = c(
    "regulacija ai", "regulativa",
    "zakon o ai", "zakonski okvir",
    "eu regulativa", "ai act",
    "etička pitanja", "etika ai",
    "sindikat", "sindikalni",
    "zaštita radnika", "prava radnika",
    "socijalna zaštita"
  ),
  PRODUCTIVITY = c(
    "produktivnost", "povećanje produktivnosti",
    "učinkovitost", "efikasnost",
    "ušteda", "smanjenje troškova",
    "konkurentnost", "konkurentna prednost",
    "gospodarski rast", "ekonomski rast",
    "optimizacija"
  ),
  INEQUALITY = c(
    "nejednakost", "rastuća nejednakost",
    "digitalni jaz", "tehnološki jaz",
    "socijalna nejednakost",
    "polarizacija",
    "jaz u plaćama",
    "ranjive skupine", "marginalizirani",
    "srednja klasa", "nestanak srednje klase"
  ),
  FEAR_RESISTANCE = c(
    "strah od ai", "strah od gubitka", "strah od tehnologij",
    "prijetnja", "opasnost",
    "apokalipsa", "distopija", "katastrofa",
    "upozorenje", "alarm",
    "otpor prema", "protivljenje",
    "neizvjesnost", "nesigurnost",
    "panika", "zabrinutost"
  )
)

# Detect frames
for (fname in names(frame_dictionaries)) {
  pattern <- paste(frame_dictionaries[[fname]], collapse = "|")
  corpus_data[[paste0("frame_", fname)]] <- stri_detect_regex(
    corpus_data$.text_lower, pattern
  )
}

frame_cols <- paste0("frame_", names(frame_dictionaries))
```

## Outlet Classification

For the outlet-type heterogeneity analysis, we classify Croatian web sources
into six categories based on the source name in the database. *Tabloid*
outlets (24sata, Index, Net.hr) emphasise high-volume, sensational coverage.
*Quality* outlets (Jutarnji list, Večernji list, N1, Telegram, Tportal,
Dnevnik) typically run longer and more analytical pieces. *Regional* outlets
(Slobodna Dalmacija, Novi list) cover local perspectives. *Public*
broadcasting (HRT) operates under regulatory obligations of balanced
coverage. *Tech* outlets (Bug, Netokracija) target a technologically
literate audience. *Business* outlets (Poslovni dnevnik, Lider, Forbes
Hrvatska) frame stories economically. Sources that do not match any of these
patterns are grouped as *Other*.

The classification is implemented through case-insensitive regular
expressions matched against the source name. The mapping is deliberately
conservative: each outlet appears in at most one category, and ambiguous
cases are left in *Other* rather than forced into a category.

```{r}
#| label: outlet-classification

outlet_patterns <- tribble(
  ~pattern,              ~outlet_type,
  "24sata",              "Tabloid",
  "index",               "Tabloid",
  "net\\.hr",            "Tabloid",
  "jutarnji",            "Quality",
  "vecernji",            "Quality",
  "dnevnik",             "Quality",
  "n1",                  "Quality",
  "tportal",             "Quality",
  "telegram",            "Quality",
  "slobodna.*dalmacija", "Regional",
  "novi.*list",          "Regional",
  "hrt",                 "Public",
  "bug",                 "Tech",
  "netokracija",         "Tech",
  "poslovni",            "Business",
  "lider",               "Business",
  "forbes.*hr",          "Business"
)

corpus_data$outlet_type <- "Other"
for (i in seq_len(nrow(outlet_patterns))) {
  hit <- stri_detect_regex(
    stri_trans_tolower(corpus_data$FROM), outlet_patterns$pattern[i]
  )
  corpus_data$outlet_type[hit] <- outlet_patterns$outlet_type[i]
}

corpus_data$outlet_type <- factor(
  corpus_data$outlet_type,
  levels = c("Tabloid", "Quality", "Regional", "Public",
             "Tech", "Business", "Other")
)
```

## Descriptive Summary

Table \@ref(tbl-summary) summarises the working corpus. The total number of
items, the share covered before and after the ChatGPT launch in December
2022, and the typical article length give a first sense of the data on
which the event-study analysis rests. The pre-ChatGPT and post-ChatGPT
splits in the final two rows already hint at the asymmetric volume shift
that the formal analysis confirms below.

```{r}
#| label: tbl-summary
#| tbl-cap: "Corpus descriptive statistics"

summary_tbl <- tibble(
  Metric = c(
    "Total articles",
    "Unique sources",
    "Date range",
    "Mean word count",
    "Median word count",
    "Pre ChatGPT articles (before Dec 2022)",
    "Post ChatGPT articles (Dec 2022 onward)"
  ),
  Value = c(
    format(nrow(corpus_data), big.mark = ","),
    format(n_distinct(corpus_data$FROM), big.mark = ","),
    paste(min(corpus_data$DATE), "to", max(corpus_data$DATE)),
    round(mean(corpus_data$word_count, na.rm = TRUE)),
    round(median(corpus_data$word_count, na.rm = TRUE)),
    format(sum(corpus_data$DATE < as.Date("2022-12-01")), big.mark = ","),
    format(sum(corpus_data$DATE >= as.Date("2022-12-01")), big.mark = ",")
  )
)

kable(summary_tbl) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

# Empirical Strategy

## Event Definition and Identification

We define the treatment date as December 1, 2022, the first full day after
ChatGPT's public release. The identifying assumption is that the timing of the
ChatGPT release was not anticipated by Croatian media outlets, so a sharp
break in coverage at that date can be attributed to the shock rather than to
a smooth pre-existing trend. Two secondary events are used for robustness
purposes only: the GPT-4 release on March 14, 2023 and the EU AI Act adoption
on March 13, 2024.

```{r}
#| label: event-setup

# Event dates
CHATGPT_DATE  <- as.Date("2022-12-01")
GPT4_DATE     <- as.Date("2023-03-14")
EU_AI_ACT     <- as.Date("2024-03-13")

# Treatment indicator
corpus_data$post_chatgpt <- as.integer(corpus_data$DATE >= CHATGPT_DATE)

# Relative time in months (for event study)
corpus_data$rel_month <- interval(CHATGPT_DATE, corpus_data$year_month) %/% months(1)
```

## Specification

We estimate the following event study specification at the monthly level for
each frame $f$:

$$
Y_{ft} = \alpha_f + \sum_{k \neq -1} \beta_k^f \cdot \mathbf{1}[t - t^* = k] + \gamma_f \cdot t + \varepsilon_{ft}
$$

where $Y_{ft}$ is the share of articles containing frame $f$ in month $t$,
$t^*$ is December 2022, and $k$ indexes months relative to the event. We
normalize to $k = -1$ (November 2022). The coefficient $\beta_k^f$ captures
the change in frame $f$ prevalence in month $k$ relative to the last
pre-treatment month, conditional on a linear time trend.

For aggregate attention we additionally estimate an interrupted time series
specification of monthly article counts on a linear trend, a post-treatment
indicator, and their interaction. We complement this with a Sup-F
structural-break test to verify that the ChatGPT date is statistically
distinguishable from a smooth trend, and a placebo test at December 2021
(reported in the Robustness section) to rule out seasonal artifacts.

## Monthly Frame Panel

The unit of observation for the frame-level event study is the
month–frame cell. For each calendar month and each of the eight frames, we
compute the share of corpus articles in that month carrying the frame. This
produces a balanced monthly panel running from January 2021 to December 2023
that we use throughout the Results section. Aggregating to monthly shares
(rather than article-level indicators) smooths daily noise, removes
mechanical correlation introduced by the changing volume of articles, and
puts the eight frames on the same scale, facilitating comparison across
frames and across time. The same monthly aggregation is applied to article
volume for the interrupted-time-series and structural-break analyses.

```{r}
#| label: frame-panel

# Build monthly panel of frame shares
frame_monthly <- corpus_data |>
  group_by(year_month) |>
  summarise(
    n_total = n(),
    across(all_of(frame_cols), ~ sum(.x, na.rm = TRUE)),
    .groups = "drop"
  ) |>
  filter(!is.na(year_month)) |>
  pivot_longer(
    cols = all_of(frame_cols),
    names_to = "frame",
    values_to = "count"
  ) |>
  mutate(
    frame = str_remove(frame, "frame_"),
    share = count / n_total,
    pct   = share * 100,
    post  = as.integer(year_month >= CHATGPT_DATE),
    t     = as.numeric(difftime(year_month, min(year_month), units = "days")) / 30,
    rel_month = interval(CHATGPT_DATE, year_month) %/% months(1)
  )
```

# Results

## Volume Shock

We begin with the most basic question: did media attention to AI and labour
change discontinuously at the ChatGPT launch? Figure \@ref(fig-volume-shock)
plots the monthly count of articles in the corpus, with the fitted values from
an interrupted-time-series (ITS) regression overlaid. Three reference dates
are marked: the ChatGPT launch in December 2022, the GPT-4 release in March
2023, and the adoption of the EU AI Act in March 2024.

```{r}
#| label: fig-volume-shock
#| fig-cap: "Monthly article volume in the AI–labour corpus, with fitted interrupted-time-series trend. Reference lines mark the ChatGPT launch (December 2022), the GPT-4 release (March 2023), and the EU AI Act adoption (March 2024)."
#| fig-height: 5

monthly <- corpus_data |>
  count(year_month) |>
  filter(!is.na(year_month)) |>
  mutate(
    post = as.integer(year_month >= CHATGPT_DATE),
    t = as.numeric(difftime(year_month, min(year_month), units = "days")) / 30
  )

# Interrupted time series regression
its_model <- lm(n ~ t + post + I(t * post), data = monthly)

monthly$fitted <- predict(its_model)

events_df <- tibble(
  date  = c(CHATGPT_DATE, GPT4_DATE, EU_AI_ACT),
  label = c("ChatGPT\nLaunch", "GPT-4\nRelease", "EU AI Act\nAdopted")
)

ggplot(monthly, aes(x = year_month, y = n)) +
  geom_col(aes(fill = factor(post)), show.legend = FALSE) +
  scale_fill_manual(values = c("0" = "gray80", "1" = "gray45")) +
  geom_line(aes(y = fitted), color = "black", linewidth = 0.9) +
  geom_vline(data = events_df, aes(xintercept = date),
             linetype = "dashed", color = "black", linewidth = 0.4) +
  geom_label(data = events_df,
             aes(x = date, y = max(monthly$n) * 0.95, label = label),
             size = 2.8, fill = "white", color = "black",
             alpha = 0.9, lineheight = 0.9, label.size = 0.2) +
  scale_x_date(date_breaks = "3 months", date_labels = "%b\n%Y") +
  labs(x = NULL, y = "Articles per month")
```

The pattern is striking. From January 2021 through November 2022, monthly
article counts fluctuate around a low and slowly-rising baseline. From
December 2022 onwards, the level shifts sharply upward and remains elevated
for the rest of the sample. The ITS estimates in Table \@ref(tbl-its-volume)
quantify this jump.

```{r}
#| label: tbl-its-volume
#| tbl-cap: "Interrupted-time-series estimates for monthly article volume. Standard errors are HAC-robust."

its_robust <- coeftest(its_model, vcov = vcovHAC(its_model))

its_tidy <- tibble(
  Term = c("Intercept (baseline level)", "Pre-trend (per month)",
           "Level shift (ChatGPT)", "Slope change (post-trend)"),
  Estimate = round(its_robust[, 1], 2),
  SE       = round(its_robust[, 2], 2),
  t_value  = round(its_robust[, 3], 2),
  p_value  = format.pval(its_robust[, 4], digits = 3)
)

kable(its_tidy) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

The third row of Table \@ref(tbl-its-volume) (the level shift at the ChatGPT
date) is the headline number: it captures the immediate upward jump in
monthly volume conditional on the pre-existing trend, and it is large and
statistically significant. The pre-trend coefficient in the second row is
small, indicating that pre-ChatGPT growth in coverage was modest. The
post-trend slope-change coefficient in the fourth row is positive,
suggesting that the post-shock series continues to grow rather than decay
back toward the pre-shock baseline. Taken together, the ITS estimates
support a regime-change interpretation of the ChatGPT launch on aggregate
media attention.

## Structural Break Tests

The ITS estimates assume that the ChatGPT date is the right place to look for
a break. Two formal tests verify that this is so without imposing the date
beforehand. The Sup-F test computes an F-statistic for a possible break at
each candidate date in the central 70 % of the sample and reports the
maximum, with a p-value derived from the distribution of that maximum under
the null of no break. If the test rejects, there is a statistically
detectable break somewhere in the series, and the location of the maximum F
identifies the most likely date. The OLS-CUSUM test takes a different
angle: it checks whether the cumulative sum of recursive residuals from a
no-break model drifts outside the bands it would stay within if the
parameters were stable. A path that crosses the bands is evidence against
parameter stability. Together the two tests give independent verification
that the December 2022 break is not an artefact of imposing the date.

Table \@ref(tbl-structural-break) reports the Sup-F result and the
endogenously-detected breakpoint, and Figure \@ref(fig-cusum) shows the
OLS-CUSUM path.

```{r}
#| label: tbl-structural-break
#| tbl-cap: "Structural-break tests for monthly article volume. The Sup-F test searches for an unknown breakpoint in the central 70% of the sample; the breakpoint algorithm reports the date that maximises the test statistic."

# Build monthly time series object (kept for reference, not displayed)
monthly_ts <- ts(monthly$n, start = c(year(min(monthly$year_month)),
                                       month(min(monthly$year_month))),
                 frequency = 12)

if (nrow(monthly) > 10) {
  fs_test <- Fstats(n ~ t, data = monthly, from = 0.15, to = 0.85)
  bp_detected <- breakpoints(n ~ t, data = monthly, h = 0.15)

  bp_dates <- if (length(bp_detected$breakpoints) > 0) {
    paste(format(monthly$year_month[bp_detected$breakpoints], "%b %Y"),
          collapse = ", ")
  } else {
    "none detected"
  }

  sb_tbl <- tibble(
    Statistic = c("Sup-F test statistic",
                  "Sup-F p-value",
                  "Endogenous breakpoint(s)",
                  "ChatGPT reference date"),
    Value = c(format(round(max(fs_test$Fstats), 2), nsmall = 2),
              format.pval(sctest(fs_test)$p.value, digits = 3),
              bp_dates,
              "Dec 2022")
  )

  kable(sb_tbl) |>
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
}
```

```{r}
#| label: fig-cusum
#| fig-cap: "OLS-CUSUM test for parameter stability of the monthly volume series. The cumulative sum of recursive residuals (solid line) should stay within the dashed reference bands under the null of no structural change."
#| fig-height: 4

if (nrow(monthly) > 10) {
  cusum_test <- efp(n ~ t, data = monthly, type = "OLS-CUSUM")
  par(mar = c(4, 4, 1, 1))
  plot(cusum_test, main = "",
       xlab = "Observation index", ylab = "Cumulative sum",
       col = "black")
}
```

The Sup-F p-value is well below conventional significance thresholds, so the
null of no break is rejected, and the endogenously-detected breakpoint
coincides with the ChatGPT reference date. The OLS-CUSUM path leaves the
no-break confidence bands shortly after that date, which is the signature of
a structural shift rather than gradual drift. The two tests therefore
corroborate the visual impression of Figure \@ref(fig-volume-shock): the
ChatGPT launch is associated with a statistically detectable regime change
in aggregate media attention.

## Event Study Coefficients

The volume results show that overall attention shifted, but they say nothing
about which interpretive lenses became more or less prominent. Figure
\@ref(fig-event-study-frames) displays event-study coefficients for each
of the eight frames. Each panel plots the change in monthly frame prevalence
relative to November 2022 (the month before the ChatGPT launch), with 95 %
confidence bands. The horizontal axis runs from twenty months before to
seventeen months after the launch; the vertical reference line at zero marks
the launch itself.

```{r}
#| label: fig-event-study-frames
#| fig-cap: "Event-study coefficients by interpretive frame, relative to November 2022 (the month before the ChatGPT launch). Each panel plots monthly point estimates and 95% confidence bands from a frame-specific regression of monthly frame share on event-time indicators and a linear time trend."
#| fig-height: 10
#| fig-width: 11

# Estimate event study for each frame
run_event_study <- function(df, frame_name) {
  fdata <- df |> filter(frame == frame_name)

  # Trim endpoints with very few obs
  fdata <- fdata |>
    filter(rel_month >= -20 & rel_month <= 17)

  # Create event time factor, omitting k = -1
  fdata$rel_fct <- factor(fdata$rel_month)
  fdata$rel_fct <- relevel(fdata$rel_fct, ref = as.character(-1))

  model <- lm(pct ~ rel_fct + t, data = fdata)

  tidy_out <- broom::tidy(model, conf.int = TRUE) |>
    filter(str_detect(term, "rel_fct")) |>
    mutate(
      k = as.integer(str_extract(term, "-?\\d+")),
      frame = frame_name
    )

  return(tidy_out)
}

all_frames <- names(frame_dictionaries)
es_results <- bind_rows(lapply(all_frames, function(f) run_event_study(frame_monthly, f)))

# Plot
ggplot(es_results, aes(x = k, y = estimate)) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed", linewidth = 0.3) +
  geom_vline(xintercept = 0, color = "black", linetype = "dotted", linewidth = 0.4) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
              alpha = 0.25, fill = "gray70") +
  geom_point(size = 1.4, color = "black") +
  geom_line(color = "black", linewidth = 0.4) +
  facet_wrap(~ frame, ncol = 2, scales = "free_y") +
  labs(
    x = "Months relative to ChatGPT launch (Dec 2022 = 0)",
    y = "Change in frame share (percentage points)"
  )
```

The frame-by-frame panels reveal substantial heterogeneity. The pre-launch
coefficients hover around zero in most panels, which is the visual signature
of parallel pre-trends and a precondition for interpreting post-launch
movements as effects of the shock rather than continuations of pre-existing
dynamics. After the ChatGPT date, several frames step up sharply and stay
elevated, while others move only modestly or not at all. The threat-leaning
frames (job loss, fear or resistance, inequality) show the largest and most
persistent post-launch elevations, while the opportunity-leaning frames (job
creation, productivity, transformation) move less and recover faster.
Skills and regulation activate too but on different time profiles, with
regulation responding most clearly around the EU AI Act discussions.

## Pre-Trend Validation

The visual impression of flat pre-launch coefficients is reassuring, but it
needs a formal test. A credible event-study design requires that the
pre-treatment trajectory be statistically indistinguishable from no effect;
otherwise post-launch movements could simply be the continuation of an
already-rising or already-falling trend that has nothing to do with the
shock. Table \@ref(tbl-pretrend-tests) reports an F-test for each frame in
which the joint null is that all pre-period event-time coefficients (from
month $-20$ through month $-2$) equal zero, conditional on a linear time
trend. A p-value above conventional thresholds (we use 0.10) is good news:
it means we cannot reject the null of flat pre-trends and the parallel-trends
assumption is not contradicted by the data.

```{r}
#| label: tbl-pretrend-tests
#| tbl-cap: "Pre-trend F tests by frame (H0: all pre-period coefficients equal zero)"

pretrend_tests <- lapply(all_frames, function(f) {

  fdata <- frame_monthly |>
    filter(frame == f, rel_month >= -20 & rel_month <= -2)

  if (nrow(fdata) < 5) return(NULL)

  fdata$rel_fct <- factor(fdata$rel_month)
  fdata$rel_fct <- relevel(fdata$rel_fct, ref = as.character(-2))

  model_full     <- lm(pct ~ rel_fct + t, data = fdata)
  model_restricted <- lm(pct ~ t, data = fdata)

  f_test <- anova(model_restricted, model_full)

  tibble(
    Frame     = f,
    F_stat    = round(f_test$F[2], 3),
    p_value   = round(f_test$`Pr(>F)`[2], 3),
    Pre_trend = ifelse(f_test$`Pr(>F)`[2] > 0.10, "No (good)", "Yes (concern)")
  )
})

pretrend_tbl <- bind_rows(pretrend_tests)
kable(pretrend_tbl) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

The fourth column flags each frame as either passing the pre-trend test (a
high p-value, labelled "No (good)" because we cannot detect a pre-trend) or
showing some evidence of a pre-trend ("Yes (concern)"). The expected outcome
is that most frames pass: a non-trivial pre-trend would mean that frame
prevalence was already drifting before the shock, which would complicate the
interpretation of post-shock movements. Where a frame fails the test, the
event-study plot for that frame should be read with caution and the
post-shock estimates treated as suggestive rather than definitive. In our
data, most frames pass at the 10% level; the few that do not still display
post-shock movements that are an order of magnitude larger than any pre-shock
drift, so the substantive interpretation is robust.

## Average Post-Treatment Effects

The event-study plots show the time path of each frame's response, but a
single summary number per frame is also useful for comparing magnitudes
across the eight frames at a glance. Table \@ref(tbl-ate-frames) reports the
average treatment effect (ATE) for each frame, defined as the coefficient on
a post-ChatGPT dummy in a simple regression of monthly frame share on the
dummy and a linear time trend. Standard errors are HAC-robust to handle
serial correlation in the monthly series. Frames are ordered by the absolute
size of the effect.

```{r}
#| label: tbl-ate-frames
#| tbl-cap: "Average post-ChatGPT effect on the prevalence of each frame, expressed in percentage points. Pre-mean and post-mean give the average frame share before and after the launch; ATE is the post-dummy coefficient from a regression with a linear time trend; standard errors are HAC-robust."

ate_results <- lapply(all_frames, function(f) {

  fdata <- frame_monthly |> filter(frame == f)

  model <- lm(pct ~ post + t, data = fdata)
  robust <- coeftest(model, vcov = vcovHAC(model))

  tibble(
    Frame        = f,
    Pre_mean     = round(mean(fdata$pct[fdata$post == 0]), 2),
    Post_mean    = round(mean(fdata$pct[fdata$post == 1]), 2),
    ATE_pp       = round(robust["post", 1], 2),
    SE           = round(robust["post", 2], 2),
    t_stat       = round(robust["post", 3], 2),
    p_value      = round(robust["post", 4], 4),
    Significant  = ifelse(robust["post", 4] < 0.05, "***",
                   ifelse(robust["post", 4] < 0.10, "*", ""))
  )
})

ate_tbl <- bind_rows(ate_results) |> arrange(desc(abs(ATE_pp)))

kable(ate_tbl) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

Figure \@ref(fig-ate-forest) presents the same information as a forest plot,
with each frame's ATE point estimate and 95% confidence interval shown on a
common horizontal scale. Frames are sorted from most negative to most
positive change in share.

```{r}
#| label: fig-ate-forest
#| fig-cap: "Average post-ChatGPT effects on frame shares (percentage points), with 95% confidence intervals. Frames are ordered by point estimate."
#| fig-height: 5

ate_plot_data <- bind_rows(ate_results) |>
  mutate(
    lower = ATE_pp - 1.96 * SE,
    upper = ATE_pp + 1.96 * SE
  )

ggplot(ate_plot_data, aes(x = ATE_pp, y = reorder(Frame, ATE_pp))) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black", linewidth = 0.3) +
  geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.25, color = "black",
                 linewidth = 0.5) +
  geom_point(size = 2.4, color = "black", shape = 16) +
  labs(
    x = "Change in frame share (percentage points)",
    y = NULL
  )
```

Two things stand out from the table and the forest plot. First, the
threat-leaning frames (job loss, fear or resistance, inequality) cluster on
the right side of the plot with positive and statistically significant
ATEs: their share of post-ChatGPT articles is meaningfully larger than
their share of pre-ChatGPT articles. Second, the opportunity-leaning frames
(job creation, productivity, transformation) move less, and some show
estimated ATEs near zero or even slightly negative. The skills frame
displays a moderate positive ATE, and regulation activates as well, though
its dynamics (driven heavily by EU AI Act discussion in 2024) deserve a
separate reading. The headline pattern is asymmetric activation: the
post-shock environment is dominated by frames that emphasise risk rather
than by frames that emphasise adaptation or opportunity.

## Asymmetry Between Threat and Opportunity Narratives

We expect the ChatGPT shock to activate threat narratives (job loss,
fear/resistance, inequality) more strongly than opportunity narratives (job
creation, productivity, transformation). We construct composite indices and
test this formally.

To formalise the asymmetric-activation pattern that emerges from the
frame-level results, we construct two composite indices. The *threat*
index aggregates the share of articles carrying any of the job-loss, fear,
or inequality frames; the *opportunity* index does the same for
job-creation, productivity, and transformation. Figure
\@ref(fig-threat-vs-opportunity) plots both indices over the sample period.

```{r}
#| label: fig-threat-vs-opportunity
#| fig-cap: "Composite threat and opportunity indices over time. The threat index aggregates the share of articles carrying any of the job-loss, fear, or inequality frames; the opportunity index does the same for job-creation, productivity, and transformation. Thin lines show monthly observations; thick lines are LOESS smooths."
#| fig-height: 5

composite_monthly <- corpus_data |>
  group_by(year_month) |>
  summarise(
    n = n(),
    threat = sum(frame_JOB_LOSS | frame_FEAR_RESISTANCE | frame_INEQUALITY,
                 na.rm = TRUE) / n() * 100,
    opportunity = sum(frame_JOB_CREATION | frame_PRODUCTIVITY | frame_TRANSFORMATION,
                      na.rm = TRUE) / n() * 100,
    .groups = "drop"
  ) |>
  filter(!is.na(year_month)) |>
  pivot_longer(cols = c(threat, opportunity), names_to = "index", values_to = "pct") |>
  mutate(index = factor(index,
                        levels = c("threat", "opportunity"),
                        labels = c("Threat (job loss + fear + inequality)",
                                   "Opportunity (job creation + productivity + transformation)")))

ggplot(composite_monthly, aes(x = year_month, y = pct,
                              linetype = index, group = index)) +
  geom_line(linewidth = 0.4, alpha = 0.45, color = "gray45") +
  geom_smooth(method = "loess", span = 0.3, se = FALSE,
              linewidth = 1.0, color = "black") +
  geom_vline(xintercept = CHATGPT_DATE, linetype = "dotted",
             color = "black", linewidth = 0.4) +
  annotate("label", x = CHATGPT_DATE, y = max(composite_monthly$pct) * 0.92,
           label = "ChatGPT", size = 3, color = "black",
           fill = "white", label.size = 0.2) +
  scale_linetype_manual(values = c("solid", "dashed")) +
  labs(x = NULL, y = "% of articles")
```

The two series move broadly together before December 2022 — both rise and
fall at roughly the same pace, and the threat index sits modestly above the
opportunity index. After the ChatGPT date, the two indices diverge: the
threat index rises sharply and remains elevated, while the opportunity
index moves more modestly and recovers more quickly. The visual gap between
the two smoothed lines widens noticeably from December 2022 onward and
remains wide through the end of the sample.

Table \@ref(tbl-asymmetry-test) tests this gap formally. We define the
threat–opportunity *gap* as the difference between the two indices in each
month, and regress it on a post-ChatGPT dummy with a linear time trend and
HAC-robust standard errors.

```{r}
#| label: tbl-asymmetry-test
#| tbl-cap: "Test of asymmetric frame activation: the dependent variable is the threat–opportunity gap (in percentage points) and the post-ChatGPT coefficient measures how much the gap widens after the launch."

# Build monthly data for the test
asym_data <- corpus_data |>
  group_by(year_month) |>
  summarise(
    n = n(),
    threat = sum(frame_JOB_LOSS | frame_FEAR_RESISTANCE | frame_INEQUALITY,
                 na.rm = TRUE) / n() * 100,
    opportunity = sum(frame_JOB_CREATION | frame_PRODUCTIVITY | frame_TRANSFORMATION,
                      na.rm = TRUE) / n() * 100,
    .groups = "drop"
  ) |>
  filter(!is.na(year_month)) |>
  mutate(
    gap = threat - opportunity,
    post = as.integer(year_month >= CHATGPT_DATE),
    t = as.numeric(difftime(year_month, min(year_month), units = "days")) / 30
  )

# Test whether the gap widens post treatment
gap_model <- lm(gap ~ post + t, data = asym_data)
gap_robust <- coeftest(gap_model, vcov = vcovHAC(gap_model))

gap_tidy <- tibble(
  Term     = c("Intercept", "Post ChatGPT", "Time trend"),
  Estimate = round(gap_robust[, 1], 3),
  SE       = round(gap_robust[, 2], 3),
  t_stat   = round(gap_robust[, 3], 3),
  p_value  = format.pval(gap_robust[, 4], digits = 3)
)

kable(gap_tidy) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

The post-ChatGPT coefficient measures how much the threat–opportunity gap
widens at the launch, conditional on the linear trend. A positive and
statistically significant coefficient confirms the visual asymmetry: the
information environment becomes systematically more threat-oriented after
the shock, and not just by a fraction of a percentage point.

## Platform Heterogeneity

The database covers seven platform categories: web portals, Facebook,
Twitter (X), Instagram, YouTube, TikTok, Reddit, and public forums. If the
ChatGPT shock propagated uniformly through the digital ecosystem, the
treatment effect should be similar across platforms. If, by contrast,
platforms differ in editorial selection, audience composition, or response
speed, the effect should be heterogeneous in informative ways.

### Volume by Platform

```{r}
#| label: fig-platform-volume
#| fig-cap: "Monthly article volume by platform type. The dotted vertical line marks the ChatGPT launch (December 2022). The smooth line is a LOESS fit; vertical scales are platform-specific."
#| fig-height: 7

platform_monthly <- corpus_data |>
  filter(!is.na(SOURCE_TYPE)) |>
  count(year_month, SOURCE_TYPE) |>
  filter(!is.na(year_month))

# Keep platforms with enough data
keep_platforms <- platform_monthly |>
  group_by(SOURCE_TYPE) |>
  summarise(total = sum(n)) |>
  filter(total >= 30) |>
  pull(SOURCE_TYPE)

platform_monthly <- platform_monthly |>
  filter(SOURCE_TYPE %in% keep_platforms)

ggplot(platform_monthly, aes(x = year_month, y = n)) +
  geom_col(fill = "gray70", color = NA) +
  geom_smooth(method = "loess", se = FALSE, color = "black", linewidth = 0.7) +
  geom_vline(xintercept = CHATGPT_DATE, linetype = "dotted", color = "black",
             linewidth = 0.4) +
  facet_wrap(~ SOURCE_TYPE, scales = "free_y", ncol = 2) +
  scale_x_date(date_breaks = "6 months", date_labels = "%b\n%Y") +
  labs(x = NULL, y = "Articles per month") +
  theme(axis.text.x = element_text(size = 8))
```

The platform panels show the volume shock playing out at very different
amplitudes across platforms. Web portals exhibit a clear and sustained jump
at the launch. Twitter and YouTube also show pronounced post-shock peaks,
though their dynamics differ. Smaller platforms (TikTok, Reddit, forums)
show much noisier patterns, consistent with both lower baseline volume and
more idiosyncratic posting behaviour.

### Platform-Specific Treatment Effects

The descriptive figure suggests heterogeneity, but the formal test in Table
\@ref(tbl-platform-ate) makes the magnitudes comparable. For each platform
with at least 30 monthly observations, we estimate the same pre/post
specification used for the aggregate volume series, with HAC-robust standard
errors.

```{r}
#| label: tbl-platform-ate
#| tbl-cap: "ChatGPT treatment effect on article volume by platform. Pre- and post-monthly are average article counts before and after the launch. Ratio is post/pre. ATE is the post-dummy coefficient with HAC-robust standard error."

platform_ate <- lapply(keep_platforms, function(p) {

  pdata <- corpus_data |>
    filter(SOURCE_TYPE == p) |>
    count(year_month) |>
    filter(!is.na(year_month)) |>
    mutate(
      post = as.integer(year_month >= CHATGPT_DATE),
      t = as.numeric(difftime(year_month, min(year_month), units = "days")) / 30
    )

  if (nrow(pdata) < 6) return(NULL)

  model <- lm(n ~ post + t, data = pdata)
  robust <- tryCatch(
    coeftest(model, vcov = vcovHAC(model)),
    error = function(e) coeftest(model)
  )

  tibble(
    Platform     = p,
    Pre_monthly  = round(mean(pdata$n[pdata$post == 0]), 1),
    Post_monthly = round(mean(pdata$n[pdata$post == 1]), 1),
    Ratio        = round(mean(pdata$n[pdata$post == 1]) / max(mean(pdata$n[pdata$post == 0]), 0.1), 1),
    ATE          = round(robust["post", 1], 2),
    SE           = round(robust["post", 2], 2),
    p_value      = round(robust["post", 4], 3)
  )
})

platform_ate_tbl <- bind_rows(platform_ate) |> arrange(desc(Ratio))

kable(platform_ate_tbl) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

The platform-level ATEs confirm that the shock did not arrive evenly across
the digital ecosystem. Web portals show the largest absolute increase in
article volume; relative to baseline, several smaller platforms show even
larger ratios. Where standard errors permit conclusions, the post-ChatGPT
coefficient is positive on every active platform. The ranking is consistent
with the view that web portals carried the heaviest load of post-shock
coverage in absolute terms while social and short-form video platforms
amplified the relative jump.

## Outlet-Type Heterogeneity

A second cut at heterogeneity asks whether different *kinds* of outlets
respond differently to the same shock. We use a difference-in-differences
design comparing tabloid outlets (high baseline propensity for
sensationalism) against quality outlets as the control group, with the
threat frame (job loss + fear + inequality) as the outcome at the
article level. The identifying assumption is that, absent the ChatGPT
shock, tabloid and quality outlets would have followed parallel trends in
threat-frame usage. Time fixed effects at the year-month level absorb
common shocks. The coefficient of interest is the post × tabloid
interaction, which captures the differential response of tabloids relative
to quality outlets.

```{r}
#| label: tbl-did-outlet
#| tbl-cap: "Difference-in-differences estimates: threat frame (job loss + fear + inequality) as the outcome, tabloid outlets as the treatment group, quality outlets as the control. The model includes year-month fixed effects; standard errors are heteroskedasticity-robust."

did_data <- corpus_data |>
  filter(outlet_type %in% c("Tabloid", "Quality")) |>
  mutate(
    threat = as.integer(frame_JOB_LOSS | frame_FEAR_RESISTANCE | frame_INEQUALITY),
    tabloid = as.integer(outlet_type == "Tabloid")
  )

did_model <- feols(
  threat ~ post_chatgpt * tabloid | year_month,
  data = did_data,
  vcov = "hetero"
)

did_ct <- coeftable(did_model)
did_terms <- rownames(did_ct)
did_label <- function(x) {
  x <- gsub("post_chatgpt", "Post-ChatGPT", x)
  x <- gsub("tabloid", "Tabloid", x)
  x <- gsub(":", " × ", x)
  x
}

did_tidy <- tibble(
  Term     = did_label(did_terms),
  Estimate = round(did_ct[, 1], 4),
  SE       = round(did_ct[, 2], 4),
  t_stat   = round(did_ct[, 3], 2),
  p_value  = format.pval(did_ct[, 4], digits = 3)
)

kable(did_tidy) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

The interaction coefficient (Post-ChatGPT × Tabloid) is positive and
statistically significant: tabloid outlets activated threat narratives more
strongly than quality outlets after the shock, even after netting out time
effects common to both. The main effect on the tabloid dummy reflects the
baseline difference in threat-frame use between tabloids and quality
outlets, which is positive but smaller than the post-shock differential.

## Persistence and Decay

A first-order question for any event-study finding is whether the post-shock
elevation persists or whether it fades back toward the pre-shock baseline as
the news cycle moves on. Mean reversion would suggest novelty-driven coverage
that wears off. Persistence would suggest a more durable change in how the
topic is discussed. Figure \@ref(fig-persistence) plots three-month rolling
averages of the threat and opportunity composite indices, smoothing out
month-to-month noise to make the medium-term trajectory visible.

```{r}
#| label: fig-persistence
#| fig-cap: "Three-month rolling averages of the threat and opportunity composite indices. The dotted vertical line marks the ChatGPT launch."
#| fig-height: 5

persistence_data <- corpus_data |>
  group_by(year_month) |>
  summarise(
    threat = sum(frame_JOB_LOSS | frame_FEAR_RESISTANCE | frame_INEQUALITY,
                 na.rm = TRUE) / n() * 100,
    opportunity = sum(frame_JOB_CREATION | frame_PRODUCTIVITY | frame_TRANSFORMATION,
                      na.rm = TRUE) / n() * 100,
    .groups = "drop"
  ) |>
  filter(!is.na(year_month)) |>
  arrange(year_month) |>
  mutate(
    threat_ma3 = zoo::rollmean(threat, k = 3, fill = NA, align = "right"),
    opportunity_ma3 = zoo::rollmean(opportunity, k = 3, fill = NA, align = "right")
  )

persistence_long <- persistence_data |>
  dplyr::select(year_month, threat_ma3, opportunity_ma3) |>
  pivot_longer(-year_month, names_to = "index", values_to = "pct") |>
  mutate(index = ifelse(str_detect(index, "threat"), "Threat", "Opportunity"))

ggplot(persistence_long, aes(x = year_month, y = pct,
                             linetype = index, group = index)) +
  geom_line(linewidth = 0.8, color = "black") +
  geom_vline(xintercept = CHATGPT_DATE, linetype = "dotted",
             color = "black", linewidth = 0.4) +
  annotate("label", x = CHATGPT_DATE,
           y = max(persistence_long$pct, na.rm = TRUE) * 0.92,
           label = "ChatGPT", fill = "white", color = "black",
           size = 3, label.size = 0.2) +
  scale_linetype_manual(values = c("Threat" = "solid", "Opportunity" = "dashed")) +
  labs(x = NULL, y = "% of articles (3-month moving average)")
```

The threat-index path moves up sharply at the shock and remains elevated for
the rest of the sample. The opportunity-index path also rises, but by less,
and recovers more rapidly toward its pre-shock level. The gap between the
two smoothed series is wide and durable rather than narrow and transient.
This is the visual signature of a structural shift rather than a novelty
spike.

### Post-Shock Decay Estimation

A more formal test of persistence regresses each frame's post-shock share on
months since the shock. A negative and significant slope would indicate
gradual decay back toward baseline; a positive or insignificant slope would
indicate continuing growth or a stable plateau.

```{r}
#| label: tbl-decay
#| tbl-cap: "Post-shock decay estimation. For each frame, we regress monthly post-ChatGPT share on months since the launch. A negative significant slope (\"Decaying\") would indicate fade-out; a positive significant slope (\"Growing\") indicates continued increase; otherwise the post-shock level is stable."

decay_results <- lapply(all_frames, function(f) {

  fdata <- frame_monthly |>
    filter(frame == f, post == 1) |>
    mutate(months_since = as.numeric(difftime(year_month, CHATGPT_DATE, units = "days")) / 30)

  if (nrow(fdata) < 4) return(NULL)

  model <- lm(pct ~ months_since, data = fdata)

  tibble(
    Frame         = f,
    Initial_level = round(coef(model)[1], 2),
    Monthly_decay = round(coef(model)[2], 3),
    p_value       = round(summary(model)$coefficients[2, 4], 3),
    Pattern       = case_when(
      coef(model)[2] > 0 & summary(model)$coefficients[2, 4] < 0.10 ~ "Growing",
      coef(model)[2] < 0 & summary(model)$coefficients[2, 4] < 0.10 ~ "Decaying",
      TRUE ~ "Stable (no sig. trend)"
    )
  )
})

decay_tbl <- bind_rows(decay_results)

kable(decay_tbl) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

# Robustness

## Alternative Event Windows

The pre/post estimates pool all available months on either side of the
launch. A natural concern is that the result depends on this choice. We
re-estimate the threat-frame ATE on progressively narrower symmetric windows
around the shock — six months on each side, then nine, twelve, fifteen, and
eighteen — to see whether the magnitude is stable as we approach the
discontinuity.

```{r}
#| label: tbl-window-robustness
#| tbl-cap: "Sensitivity of the threat-frame ATE to the choice of event window. Each row uses a symmetric window of the indicated width around the ChatGPT date."

windows <- c(6, 9, 12, 15, 18)

window_results <- lapply(windows, function(w) {

  wdata <- corpus_data |>
    filter(
      DATE >= (CHATGPT_DATE - months(w)),
      DATE <= (CHATGPT_DATE + months(w))
    ) |>
    group_by(year_month) |>
    summarise(
      threat_pct = sum(frame_JOB_LOSS | frame_FEAR_RESISTANCE | frame_INEQUALITY,
                       na.rm = TRUE) / n() * 100,
      .groups = "drop"
    ) |>
    filter(!is.na(year_month)) |>
    mutate(
      post = as.integer(year_month >= CHATGPT_DATE),
      t = row_number()
    )

  model <- lm(threat_pct ~ post + t, data = wdata)
  robust <- tryCatch(
    coeftest(model, vcov = vcovHAC(model)),
    error = function(e) coeftest(model)
  )

  tibble(
    Window_months = paste0("±", w),
    ATE = round(robust["post", 1], 2),
    SE  = round(robust["post", 2], 2),
    p   = round(robust["post", 4], 3)
  )
})

window_tbl <- bind_rows(window_results)

kable(window_tbl) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

The estimated ATE is broadly stable as the window narrows. Some attenuation
is expected when the window approaches the discontinuity (fewer observations
mean noisier estimates), but the sign and approximate magnitude are
preserved across all five windows. The threat-frame finding is therefore not
an artefact of the choice of estimation window.

## Poisson Count Model

Monthly article counts are non-negative integers, so OLS is technically
inappropriate even though it is convenient. As a check, we re-estimate the
volume specification as a quasi-Poisson regression. Quasi-Poisson allows for
overdispersion (variance larger than the mean), which is common in count
data with extra-Poisson volatility, and is therefore a more appropriate
match for monthly article counts than a strict Poisson model. The dispersion
parameter is reported alongside the regression coefficients.

```{r}
#| label: tbl-poisson
#| tbl-cap: "Quasi-Poisson regression of monthly article counts on a post-ChatGPT dummy and a linear time trend. IRR is the incidence rate ratio (exp(coefficient)). The dispersion parameter is reported in the bottom row; values much greater than one indicate overdispersion relative to a strict Poisson model."

poisson_model <- glm(n ~ post + t, data = monthly, family = poisson())

# Quasi-Poisson for overdispersion
qpoisson_model <- glm(n ~ post + t, data = monthly, family = quasipoisson())

dispersion_value <- summary(qpoisson_model)$dispersion

poisson_tidy <- broom::tidy(qpoisson_model, conf.int = TRUE) |>
  mutate(
    IRR = round(exp(estimate), 2),
    across(c(estimate, std.error, statistic), ~ round(.x, 3)),
    p.value = format.pval(p.value, digits = 3)
  ) |>
  dplyr::select(term, estimate, std.error, statistic, p.value, IRR)

dispersion_row <- tibble(
  term      = "Dispersion parameter",
  estimate  = round(dispersion_value, 2),
  std.error = NA_real_,
  statistic = NA_real_,
  p.value   = NA_character_,
  IRR       = NA_real_
)

poisson_full <- bind_rows(poisson_tidy, dispersion_row)

kable(poisson_full) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

The post-ChatGPT incidence rate ratio (the IRR in the second row of the
table) translates the level shift into a multiplicative factor on the
monthly count. The estimate confirms the OLS finding qualitatively: the
post-shock environment is characterised by a multiple-fold increase in
monthly article volume, and the result is statistically significant. The
dispersion parameter is well above one, justifying the quasi-Poisson
specification over a strict Poisson model.

## Placebo Test

The structural-break test rejected no-break against an unknown alternative
and located the break at December 2022. As a complementary check, we apply
the same level-shift specification at December 2021 — exactly one year before
the actual shock and far from any salient AI event. If the estimated effect
were a seasonal artefact (for example, an end-of-year coverage spike), the
placebo regression should pick it up. If the effect is specific to the
ChatGPT launch, the placebo coefficient should be small and statistically
indistinguishable from zero.

```{r}
#| label: tbl-placebo
#| tbl-cap: "Placebo test at December 2021 (one year before the actual ChatGPT launch). Estimation is restricted to months strictly before the actual shock so that the placebo dummy is not contaminated by the true treatment."

PLACEBO_DATE <- as.Date("2021-12-01")

placebo_monthly <- monthly |>
  filter(year_month < CHATGPT_DATE) |>
  mutate(
    placebo_post = as.integer(year_month >= PLACEBO_DATE),
    t_placebo = row_number()
  )

if (nrow(placebo_monthly) >= 6) {
  placebo_model <- lm(n ~ placebo_post + t_placebo, data = placebo_monthly)
  placebo_robust <- tryCatch(
    coeftest(placebo_model, vcov = vcovHAC(placebo_model)),
    error = function(e) coeftest(placebo_model)
  )

  placebo_tidy <- tibble(
    Term     = c("Intercept", "Placebo post", "Time trend"),
    Estimate = round(placebo_robust[, 1], 2),
    SE       = round(placebo_robust[, 2], 2),
    p_value  = format.pval(placebo_robust[, 4], digits = 3)
  )

  kable(placebo_tidy) |>
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
} else {
  cat("Insufficient pre period data for placebo test.\n")
}
```

The placebo coefficient is small and far from statistical significance. The
real ChatGPT effect, by contrast, is large and precise. The contrast is the
expected pattern under a clean event-study identification: the actual launch
date carries information that the placebo date does not.

# Summary of Findings

```{r}
#| label: tbl-summary-findings
#| tbl-cap: "Summary of main results"

# Compile key numbers
n_total <- nrow(corpus_data)
n_pre <- sum(corpus_data$post_chatgpt == 0)
n_post <- sum(corpus_data$post_chatgpt == 1)
volume_ratio <- round(n_post / max(n_pre, 1) *
  (as.numeric(difftime(CHATGPT_DATE, min(corpus_data$DATE), units = "days")) /
   as.numeric(difftime(max(corpus_data$DATE), CHATGPT_DATE, units = "days"))), 1)

findings <- tibble(
  Finding = c(
    "Total corpus size",
    "Pre treatment articles",
    "Post treatment articles",
    "Volume multiplier (monthly average ratio)",
    "Structural break detected at ChatGPT date",
    "Frames with significant positive ATE",
    "Frames with significant negative ATE",
    "Threat vs opportunity gap widens post shock"
  ),
  Result = c(
    format(n_total, big.mark = ","),
    format(n_pre, big.mark = ","),
    format(n_post, big.mark = ","),
    paste0(volume_ratio, "x"),
    ifelse(exists("fs_test"),
           ifelse(sctest(fs_test)$p.value < 0.05, "Yes (p < 0.05)", "No"),
           "Not tested"),
    paste(ate_tbl$Frame[ate_tbl$p_value < 0.05 & ate_tbl$ATE_pp > 0], collapse = ", "),
    paste(ate_tbl$Frame[ate_tbl$p_value < 0.05 & ate_tbl$ATE_pp < 0], collapse = ", "),
    ifelse(as.numeric(gap_robust["post", 4]) < 0.10, "Yes", "No")
  )
)

kable(findings) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

# Discussion

The results tell a consistent story. Aggregate media attention to the
AI–labour intersection breaks sharply at the ChatGPT launch. The Sup-F
structural-break test confirms that December 2022 is statistically
distinguishable from a smooth trend, and the placebo at December 2021 rules
out a seasonal explanation. At the frame level, threat-related content (job
loss, fear or resistance, inequality) rises faster and fades more slowly than
opportunity-related content (job creation, productivity, transformation), and
the composite gap between the two indices widens significantly in the
post-shock period. Tabloid outlets activate threat narratives more strongly
than quality outlets, and the shock propagated unevenly across web portals
and social media platforms.

From a labour-economics perspective, the asymmetric activation suggests
media-amplified uncertainty about automation [@acemoglu2018race; @autor2022work].
Croatian workers, firms, and policymakers receive a pessimistic signal about
the labour-market consequences of generative AI that is hard to square with
either task-level exposure estimates [@eloundou2023gpts; @felten2021ai] or
early productivity field experiments [@brynjolfsson2023generative]. The
persistence of threat-frame elevation in our rolling-window and decay
regressions also rules out a simple "novelty wears off" story: what we observe
looks more like a durable shift in how AI and work are discussed than a
short-lived spike.

The policy implication is direct. When displacement framing rises faster and
more persistently than skills or reskilling framing, the information
environment delivers risk signals more readily than adjustment signals. That
imbalance is measurable, and it is something Croatian active-labour-market
policy and adult education communication could address. Without firm-level
data on AI-driven hiring or automation investment, a documented shift in the
balance of signals reaching workers is still a meaningful input for
policymakers designing retraining and reskilling programmes.

The analysis has four limitations. First, frames are detected via Croatian
keyword dictionaries with stem-based regular expressions and are not validated
against a human-coded gold standard, so per-frame precision and recall are
unmeasured. The diagnostic columns built upstream in `R/02_add_diagnostics.R`
(`groupA_hit_count`, `groupB_hit_count`, `groupA_context`, `groupB_context`)
show which keywords matched and where, which provides some transparency but
does not eliminate the concern. We note, however, that dictionary measurement
error is plausibly orthogonal to the ChatGPT date: it would attenuate
estimated effects toward zero rather than create spurious post-shock
activation, so the *direction* of the findings is unlikely to be an artifact
of the detection method. Second, the corpus records media supply rather than
media consumption, and we cannot directly measure whether readers were exposed
to or influenced by the framing patterns we document. Third, the Determ
AUTO_SENTIMENT field has unknown labelling methodology and is accordingly
treated as a robustness check rather than a primary outcome. Fourth, the
analysis covers Croatian-language media only; replication on Slovenian,
Serbian, or Bosnian press would be a natural next step.

Three directions stand out for future research. Linking the framing series to
Croatian Employment Service (HZZ) monthly vacancy and registered-unemployment
data at the occupational level would take the analysis from supply-side
measurement to direct estimation of how the information environment co-moves
with actual labour-market behaviour. Replicating the design on Slovenian,
Serbian, or Bosnian press would test whether the asymmetric-activation pattern
is specific to Croatia or broader across the region. An occupation-level
extension mapping which jobs draw the most post-shock narrative attention, and
how that compares to task-level AI exposure indices, would address whether
media coverage accurately reflects occupational risk or systematically
misdirects worker attention.

<!--
COLLABORATOR NOTE: the Discussion is intentionally a first pass.
Two areas would benefit from your domain knowledge:
  (i)  Policy interpretation. Sharpen paragraph 3 with concrete Croatian
       active-labour-market policy levers (HZZ retraining schemes, the
       National Plan for the Development of Lifelong Learning, EU AI Act
       implementation timeline). The current draft is deliberately generic.
  (ii) Limitations framing. The dictionary-detection caveat is the most
       likely reviewer pressure point. If you can describe (a) any spot
       check you have done on the diagnostic columns, or (b) a brief sanity
       check on a small random sample, that would strengthen paragraph 4(i).
-->


# Technical Appendix

```{r}
#| label: session-info

sessionInfo()
```
